---
title: "Preparing the data for estimating volumes"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
execute:
  error: true
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    keep-tex: true
    pdf-engine: lualatex
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

\newcommand{\ie}{*i.e.,*}
\newcommand{\F}{\mathscr{F}}
\newcommand{\N}{\mathbfscr{N}}

## Introduction

The data are provided in @Didion2024. After talking to the main author (Markus Didion), it is recommended to use the variable `V_finewoodytotal` instead of `V_finebranch` + estimate of tree top, as the definition of `V_finebranch` was not always stable (sometimes set to 0 despite there are small branches, and reported to `V_finewoodytotal`).

Given that there is no bole volume provided and that the authors leave the choice of the method to the users, I decided to use the package `TapeS`. Originally I was using `TapeR`, but, after discussion, Markus Didion recommended the newer package which is species-specific, while `TapeR` was only trained on *Abies alba*. Surprisingly, `TapeR` gave results quite close to simply piling up cones or cylinders! `TapeS` is also very similar to `TapeR` in estimations too. Note that I use the version `0.14.1` of `TapeS`.

## Prepare the data

### Loading the data

The following packages are required to load and handle the data and print summary tables:
```{r}
#| output: false
#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(stringi)
library(terra)
library(TapeS)
library(gt)
```

To load the data, more details are provided in `01_prepare_french_data.qmd`

```{r}
#### Prepare data
## Loading
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
	path_data = paste0(mnt_point, "data_orig/")
	if (!dir.exists(path_data))
		stop(paste0("Folder <", path_data, "> does not exist! Check mounting point and mounted folder"))
} else if (os == "Windows") {
	path_data = "//Del1509n015/2024_FairCarbon/data_orig/"
	if (!dir.exists(path_data))
		stop(paste0("Folder <", path_data, "> does not exist!"))
} else {
	stop(paste("Unknown Operating System:", os))
}

# Swiss data
filename_switzerland = paste0(path_data, "switzerland/stembranchchwood_2024-04-25.txt")
if (!file.exists(filename_switzerland))
	stop(paste0("The file <", filename_switzerland, "> does not exist"))

swiss_dt = fread(filename_switzerland)
```

### Adding species informations
Species have been prepared in `01_prepare_french_data.qmd` accordingly with the Swiss species code.

```{r}
#### Add species to swiss_dt
## Load species_dt, in order to add species' name
species_dt = readRDS("/mnt/local_share/data/all_species.rds")

## Add species scientific names
swiss_dt = merge.data.table(x = swiss_dt,
	y = species_dt[!is.na(sp_code_swiss), .(sp_code_swiss, speciesName_sci, genus = genre, fct_type)],
	by.x = "TreeSpecies", by.y = "sp_code_swiss")
```

### Extract coordinates to create the point shapefile (`SpatVector`)

Extract all the informations at the plot level, such as structure of the stand (even/uneven aged), main species, elevation and coordinates, ...

```{r}
#### Extract info plots
## Keep columns of interest
coords_swiss = unique(swiss_dt[, .(SiteID, Lat, Long, InvYear, Elevation, NFI_PR, StandAge, StandComp)])
swiss_dt[, c("Elevation", "NFI_PR", "StandAge", "StandComp", "TreeAge", "NFI_mainspecies",
	"Lat", "Long", "DBH", "H_total") := NULL]

setnames(coords_swiss, new = c("plot_id", "lat", "lon", "year", "altitude", "region", "stand_age", "composition"))

## Change modalities, based on Didion et al. 2024
coords_swiss[, .N, by = stand_age] # I guess 0 is even-aged
coords_swiss[, .N, by = composition][order(composition)] # I guess 1 is pure, 2 is confi 75%, 3 is broadleaf 75

# Stand age
coords_swiss[, stand_age := as.character(stand_age)]
coords_swiss[stand_age == "0", stand_age := "even"]
coords_swiss[stand_age == "1", stand_age := "uneven"]

# Composition
coords_swiss[, composition := as.character(composition)]
coords_swiss[composition == "1", composition := "pure"]
coords_swiss[composition == "2", composition := "conifer_75"]
coords_swiss[composition == "3", composition := "broadleaf_75"]
coords_swiss[composition == "4", composition := "mixed"]

## Correction of the plot id, leading 0 were removed when loading
coords_swiss[, plot_id := stri_pad(str = plot_id, width = 8, pad = "0")]

## Create SpatVector
coords_swiss = vect(x = coords_swiss, crs = "EPSG:4326") # There are doublons in terms of (x, y) coords of courses!
```

### Reshaping the data

To use the package `TapeS`, data must be reshaped in the 'melt' format rather than large format. There are also intermediate diameter--height combinations to add, such as the dbh, the 7 cm diameter, and the top diameter.
