---
title: "Preparing the data for estimating volumes"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
execute:
  error: true
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    keep-tex: true
    pdf-engine: lualatex
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

\newcommand{\ie}{*i.e.,*}
\newcommand{\F}{\mathscr{F}}
\newcommand{\N}{\mathbfscr{N}}

## Introduction

The data are provided in @Didion2024. After talking to the main author (Markus Didion), it is recommended to use the variable `V_finewoodytotal` instead of `V_finebranch` + estimate of tree top, as the definition of `V_finebranch` was not always stable (sometimes set to 0 despite there are small branches, and reported to `V_finewoodytotal`).

Given that there is no bole volume provided and that the authors leave the choice of the method to the users, I decided to use the package `TapeS`. Originally I was using `TapeR`, but, after discussion, Markus Didion recommended the newer package which is species-specific, while `TapeR` was only trained on *Abies alba*. Surprisingly, `TapeR` gave results quite close to simply piling up cones or cylinders! `TapeS` is also very similar to `TapeR` in estimations too. Note that I use the version `0.14.1` of `TapeS`.

## Prepare the data

### Loading the data

The following packages are required to load and handle the data and print summary tables:
```{r}
#| output: false
#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(stringi)
library(terra)
library(TapeS)
library(gt)
```

To load the data, more details are provided in `01_prepare_french_data.qmd`

```{r}
#### Prepare data
## Loading
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
	path_data = paste0(mnt_point, "data_orig/")
	if (!dir.exists(path_data))
		stop(paste0("Folder <", path_data, "> does not exist! Check mounting point and mounted folder"))
} else if (os == "Windows") {
	path_data = "//Del1509n015/2024_FairCarbon/data_orig/"
	if (!dir.exists(path_data))
		stop(paste0("Folder <", path_data, "> does not exist!"))
} else {
	stop(paste("Unknown Operating System:", os))
}

# Swiss data
filename_switzerland = paste0(path_data, "switzerland/stembranchchwood_2024-04-25.txt")
if (!file.exists(filename_switzerland))
	stop(paste0("The file <", filename_switzerland, "> does not exist"))

swiss_dt = fread(filename_switzerland)
```

### Adding species informations
Species have been prepared in `01_prepare_french_data.qmd` accordingly with the Swiss species code.

```{r}
#### Add species to swiss_dt
## Load species_dt, in order to add species' name
species_dt = readRDS("/mnt/local_share/data/all_species.rds")

## Add species scientific names
swiss_dt = merge.data.table(x = swiss_dt,
	y = species_dt[!is.na(sp_code_swiss), .(sp_code_swiss, speciesName_sci, genus = genre, fct_type)],
	by.x = "TreeSpecies", by.y = "sp_code_swiss")
```

### Extract coordinates to create the point shapefile (`SpatVector`)

Extract all the informations at the plot level, such as structure of the stand (even/uneven aged), main species, elevation and coordinates, ...

```{r}
#### Extract info plots
## Keep columns of interest
coords_swiss = unique(swiss_dt[, .(SiteID, Lat, Long, InvYear, Elevation, NFI_PR, StandAge, StandComp)])
swiss_dt[, c("Elevation", "NFI_PR", "StandAge", "StandComp", "TreeAge", "NFI_mainspecies",
	"Lat", "Long") := NULL]

setnames(coords_swiss, new = c("plot_id", "lat", "lon", "year", "altitude", "region", "stand_age", "composition"))

## Change modalities, based on Didion et al. 2024
coords_swiss[, .N, by = stand_age] # I guess 0 is even-aged
coords_swiss[, .N, by = composition][order(composition)] # I guess 1 is pure, 2 is confi 75%, 3 is broadleaf 75

# Stand age
coords_swiss[, stand_age := as.character(stand_age)]
coords_swiss[stand_age == "0", stand_age := "even"]
coords_swiss[stand_age == "1", stand_age := "uneven"]

# Composition
coords_swiss[, composition := as.character(composition)]
coords_swiss[composition == "1", composition := "pure"]
coords_swiss[composition == "2", composition := "conifer_75"]
coords_swiss[composition == "3", composition := "broadleaf_75"]
coords_swiss[composition == "4", composition := "mixed"]

## Correction of the plot id, leading 0 were removed when loading
coords_swiss[, plot_id := stri_pad(str = plot_id, width = 8, pad = "0")]

## Create SpatVector
coords_swiss = vect(x = coords_swiss, crs = "EPSG:4326") # There are doublons in terms of (x, y) coords of courses!
```

### Reshaping the data

To use the package `TapeS`, data must be reshaped in the 'melt' format rather than large format. There are also intermediate diameter--height combinations to add, such as the dbh, the 7 cm diameter, and the top diameter.

```{r}
#### Reshape data
## Melt
swiss_dt = melt(swiss_dt, measure = patterns("^DM[[:digit:]]"), variable.name = "diameter_height",
	value.name = "diameter", variable.factor = FALSE)

## Define new variables, convert some to m, and rename some columns
swiss_dt[, c("diameter", "dbh_m", "height", "L_coarsestem", "L_coarsestemfinal", "L_top",
		"D_coarsestemfinal", "D_top") :=
	.(diameter/1e3, DBH/1e3, H_total/10, L_coarsestem/10, L_coarsestemfinal/10, L_top/10,
		D_coarsestemfinal/1e3, D_top/1e3)]

swiss_dt[, c("DBH", "H_total") := NULL]

# Convert volumes to cubic metres
swiss_dt[, c("V_coarsebranch", "V_finebranch", "V_finewoodytotal") :=
	.(V_coarsebranch/1e3, V_finebranch/1e3, V_finewoodytotal/1e3)]

setnames(swiss_dt, old = c("SiteID", "InvYear", "TreeID", "TreeSpecies"),
	new = c("plot_id", "year", "tree_id", "sp_code"))

## Extract heights of dbh measurements
swiss_dt[diameter_height == "DM065", diameter_height_m := 0.65]
swiss_dt[diameter_height != "DM065",
	diameter_height_m := as.numeric(stri_replace(str = diameter_height, regex = "DM", replacement = ""))]

## Plot_id do not correspond to what is written in Didion et al. 2024 (int conversion => lost leading 0)
swiss_dt[, plot_id := stri_pad(str = plot_id, width = 8, pad = "0")]

## Remove useless lines
swiss_dt = swiss_dt[!is.na(diameter)] # Diameters that do not exist or where not given

# Check that sp code is unique per tree_id, remove unknown species (NAs)
range(swiss_dt[, length(unique(sp_code)), by = tree_id][, V1])
swiss_dt = swiss_dt[!is.na(sp_code)] # Missing species information

## Sort by tree id, and then diameter height (tree_id is already a unique identifier, no need for plot_id)
setkey(swiss_dt, tree_id, diameter_height_m)

## Remove individuals with is.na(D_coarsestemfinal) and L_coarsestemfinal > 0
rm_indiv = swiss_dt[is.na(D_coarsestemfinal) & (L_coarsestemfinal > 0), unique(tree_id)]
swiss_dt = swiss_dt[!(.(rm_indiv))]

## Remove individuals with is.na(D_top)
rm_indiv = swiss_dt[is.na(D_top), unique(tree_id)]
swiss_dt = swiss_dt[!(.(rm_indiv))]

## Remove individuals that are not fully measured (i.e., missing V_finewoodytotal and V_coarsebranch)
swiss_dt = swiss_dt[!is.na(V_finewoodytotal) & !is.na(V_coarsebranch)] # NAs != 0 as specified in EFM article
swiss_dt[, length(unique(tree_id))] # 16437 individuals

## Add three lines per tree...
tree_level = unique(swiss_dt, by = "tree_id")

# ...one for diameter at breast height at 1.3 m
diam130 = tree_level[, .(
	plot_id, year, tree_id, sp_code,
	L_coarsestem, L_coarsestemfinal, L_top,
	D_coarsestemfinal, D_top,
	V_coarsebranch, V_finebranch, V_finewoodytotal,
	diameter_height = "DM1.3",
	diameter = dbh_m,
	dbh_m, height,
	diameter_height_m = 1.3,
	speciesName_sci,  genus, fct_type
)]

# ...one for diameter 7 at L_coarsestem
diam007 = tree_level[, .(
	plot_id, year, tree_id, sp_code,
	L_coarsestem, L_coarsestemfinal, L_top,
	D_coarsestemfinal, D_top,
	V_coarsebranch, V_finebranch, V_finewoodytotal,
	diameter_height = "DM_diam7",
	diameter = 0.07,
	dbh_m, height,
	diameter_height_m = L_coarsestem,
	speciesName_sci,  genus, fct_type
)]

pb = diam007[L_coarsestem == 0, tree_id]
swiss_dt[.(pb), all(L_top == height)] # So all this trees have L_top equals to height
rm(pb)

diam007 = diam007[L_coarsestem > 0] # Otherwise it would put 7 cm at height 0!

# ...one for diameter D_top at height - L_top/2
diamTop = tree_level[, .(
	plot_id, year, tree_id, sp_code,
	L_coarsestem, L_coarsestemfinal, L_top,
	D_coarsestemfinal, D_top,
	V_coarsebranch, V_finebranch, V_finewoodytotal,
	diameter_height = "DM_diamTop",
	diameter = D_top,
	dbh_m, height,
	diameter_height_m = height - L_top / 2,
	speciesName_sci,  genus, fct_type
)]

## Bind, clean and sort data
swiss_dt = rbindlist(list(swiss_dt, diam130, diam007, diamTop), use.names = TRUE)

# Remove individuals with max(diameter) below 7 cm (bole volume does not exist for them)
max_diam = swiss_dt[, .(no_bole = max(diameter) <= 0.07), by = tree_id]
pb = max_diam[(no_bole), tree_id]

swiss_dt = swiss_dt[!(tree_id %in% pb)]

# Set key
setkey(swiss_dt, sp_code)

## Compute number of individuals per species
n_indiv = unique(swiss_dt[, .(sp_code, speciesName_sci), by = tree_id])
n_indiv = n_indiv[, .N, by = .(sp_code, speciesName_sci)][order(-N)]
n_indiv[, sum(N)] # 16250 individuals
```

## Compute bole volumes and estimate taper break height (hdec)

As said in the introduction, bole volumes are not provided. I estimate them using the package `TapeS`, verion 0.14.1. The documentation can be found [here](https://cran.r-project.org/web/packages/TapeS/refman/TapeS.html), and I used the [vignette](https://cran.r-project.org/web/packages/TapeS/vignettes/tapes.html) as a guide; for instance it seems that the diameters must be in cm, and heights in m, although it is speciefied nowhere in the documentation. The package is based on @Kublin2013 and @Vonderach2018.

The taper break height, hdec (for "hauteur de découpe", French NFI wording, or "hauteur de décrochement", Emerge wording) corresponds to the height at which a sharp decrease (at least 10%) in diameter occurs. This measurement seems to be specific to the French NFI and does not exist in the Oudin's protocol [@Oudin1930], nor does it exist in the EFM [@Didion2024].

I do not need hdec so far, but a joint model (multivariate distribution or copula, see @Nelsen2006) would require hdec as the bole volume formula requires it!

```{r}
#### Compute bole volume using TapeS package (https://link.springer.com/article/10.1007/s10342-013-0715-0)
## List species parametrised in TapeS https://cran.r-project.org/web/packages/TapeS/index.html
ls_tapes = tprSpeciesCode(inSp = NULL, outSp = NULL)
setDT(ls_tapes)
ls_tapes = ls_tapes[, .(ID, scientific)]

# Correction in the package, wrong leading space
ls_tapes[scientific == " Picea sitchensis", scientific := "Picea sitchensis"]

setorder(ls_tapes, scientific)

# List problematic species
sp_pb = n_indiv[!(speciesName_sci %in% ls_tapes[, scientific]), speciesName_sci]
sp_pb_genus = unique(stri_sub(str = sp_pb, to = stri_locate_first(str = sp_pb, regex = " ")[, "start"] - 1))

for (sp in sp_pb_genus)
{
	print(sp_pb[stri_detect(str = sp_pb, regex = sp)])
	print(ls_tapes[stri_detect(str = scientific, regex = sp)])
}

# Assumptions and names modifications (in ls_tapes rather than swiss_dt of course!)
# I assume that Fraxinus sp. in Switzerland is Fraxinus excelsior,
# https://www.wsl.ch/lud/portree/_sppDoc/Fraxinus_excelsior_v2.pdf
ls_tapes[scientific == "Fraxinus excelsior", scientific := "Fraxinus sp."]

# I assume that Populus spp. applies to Populus tremula in Switzerland given that balsamifera ruled out
ls_tapes[scientific == "Populus spp.", scientific := "Populus tremula"]

# I assume that Betula spp. applies to Betula pendula in Switzerland
ls_tapes[scientific == "Betula spp.", scientific := "Betula pendula"]

# I assume that Ulmus spp. applies to Ulmus glabra in Switzerland
ls_tapes[scientific == "Ulmus spp.", scientific := "Ulmus glabra"]

# List species that are not referenced in the TapeS package
sp_pb = n_indiv[!(speciesName_sci %in% ls_tapes[, scientific]), speciesName_sci]
# sp_pb = "Pinus cembra", "Pinus mugo", "Chamaecyparis sp.", "Juglans sp.", "Picea sp." 

n_indiv[speciesName_sci %in% sp_pb, sum(N)]
n_indiv = n_indiv[!(speciesName_sci %in% sp_pb)]

setkey(ls_tapes, scientific)
setkey(n_indiv, speciesName_sci)

## Fit taper model to estimate tree profile and compute bole volumes (threshold diameter = 7 cm)
bole_volume_dt = unique(swiss_dt[, .(sp_code, tree_id, height)])

bole_volume_dt[, eval_h := .(list(seq(2, max(height), 1))), by = tree_id]

bole_volume_dt[, bole_volume_m3 := -Inf]
bole_volume_dt[, pred_diam_m := as.list(NULL)]

setkey(bole_volume_dt, sp_code, tree_id)

swiss_dt[D_top == 0, length(unique(tree_id))] # 416 concerned individuals

# Fit for species that are parametrised in TapeS
for (sp in n_indiv[, speciesName_sci])
{
	print(paste("Running species", sp))
	tapes_id = ls_tapes[sp, ID]

	temp = swiss_dt[.(n_indiv[sp, sp_code])]
	temp = temp[diameter > 0] # Sometimes D_top == 0...
	setorder(temp, tree_id, diameter_height_m)

	obj = tprTrees(spp = rep(tapes_id, n_indiv[sp, N]),
		Hm = temp[, .(diameter_height_m = list(diameter_height_m)), by = tree_id][, diameter_height_m],
		Dm = temp[, .(diameter = list(100*diameter)), by = tree_id][, diameter], #! Diameter in cm
		Ht = temp[, .(height = unique(height)), by = tree_id][, height])

	# Compute bole volume based on taper functions
	bv = tprVolume(obj, AB = list(A = 0, B = 7), iAB = c("h", "dob"), bark = TRUE)
	bole_volume_dt[.(n_indiv[sp, sp_code]), bole_volume_m3 := bv]

	# Predict diameters every metre (matrix filled with 0 when beyond tree height) and convert into metres
	pred_diam_mat = tprDiameter(obj = obj, Hx = seq(2, max(Ht(obj)), by = 1), bark = TRUE,
		interval = "none", cp = TRUE, mono = TRUE, Rfn = NULL)/100 # Start at 2 m, to avoid buttlog

	# Remove trailing zeros and coerce to list
	if (n_indiv[.(sp), N] > 1) # It means there is more than one individual (pb Larix kaempferi)
		diam_ls = apply(pred_diam_mat, 1, function(x) x[seq_len(max(which(x != 0)))])
	
	if (n_indiv[.(sp), N] == 1) # It means there there is only one individual (pb Larix kaempferi)
		diam_ls = list(pred_diam_mat)

	# Add to bole volume
	bole_volume_dt[.(n_indiv[sp, sp_code]), pred_diam_m := diam_ls]
}

# Check NAs
bole_volume_dt[, any(is.na(bole_volume_m3))] # FALSE

# Check the non computed volumes (only non parametrised species)
pb = bole_volume_dt[is.infinite(bole_volume_m3), unique(sp_code)]
n_indiv[, any(sp_code %in% pb)] # In other words, -Inf only for species not listed, such as Pinus cembra...

# Fit non-parametrised species using TapeS' generic models, Coniferales and Magnoliopsida trees, resp.
sp_pb = unique(swiss_dt[.(pb), .(speciesName_sci, fct_type, .N), by = sp_code])
setkey(sp_pb, speciesName_sci)

for (sp in sp_pb[, speciesName_sci])
{
	print(paste("Running species", sp))
	if (sp_pb[sp, fct_type] == "conifer")
		tapes_id = ls_tapes["Coniferales trees", ID]

	if (sp_pb[sp, fct_type] == "broadleaf")
		tapes_id = ls_tapes["Magnoliopsida trees", ID]

	temp = swiss_dt[.(sp_pb[sp, sp_code])]
	temp = temp[diameter > 0] # Sometimes D_top == 0...
	setorder(temp, tree_id, diameter_height_m)

	obj = tprTrees(spp = rep(tapes_id, temp[, length(unique(tree_id))]),
		Hm = temp[, .(diameter_height_m = list(diameter_height_m)), by = tree_id][, diameter_height_m],
		Dm = temp[, .(diameter = list(100*diameter)), by = tree_id][, diameter], #! Diameter in cm
		Ht = temp[, .(height = unique(height)), by = tree_id][, height])

	# Compute bole volume based on taper functions
	bv = tprVolume(obj, AB = list(A = 0, B = 7), iAB = c("h", "dob"), bark = TRUE)
	bole_volume_dt[.(sp_pb[sp, sp_code]), bole_volume_m3 := bv]

	# Predict diameters every metre (matrix filled with 0 when beyond tree height) and convert into metres
	pred_diam_mat = tprDiameter(obj = obj, Hx = seq(2, max(Ht(obj)), by = 1), bark = TRUE,
		interval = "none", cp = TRUE, mono = TRUE, Rfn = NULL)/100 # Start at 2 m, to avoid buttlog

	# Remove trailing zeros and coerce to list
	if (sp_pb[.(sp), N] > 1) # It means there is more than one individual (pb Larix kaempferi)
		diam_ls = apply(pred_diam_mat, 1, function(x) x[seq_len(max(which(x != 0)))])
	
	if (sp_pb[.(sp), N] == 1) # It means there there is only one individual (pb Larix kaempferi)
		diam_ls = list(pred_diam_mat)

	# Add to bole volume
	bole_volume_dt[.(sp_pb[sp, sp_code]), pred_diam_m := diam_ls]
}

## Check the validity of the computation
bole_volume_dt[, any(is.na(bole_volume_m3)) | any(is.infinite(bole_volume_m3))] # FALSE
round(bole_volume_dt[, range(bole_volume_m3)], digit = 3) # 0.004 -- 15.543

## Compute hdec
hdec = bole_volume_dt[, {
	pd <- unlist(pred_diam_m)
	pct_change <- diff(pd) / head(pd, -1)
	drop_idx <- which(pct_change <= -0.10)[1L]
	eh <- unlist(eval_h)
		.(first_drop_h = eh[drop_idx] + 0.5, # + 0.5 because one metre precision, so chosing mean dist
		first_pct_drop = round(pct_change[drop_idx] * 100, 2))
	}, by = .(sp_code, tree_id)]

## Add the computed volumes to Swiss dataset
swiss_dt = merge.data.table(swiss_dt, bole_volume_dt, by = c("sp_code", "tree_id"))

## Add the estimated hdec and remove pred_diam_m column
swiss_dt = merge.data.table(swiss_dt, hdec[, .(sp_code, tree_id, hdec = first_drop_h)],
	by = c("sp_code", "tree_id"))

swiss_dt[, pred_diam_m := NULL]
swiss_dt[, eval_h := NULL]

## Compute bole volumes using cylinders
# Volume trunk using cylindre formula with median diameter: log 0 to 2 --> DM1, log 2 to 4 --> DM3, etc...
swiss_dt[, vol_log_cyl := pi*diameter^2/4*2] # The *2 comes from the fact that each log is 2m long
swiss_dt[, bole_volume_cyl_m3 := sum(vol_log_cyl) + pi*D_coarsestemfinal^2/4*L_coarsestemfinal, by = tree_id]

# Total volume
swiss_dt[, total_volume_m3_cyl := bole_volume_cyl_m3 + V_coarsebranch + V_finewoodytotal]
swiss_dt[, total_volume_m3 := bole_volume_m3 + V_coarsebranch + V_finewoodytotal]
```

As can be seen, the stack cylindres overestimate the bole volume compared to taper functions. Whatever is the closest to the truth, there is no biais from `TapeS` (intercept is near 0).

```{r}
plot(unique(swiss_dt[, .(tree_id, total_volume_m3_cyl)])[, total_volume_m3_cyl],
	unique(swiss_dt[, .(tree_id, total_volume_m3)])[, total_volume_m3],
	pch = 19, cex = 0.65, axes = FALSE, xlab = "Cyl", ylab = "TapeS")
axis(1)
axis(2, las = 1)
abline(a = 0, b = 1, lwd = 4, col = "#CF212A")

lm(formula = total_volume_m3_cyl ~ total_volume_m3, data = swiss_dt)
```

## Save the prepared data

```{r}
saveRDS(swiss_dt, paste0(mnt_point, "data/switzerland.rds"))
if (!dir.exists(paste0(mnt_point, "data/coords_switzerland")))
	writeVector(x = coords_swiss, filename = paste0(mnt_point, "data/coords_switzerland"),
		filetype = "ESRI Shapefile", overwrite = TRUE)
```
