---
title: "Merge data FR -- CH and export for TikZ"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

## Intro

I merge the data for fitting and plotting. Most of the plots to present the data will be done using TikZ. This implies to format and export the data in csv for LaTeX. So far, I pla the following figures:

- A map (this one is actually not done in TikZ),
- Horizontal histogram of individual per species, for total volume (cumulative for TikZ by origin),
- Distributions of circumference, height, and slenderness for French NFI, EFM, and Emerge.

## Load and merge data

Data must be prepared by `01_[...]_prepare_data.qmd`. Note that the Emerge dataset uses the NFI species codes.

```{r}
#| output: false
#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(stringi)
library(terra)
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

opt = "_inra"

## Loading Emerge data
inra = readRDS(paste0(path_data, "nfi-inra_dt", opt, ".rds"))
inra = inra[, .(speciesName_sci, tree_id = unique_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = taper_height_flo, bole_volume_m3, total_volume_m3)]
setkey(inra, speciesName_sci)

## Loading Swiss data
swiss = readRDS(paste0(mnt_point, "data/switzerland.rds"))
swiss = unique(swiss[, .(speciesName_sci, tree_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = hdec, bole_volume_m3, total_volume_m3)])
setkey(swiss, speciesName_sci)

## Modern Emerge data
emerge = readRDS(paste0(mnt_point, "data/emerge_2009-2010.rds"))
emerge[, year := as.integer(stringi::stri_replace_all(str = dataset, replacement = "", regex = "emerge_"))]
emerge = emerge[, .(speciesName_sci, tree_id, fct_type, year, circumference_m, height, taper_height,
	bole_volume_m3 = bole_volume_conic_m3, total_volume_m3)]

## NFI data used to fit bole volumes (prepared by Florence)
nfi_data = readRDS("~/work/FairCarbon/gohon_florence/arbComp.rds")

## Species list
# Load data
species_dt = readRDS(paste0(mnt_point, "data/all_species.rds"))[, .(speciesName_sci, nfi = essence_ifn,
	sp_code_swiss, fct_type)]

# Change class to avoid merge warnings
species_dt[, nfi := as.integer(nfi)]

# Reshaping to long format
species_dt = melt.data.table(species_dt, id.vars = c("speciesName_sci", "fct_type"), variable.name = "origin",
	value.name = "sp_code", na.rm = TRUE, variable.factor = FALSE)

# Rename modalities
species_dt[stri_detect(str = origin, regex = "sp_code_"),
	origin := stri_sub(str = origin, from = stri_locate(str = origin, regex = "sp_code_")[, "end"] + 1)]

# Two species are missing in species_dt
species_dt = rbindlist(list(species_dt,
	data.table(speciesName_sci = c("Acer pseudoplatanus", "Acer campestre", "fruits",
		"Exotic broadleaves sp.", "Broadleaves sp."), fct_type = rep("broadleaf", 5),
	origin = rep("nfi", 5), sp_code = c(15, 21, 23, 29, 49))))

# Set keys
setkey(species_dt, speciesName_sci, origin)
```

I merge the three datasets into one. Note that species with less than 150 individuals will not be fit directly later on. I will, for some, group them by genera, such as *Quercus sp.*, *Fraxinus sp.*, etc.
```{r}
## Bind everything
tree_dt = rbindlist(list(inra = inra, swiss = swiss, emerge = emerge), idcol = "origin", fill = TRUE)

# Remove NAs
tree_dt = tree_dt[!is.na(bole_volume_m3)]

# Remove species with less than 150 individuals
tree_dt[, n_indiv := .N, by = speciesName_sci]

# Add column direct_fit to decide if a species is fitted directly or by group
tree_dt[, direct_fit := ifelse(n_indiv > 150, TRUE, FALSE)]
```

I subset the columns of the nfi dataset `nfi_data` and rename the columns so that it matches the total volume dataset `tree_dt`.

```{r}
#### Reshape NFI data to match tree_dt format
## Add species names
# Cast to integer for merging
nfi_data[, ess := as.integer(ess)]

# Mergin
nfi_data = merge.data.table(nfi_data,
	species_dt[.(unique(speciesName_sci), "nfi"), .(sp_code, speciesName_sci, fct_type)],
	by.x = "ess", by.y = "sp_code")

nfi_data = nfi_data[, .(origin = "nfi", speciesName_sci, tree_id = a, plot_id = npp,
	fct_type, circumference_m = c13, height = htot)]
```

Now, I merge all the datasets to plot the circumference, height, and slenderness distributions, and assign a colour to each based on `Met.Brewer("Lakota")`.

```{r}
chs_data = rbindlist(list(nfi_data,
	tree_dt[, .(origin, speciesName_sci, tree_id, plot_id, fct_type, circumference_m, height)]))

ls_colours = met.brewer("Kandinsky")[c(1, 3, 4)]
names(ls_colours) = c("swiss", "nfi", "inra")
# ls_colours = met.brewer("Egypt", n = 3)
# names(ls_colours) = c("inra", "nfi", "swiss")

chs_data[, colour := ls_colours["nfi"]]
chs_data[origin == "swiss", colour := ls_colours["swiss"]]
chs_data[origin == "inra", colour := ls_colours["inra"]]
```

## Complete NFI's data with coordinates

Florence did not add the coordinates to her query. I am using the internal NFI R package `inventR` to get them.

```{r}
login = inventR::connect_db()

coords_am = inventR::exec_req(conn = login, req = "
	SELECT
		coords.NPP AS plot_id, coords.XL, coords.YL

	FROM
		inv_exp_am.e1point AS coords

	ORDER BY
		plot_id;
	", DT = TRUE)

coords_nm = inventR::exec_req(conn = login, req = "
	SELECT
		coords.NPP AS plot_id, coords.XL, coords.YL

	FROM
		inv_exp_nm.e1point AS coords

	ORDER BY
		plot_id;
	", DT = TRUE)

logout = inventR::disconnect_db(login)

ll = unique(nfi_data[, .(plot_id, tree_id)])
setkey(ll)

coords_am[, plot_id := stri_sub(str = plot_id, to = stri_locate(str = plot_id, regex = "---")[, "start"] - 1)]

coords_nfi = rbindlist(list(am = coords_am, nm = coords_nm), idcol = "origin")
coords_nfi = na.omit(coords_nfi)

coords_nfi = coords_nfi[plot_id %in% nfi_data[, unique(plot_id)]]
coords_nfi = vect(x = coords_nfi, geom = c("xl", "yl"), crs = "EPSG:27572") # Lambert zone II, I checked
```

## Presentation of the data

I add few columns.

```{r}
chs_data[, genus := stri_sub(str = speciesName_sci,
	to = stri_locate_first(str = speciesName_sci, regex = " ")[, "start"] - 1)]
```

### Maps of plots location

The plots are located:

```{r}
#| echo: false
#| label: fig-map
#| fig-cap: "Location of the plots, with NFI for the Naional Forest Invetory of France, and EFM the Experimental Forest Monitoring of Switzerland [@Didion2024]"
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 9

#### Coordinates
## Read point shapefiles
coords_inra = vect(paste0(mnt_point, "data/coords_inra_full"))
coords_swiss = vect(paste0(mnt_point, "data/coords_switzerland"))

## Read shapefiles countries
switzerland = project(vect(paste0(mnt_point, "data/shapefiles/switzerland/CHE_adm0.shp")), coords_inra)
france = project(vect(paste0(mnt_point, "data/shapefiles/france/FRA_adm0.shp")), coords_inra)
belgium = project(vect(paste0(mnt_point, "data/shapefiles/belgium/gadm41_BEL_0.shp")), coords_inra)
italy = project(vect(paste0(mnt_point, "data/shapefiles/italy/gadm41_ITA_0.shp")), coords_inra)
luxembourg = project(vect(paste0(mnt_point, "data/shapefiles/luxembourg/gadm41_LUX_0.shp")), coords_inra)
germany = project(vect(paste0(mnt_point, "data/shapefiles/germany/gadm41_DEU_0.shp")), coords_inra)
spain = project(vect(paste0(mnt_point, "data/shapefiles/spain/gadm41_ESP_0.shp")), coords_inra)

fr_ch = project(rbind(france, switzerland), coords_inra)

## Plot
pdf("../../../writing/article_allometry/pictures/map2.pdf")
plot(fr_ch, lwd = 2, axes = FALSE)
points(coords_nfi, col = ls_colours["nfi"], pch = 20, cex = 0.35)
points(coords_inra, col = ls_colours["inra"], pch = 20, cex = 1.15)
points(project(coords_swiss, coords_inra), col = ls_colours["swiss"], pch = 20, cex = 0.55)
plot(fr_ch, lwd = 2, add = TRUE, bg = "transparent")
plot(luxembourg, lwd = 2, add = TRUE)
plot(germany, lwd = 2, add = TRUE)
plot(belgium, lwd = 2, add = TRUE)
plot(italy, lwd = 2, add = TRUE)
plot(spain, lwd = 2, add = TRUE)
# legend(x = "bottomleft", legend = c("Emerge", "NFI", "EFM"), fill = ls_colours[c("inra", "nfi", "swiss")], bty = "n", inset = c(0, 0.1))
dev.off()
```

### Number of individuals per species and dataset

The data are composed by the following species:
```{r}
#| echo: false
#| label: fig-data
#| fig-cap: Number of individuals per genus (French data), with conifers in blue and broadleaves in orange. Watch out, it is a $\\log_{10}$ scale
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 9

## Create nb indiv data
species_nb = unique(chs_data[speciesName_sci != "fruits"][origin %in% c("swiss", "emerge", "inra"),
	.(N = .N, colour), by = .(genus, origin)])

species_nb[, tot := sum(N), by = genus]
setorder(species_nb, tot)

par(mar = c(5, 12, 2, 2))

## Use barplot fct (for preview)
ll = dcast.data.table(species_nb, colour ~ genus, value.var = "N", fill = 0)
# cols = colnames(ll[, .SD, .SDcols = !patterns("^colour")])
# ll[, (cols) := lapply(X = .SD, FUN = function(x) fcase(x == 0, 0, x > 0, log1p(x))), .SDcols = cols]
ll = as.matrix(ll, rownames = "colour")

ord = species_nb[, sum(N), by = genus][order(V1)]

ll = ll[, ord[, genus]]

# Get the stacked barplot
barplot(ll,
	col = rownames(ll),
	border = "white",
	space = 0.04,
	font.axis = 2,
	log = "x",
	offset = 1,
	las = 1,
	horiz = TRUE, # Horizontal bars
	xlab = "Number of individuals")

fwrite(species_nb[, .(genus, origin, N)], "../../../writing/article_allometry/data/n_indiv.csv")
```

### Distribution of the data in terms of circumference, height, slenderness

Densities are computed here, with 1024 points, and then exported to csv files for TikZ plotting.

```{r}
#| echo: false
#| label: fig-distrib
#| fig-cap: Distribution of the data
#| fig-subcap: 
#|   - "Densities of circumferences"
#|   - "Densities of heights"
#|   - "Densities of slenderness (height/circumference)"
#| layout-nrow: 3
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 9

# --- Circumferences
dd_fr = density(chs_data[origin == "nfi", circumference_m]) # French NFI
dd_ch = density(chs_data[origin == "swiss", circumference_m]) # EFM data
dd_in = density(chs_data[origin %in% c("emerge", "inra"), circumference_m], na.rm = TRUE) # Inra/Emerge data

plot(dd_fr, col = ls_colours["nfi"], axes = FALSE, lwd = 3, xlab = "Circumference (m)", main = "",
	ylim = c(0, max(dd_fr$y, dd_in$y, dd_ch$y)))
polygon(dd_fr, col = paste0(ls_colours["nfi"], "AA"))

lines(dd_in, col = ls_colours["inra"], lwd = 3)
polygon(dd_in, col = paste0(ls_colours["inra"], "AA"))

lines(dd_ch, col = ls_colours["swiss"], lwd = 3)
polygon(dd_ch, col = paste0(ls_colours["swiss"], "AA"))

axis(1)
axis(2, las = 1)
legend(x = "topright", legend = c("Emerge", "NFI", "EFM"), fill = ls_colours[c("inra", "nfi", "swiss")], bty = "n")

# --- Heights
dd_fr = density(chs_data[origin == "nfi", height]) # French NFI
dd_ch = density(chs_data[origin == "swiss", height]) # EFM data
dd_in = density(chs_data[origin %in% c("emerge", "inra"), height], na.rm = TRUE) # Inra/Emerge data

plot(dd_fr, col = ls_colours["nfi"], axes = FALSE, lwd = 3, xlab = "Height (m)", main = "",
	ylim = c(0, max(dd_fr$y, dd_in$y, dd_ch$y)))
polygon(dd_fr, col = paste0(ls_colours["nfi"], "AA"))

lines(dd_in, col = ls_colours["inra"], lwd = 3)
polygon(dd_in, col = paste0(ls_colours["inra"], "AA"))

lines(dd_ch, col = ls_colours["swiss"], lwd = 3)
polygon(dd_ch, col = paste0(ls_colours["swiss"], "AA"))

axis(1)
axis(2, las = 1)
legend(x = "topright", legend = c("Emerge", "NFI", "EFM"), fill = ls_colours[c("inra", "nfi", "swiss")], bty = "n")

# --- Slenderness
dd_fr = density(chs_data[origin == "nfi", height/circumference_m]) # French NFI
dd_ch = density(chs_data[origin == "swiss", height/circumference_m]) # EFM data
dd_in = density(chs_data[origin %in% c("emerge", "inra"), height/circumference_m], na.rm = TRUE) # Inra/Emerge data

fwrite(data.table(xfr = dd_fr$x, yfr = dd_fr$y, xin = dd_in$x, yin = dd_in$y,
	xch = dd_ch$x, ych = dd_ch$y), "../../../writing/article_allometry/data/densities_slenderness.csv")

plot(dd_fr, col = ls_colours["nfi"], axes = FALSE, lwd = 3, xlab = "Slenderness (height/circumference)", main = "",
	ylim = c(0, max(dd_fr$y, dd_in$y, dd_ch$y)))
polygon(dd_fr, col = paste0(ls_colours["nfi"], "AA"))

lines(dd_in, col = ls_colours["inra"], lwd = 3)
polygon(dd_in, col = paste0(ls_colours["inra"], "AA"))

lines(dd_ch, col = ls_colours["swiss"], lwd = 3)
polygon(dd_ch, col = paste0(ls_colours["swiss"], "AA"))

axis(1)
axis(2, las = 1)
legend(x = "topright", legend = c("Emerge", "NFI", "EFM"), fill = ls_colours[c("inra", "nfi", "swiss")], bty = "n")
```
