---
title: "Analyse data to estimate volumes"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
execute:
  error: true
bibliography: references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    keep-tex: true
    pdf-engine: lualatex
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

\newcommand{\ie}{*i.e.,*}
\newcommand{\F}{\mathscr{F}}
\newcommand{\N}{\mathbfscr{N}}

\newcommand{\vtot}{v_{\text{tot}}}
\newcommand{\vbole}{v_{\text{bole}}}
\newcommand{\vbr}{v_{\text{branches}}}

```{r}
#| output: false
#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
```

## Introduction
The data have been prepared in `01_prepare_data.qmd`. In this document, I fit a model on both INRA and French NFI data. In this section, I explore the data, and in the next section I define the model. Lastly  I explore few results.

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "/data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading
stanData = readRDS(paste0(path_data, "stanData_nfi-inra.rds"))
forest_dt = readRDS(paste0(path_data, "nfi-inra_dt.rds"))
ind_species = readRDS(paste0(path_data, "ind_species_nfi-inra.rds"))

colours = paste0(met.brewer("Egypt")[1:4])
names(colours) = c("conifer", "broadleaf", "inra", "nfi")

forest_dt[fct_type == "broadleaf", fct_colour := colours["broadleaf"]]
forest_dt[fct_type == "conifer", fct_colour := colours["conifer"]]
forest_dt[origin_data == "inra", origin_colour := colours["inra"]]
forest_dt[origin_data == "nfi", origin_colour := colours["nfi"]]
```

```{r}
colnames(forest_dt)

forest_dt[!is.na(crown_volume_m3), .N, by = fct_type]

png("./test.png", res = 120, height = 1080, width = 1080)
plot(forest_dt[, log(circumference_m)], forest_dt[, log(bole_volume_m3)], col = forest_dt[, fct_colour], pch = 19,
	xlab = "Circumference", ylab = "Bole volume")
legend(x = "topright", legend = c("Broadleaf", "Conifer"), fill = colours[c("broadleaf", "conifer")])
dev.off()

png("./test2.png", res = 120, height = 1080, width = 1080)
plot(forest_dt[!is.na(crown_volume_m3), log(circumference_m)], forest_dt[!is.na(crown_volume_m3), log(crown_volume_m3)],
	col = forest_dt[!is.na(crown_volume_m3), fct_colour], pch = 19, xlab = "Circumference", ylab = "Crown volume")
legend(x = "topright", legend = c("Broadleaf", "Conifer"), fill = colours[c("broadleaf", "conifer")])
dev.off()

lm_bole = lm(forest_dt[, log(bole_volume_m3)] ~ 1 + forest_dt[, log(circumference_m)] + forest_dt[, log(circumference_m)^4])
lm_crown = lm(forest_dt[, log(crown_volume_m3)] ~ 1 + forest_dt[, log(circumference_m)] + forest_dt[, log(circumference_m)^4])
```

## Methods
I first try a simple model:
$$
\vtot = \vbole + \vbr
$$
where the bole volume, $\vbole$, and the branch volume, $\vbr$, both follow a lognormal distribution. For a definition of this volumes see `01_prepare_data.qmd`. The sum of the lognormals can be approximated via a method described in @Lo2012.

```{stan output.var = "nfi_inra", cache = TRUE}
data {
	// Dimensions and indices
	int <lower = 1> N; // Total number of individuals
	int <lower = 1, upper = N> N_inra; // Number of individuals with a measured crown
	int <lower = 1, upper = N> S; // Number of species
	int <lower = 0, upper = S> n_sp_broad_inra; // Number of broadleaf species
	int <lower = 0, upper = S - n_sp_broad_inra> n_sp_conif_inra; // Number of conifer species

	int <lower = 1, upper = N - 1> ind_start_broad; // Broadleaf index start
	int <lower = 1, upper = N> ind_end_broad; // Broadleaf index end
	int <lower = 1, upper = N - 1> ind_start_conif; // Conifer index start
	int <lower = 1, upper = N> ind_end_conif; // Conifer index end
	
	array[n_sp_broad_inra] int ind_start_broad_inra; // Broadleaf crown species index start
	array[n_sp_broad_inra] int ind_end_broad_inra; // Broadleaf crown species index end
	array[n_sp_conif_inra] int ind_start_conif_inra; // Conifer crown species index start
	array[n_sp_conif_inra] int ind_end_conif_inra; // Conifer crown species index end

	// Predictors
	vector <lower = 0> [N] circumference_m;

	// Response variables
	vector <lower = 0> [N] bole_volume_m3;
	vector <lower = 0> [N] crown_volume_m3;
}

transformed data {
	// Predictors
	vector [N] log_circumference_m = log(circumference_m);

	// Response variables
	vector [N] log_bole_volume_m3 = log(bole_volume_m3);
	vector [N] log_crown_volume_m3 = log(crown_volume_m3);
}

parameters {
	// Intercepts
	vector [2] a_bole;
	vector [2] a_crown;

	// Slopes
	vector [2] b_bole;
	vector [2] b_crown;
	vector [2] c_bole;
	vector [2] c_crown;

	// Variances
	vector <lower = 0> [2] sigma_bole;
	vector <lower = 0> [2] sigma_crown;
}

model {
	// Priors
	// --- Intercepts
	target += normal_lpdf(a_bole | 0, 2);
	target += normal_lpdf(a_crown | 0, 2);

	// --- Slopes
	target += normal_lpdf(b_bole | 0, 2);
	target += normal_lpdf(b_crown | 0, 2);

	target += normal_lpdf(c_bole | 0, 2);
	target += normal_lpdf(c_crown | 0, 2);

	// --- Variances
	target += gamma_lpdf(sigma_bole | 1.0^2/5.0, 1.0/5.0);
	target += gamma_lpdf(sigma_crown | 1.0^2/5.0, 1.0/5.0);

	// Likelihood
	// --- Broadleaves, bole
	target += normal_lpdf(log_bole_volume_m3[ind_start_broad:ind_end_broad] | a_bole[1] + 
		b_bole[1]*log_circumference_m[ind_start_broad:ind_end_broad] +
		c_bole[1]*log_circumference_m[ind_start_broad:ind_end_broad].^4, sigma_bole[1]);
	
	// --- Conifers, bole
	target += normal_lpdf(log_bole_volume_m3[ind_start_conif:ind_end_conif] | a_bole[2] + 
		b_bole[2]*log_circumference_m[ind_start_conif:ind_end_conif] +
		c_bole[2]*log_circumference_m[ind_start_conif:ind_end_conif].^4, sigma_bole[2]);

	// --- Broadleaves, crown
	for (i in 1:n_sp_broad_inra)
		target += normal_lpdf(log_crown_volume_m3[ind_start_broad_inra[i]:ind_end_broad_inra[i]] | a_crown[1] + 
			b_crown[1]*log_circumference_m[ind_start_broad_inra[i]:ind_end_broad_inra[i]] +
			c_crown[1]*log_circumference_m[ind_start_broad_inra[i]:ind_end_broad_inra[i]].^4, sigma_crown[1]);
	
	// --- Conifers, crown
	for (i in 1:n_sp_conif_inra)
		target += normal_lpdf(log_crown_volume_m3[ind_start_conif_inra[i]:ind_end_conif_inra[i]] | a_crown[2] + 
			b_crown[2]*log_circumference_m[ind_start_conif_inra[i]:ind_end_conif_inra[i]] +
			c_crown[2]*log_circumference_m[ind_start_conif_inra[i]:ind_end_conif_inra[i]].^4, sigma_crown[2]);
}
```


```{r}
n_chains = 4
# woodstock_seed = 1969 - 08 - 18
iter_warmup = 350
iter_sampling = 500

nfi_inra = cmdstan_model("./test_nfi-inra.stan")

stanData$total_volume_m3 = NULL
stanData$crown_volume_m3[is.na(stanData$crown_volume_m3)] = Inf
stanData$ind_start_broad = min(stanData$ind_start_broad)
stanData$ind_end_broad = max(stanData$ind_end_broad)
stanData$ind_start_conif = min(stanData$ind_start_conif)
stanData$ind_end_conif = max(stanData$ind_end_conif)

## Fit
if (!file.exists("fit_inra-nfi.rds"))
{
	fit = nfi_inra$sample(data = stanData, chains = n_chains, parallel_chains = ifelse(n_chains < 4, n_chains, 4),
		seed = NULL, refresh = 50, max_treedepth = 12, save_warmup = TRUE,
		iter_sampling = iter_sampling, iter_warmup = iter_warmup, adapt_delta = 0.8)
	fit$save_output_files(dir = "./", basename = paste0("fit_inra-nfi"), random = FALSE)
	saveRDS(fit, "./fit_inra-nfi.rds")
} else {
	fit = readRDS("./fit_inra-nfi.rds")
}
```

## Results
As can be seen here in what follows there is no problem of convergence
```{r}
mcmc_trace(fit$draws(c("a_bole", "a_crown")))
mcmc_trace(fit$draws(c("b_bole", "b_crown")))
mcmc_trace(fit$draws(c("c_bole", "c_crown")))
mcmc_trace(fit$draws(c("sigma_bole", "sigma_crown")))

mcmc_rhat(bayesplot::rhat(fit))
mcmc_nuts_divergence(bayesplot::nuts_params(fit), lp = log_posterior(fit))
```

I now do some posterior checks:

1. What are the residuals?
2. How well does it predict an independent dataset?

