---
title: "Simple tests of bole volume models"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(inventR)
library(interp)
library(shape)
library(loo)

## Tool function
source("./toolFunctions.R")

plot_sp = function(fit, sp, forest = tree_dt[.(sp)], pred = TRUE, simplif = FALSE,
	n_bins = 4, pal = "Hiroshige", selected_variable = NULL, print_plot = TRUE)
{
	# Get parameters
	n_sampling = fit$metadata()$iter_sampling
	n_chains = fit$num_chains()

	if (n_chains != nrow(fit$metadata()$time))
	{
		warning("Some chains are missing probably because they could not start! Using the number of successful chains")
		n_chains = nrow(fit$metadata()$time)
	}

	# Add colour to the measured points according to selected variable
	if (!is.null(selected_variable))
	{
		if (!(selected_variable %in% colnames(forest)))
			stop(paste0("'", selected_variable, "' is not a column of forest"))
		
		colours = MetBrewer::met.brewer(pal, n_bins)[1:n_bins]
		forest[, colour_ind := as.numeric(cut(.SD[[selected_variable]], breaks = n_bins))] # Indices per category
		forest[, colour := colours[colour_ind]] # Map selected variable to colour using indices
	} else {
		forest[, colour := "#3355AA33"] # Default colour
	}

	# Define fitted functions (depends if simplified model 2 stages vs 3 stages)
	if (!simplif)
	{
		paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")
	} else {
		paramsVec_simplif = getParams(model_cmdstan = fit,
			params_names = c("alpha", "beta_", "gamma", "delta"), type = "mean")
		paramsVec = c(
			c = unname(paramsVec_simplif["alpha"]),
			j = 1,
			k = unname(paramsVec_simplif["beta_"]),
			m = unname(exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"]),
			n = unname(paramsVec_simplif["delta"] + paramsVec_simplif["alpha"]),
			s = unname(paramsVec_simplif["beta_"])
		)
	}

	r_func = function(x, pars)
	{
		return ((pars["m"] - pars["c"]) * exp(pars["j"] - pars["k"]*x) * (pars["k"]*x/pars["j"])^pars["j"] +
			pars["c"] - (pars["c"] - pars["n"])*exp(-pars["s"]*x));
	}
	
	psi = function(x, pars)
	{
		r = r_func(x, pars)

		return ( 1/pars["c"] * x^( 1 - (log(r) - log(pars["c"])) / log(x) ) )
	}

	# Check rhats
	pars_names = names(paramsVec)
	if (simplif)
		pars_names = c("alpha", "beta_", "gamma", "delta")
	rhats = rhat(object = fit, pars = pars_names)

	# Divergence check up
	params = getParams(model_cmdstan = fit, params_names = pars_names, type = "all")
	div = posterior::subset_draws(fit$sampler_diagnostics(), variable = "divergent__")

	if (!print_plot)
		return (list(any_div = any(div != 0), loc = div, rhats = rhats))

	# Plot figure r vs bole and fitted curve
	plot(forest[, bole_volume_m3], forest[, r], pch = 19, col = forest[, colour], axes = FALSE,
		xlab = "Bole volume", ylab = "Ratio bole/tot volumes", lwd = 0, cex = 0.75,
		main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
	if (!is.null(selected_variable))
		legend("topright", fill = colours, title = paste(selected_variable, "classes"),
			legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
	curve(r_func(x, paramsVec), add = TRUE, lwd = 3)
	axis(1)
	axis(2, las = 1)

	# Plot predictions
	if (pred)
	{
		sim = apply(X = fit$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
		plot(sim, forest[, total_volume_m3], pch = 19, col = forest[, colour], axes = FALSE,
			xlab = "Predicted total volume", ylab = "Observed total volume",
			main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
		if (!is.null(selected_variable))
			legend("topright", fill = colours, title = paste(selected_variable, "classes"),
				legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
		abline(a = 0, b = 1, lty = "dashed", lwd = 1.5, col = "#2E2E2E")
		axis(1)
		axis(2, las = 1)
		
		plot(forest[, bole_volume_m3], forest[, total_volume_m3], pch = 19, col = forest[, colour],
			axes = FALSE, xlab = "Observed bole volume", ylab = "Observed total volume",
			main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
		if (!is.null(selected_variable))
			legend("topright", fill = colours, title = paste(selected_variable, "classes"),
				legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
		abline(a = 0, b = 1, lty = "dashed", lwd = 0.85, col = "#2E2E2E")
		curve(psi(x, paramsVec), add = TRUE, lwd = 3)
		axis(1)
		axis(2, las = 1)
	}

	# Plot chains if any rhat problem
	if (any(rhats > 1.05))
	{
		if (!simplif)
		{
			lazyTrace(fit$draws("c"), main = paste0(sp, ", c"))
			lazyTrace(fit$draws("j"), main = paste0(sp, ", j"))
			lazyTrace(fit$draws("k"), main = paste0(sp, ", k"))
			lazyTrace(fit$draws("m"), main = paste0(sp, ", m"))
			lazyTrace(fit$draws("n"), main = paste0(sp, ", n"))
			lazyTrace(fit$draws("s"), main = paste0(sp, ", s"))

			paramsVec1 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 1)
			paramsVec2 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 2)
			paramsVec3 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 3)
			paramsVec4 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 4)
		} else {
			lazyTrace(fit$draws("alpha"), main = "alpha")
			lazyTrace(fit$draws("beta_"), main = "beta_")
			lazyTrace(fit$draws("gamma"), main = "gamma")
			lazyTrace(fit$draws("delta"), main = "delta")

			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 1)
			paramsVec1 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 2)
			paramsVec2 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 3)
			paramsVec3 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 4)
			paramsVec4 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
		}
		plot(forest[, bole_volume_m3], forest[, r], pch = 19, axes = FALSE, lwd = 0,
			xlab = "Bole volume", ylab = "Ratio bole/tot volumes", main = sp)
		curve(r_func(x, paramsVec), add = TRUE, lwd = 3)
		curve(r_func(x, paramsVec1), lwd = 3, add = TRUE, col = met.brewer("Egypt")[1])
		curve(r_func(x, paramsVec2), lwd = 3, add = TRUE, col = met.brewer("Egypt")[2])
		curve(r_func(x, paramsVec3), lwd = 3, add = TRUE, col = met.brewer("Egypt")[3])
		curve(r_func(x, paramsVec4), lwd = 3, add = TRUE, col = met.brewer("Egypt")[4])
		axis(1)
		axis(2, las = 1)
	}

	return (list(any_div = any(div != 0), loc = div, rhats = rhats))
}

plot_divergences = function(fit, div, simplif = FALSE)
{
	params = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "all")

	div_c = posterior::subset_draws(params[, , "c"], draw = which(div == 1))
	div_j = posterior::subset_draws(params[, , "j"], draw = which(div == 1))
	div_m = posterior::subset_draws(params[, , "m"], draw = which(div == 1))
	div_k = posterior::subset_draws(params[, , "k"], draw = which(div == 1))
	div_s = posterior::subset_draws(params[, , "s"], draw = which(div == 1))

	# Plot divergence for c and j
	plot(params[, , "c"], params[, , "j"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "j", main = "c and j")
	points(div_c, div_j, pch = 18, col = "#0F7BA2")

	# Plot divergence for c and k
	plot(params[, , "c"], params[, , "k"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "k", main = "c and k")
	points(div_c, div_k, pch = 18, col = "#0F7BA2")

	# Plot divergence for c and m
	plot(params[, , "c"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "m", main = "c and m")
	points(div_c, div_m, pch = 18, col = "#0F7BA2")

	# Plot divergence for j and k
	plot(params[, , "j"], params[, , "k"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "j", ylab = "k", main = "j and k")
	points(div_j, div_k, pch = 18, col = "#0F7BA2")

	# Plot divergence for j and m
	plot(params[, , "j"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "j", ylab = "m", main = "j and m")
	points(div_j, div_m, pch = 18, col = "#0F7BA2")

	# Plot divergence for k and m
	plot(params[, , "k"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "k", ylab = "m", main = "k and m")
	points(div_k, div_m, pch = 18, col = "#0F7BA2")
	
	# Bonus plots of susceptible correlated parameters: k and s
	plot(params[, , "k"], params[, , "s"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "k", ylab = "s", main = "k and s")
	points(div_k, div_s, pch = 18, col = "#0F7BA2")
}
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

opt = "_inra"

# Loading Emerge data
inra = readRDS(paste0(path_data, "nfi-inra_dt", opt, ".rds"))
inra = inra[, .(speciesName_sci, tree_id = unique_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = taper_height_flo, bole_volume_m3, total_volume_m3)]
setkey(inra, speciesName_sci)

# Loading Swiss data
swiss = readRDS(paste0(mnt_point, "data/switzerland.rds"))
swiss = swiss[, .(speciesName_sci, tree_id, plot_id, fct_type, year, circumference_m, height, taper_height = hdecro,
	bole_volume_m3, total_volume_m3)]
setkey(swiss, speciesName_sci)

## Modern Emerge data
emerge = readRDS(paste0(mnt_point, "data/emerge_2009-2010.rds"))
emerge[, year := as.integer(stringi::stri_replace_all(str = dataset, replacement = "", regex = "emerge_"))]
emerge = emerge[, .(speciesName_sci, tree_id, fct_type, year, circumference_m, height, taper_height,
	bole_volume_m3 = bole_volume_conic_m3, total_volume_m3)]

## Bind everything
tree_dt = rbindlist(list(inra = inra, swiss = swiss, emerge = emerge), idcol = "origin", fill = TRUE)

# Add colours France vs Swiss
ls_colours = met.brewer("Egypt")[1:4]
names(ls_colours) = c("fr_comp", "fr", "ch_comp", "ch")
tree_dt[, colour := ls_colours["fr"]] # Inra and Emerge, in blue
tree_dt[origin == "swiss", colour := ls_colours["ch"]] # Swiss, in yellow

# Remove NAs
tree_dt = tree_dt[!is.na(bole_volume_m3)]

# Remove species with less than 150 individuals
tree_dt[, n_indiv := .N, by = speciesName_sci]
tree_dt = tree_dt[n_indiv > 150]

tree_dt[, .N, by = .(speciesName_sci, colour)][order(speciesName_sci)]

# Set key to sort by species and colour (i.e., France vs Switzerland)
setkey(tree_dt, speciesName_sci, colour)
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script try to parametrise the total volume in function of the bole volume, that is to say, to use the basic definition of allometry: the dimension of an organ $y$ is proportional to an organ $x$, such as in this power-law $y = a x^b$. Here, $y$ is the total volume and $x$ is the bole volume. In @Yang2024, they do it with above-ground biomass as $y$ and stem biomass as $x$. I expect to be able to apply the same type of functions to volume, as volume and biomass are related by wood density. It should be the same if wood density is constant between the crown and the bole, and only the scale should change! According to @Knapic2011 there is little variation within tree oaks, while there can be for pines @Repola2006.

Here is an intuition on how the function @eq-yang2024 from @Yang2024, works.
$$
	r = (m - c) \exp[j - kx] \left( \frac{kx}{j} \right)^j + c - (c - n)\exp[-sx]
$${#eq-yang2024}
The parameter $j$ influences the curvature of the function as claimed in the appendix S1 of that paper, but not intuitive: The lower $j$, the stepper the function! However, it is correct to assume that $j < 1$ as otherwise the function is first convex and then concave. So the stan code should reflect this constraint, although the model is perfectly defined for $j > 1$.

```{r}
#| label: fig-var_j
#| fig-cap: Different shape of the function $r = \exp[j - kx] \left( \frac{kx}{j} \right)^j$ for k = 0.06 and j between 0 and 1.6. Dashed lines are for j > 1
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

f = function(x, j, k)
	return (exp(j - k*x) * (k*x/j)^j)

k = 0.06
Homer = met.brewer("Homer1")
curve(f(x, 0.2, k), from = 0, to = 25, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (j in 2:8)
	curve(f(x, j/5, k), add = TRUE, col = Homer[j], lwd = 3, lty = ifelse(j/5 < 1, 1, 2))
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = (1:8)/5, bty = "n")
```

```{r}
#| label: fig-var_k
#| fig-cap: Different shape of the function $r = \exp[j - kx] \left( \frac{kx}{j} \right)^j$ for j = 0.5 and k between 0.02 and 0.16.
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

j = 0.6
curve(f(x, j, 0.02), from = 0, to = 25, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (k in 2:8)
	curve(f(x, j, k/50), add = TRUE, col = Homer[k], lwd = 3)
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = (1:8)/50, bty = "n")
```

```{r}
#| label: fig-var_s
#| fig-cap: Different shape of the function @eq-yang2024 for j = 0.5, k = , m = , n = , c = and s between 0.02 and 0.16.
#| fig-subcap:
#|   - "For Mammea"
#|   - "For Triplochiton"   
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 4
#| layout-ncol: 2

f = function(x, c, j, k, m, n, s)
	return ((m - c)*exp(j - k*x) * (k*x/j)^j + c - (c - n)*exp(-s*x))

# Values for Mammea taken from Yang et al. 2024
m = 0.65
n = 0.4
c = 0.6
P = 150
j = 1 - 0.5*exp(-0.001*P)
k = 0.004557
print(paste("s =", round(10/P, 3)))

curve(f(x, c, j, k, m, n, 10/P), from = 0, to = 500, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (s in 2:8)
	curve(f(x, c, j, k, m, n, s*10/P), add = TRUE, col = Homer[s], lwd = 3)
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = round((1:8)*10/P, 3), bty = "n")


curve(f(x, c, j, k, m, n, 10/P), from = 500, to = 2000, lwd = 3, axes = FALSE, xlab = "Bole volume",
	ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (s in 2:8)
	curve(f(x, c, j, k, m, n, s*10/P), add = TRUE, col = Homer[s], lwd = 3)
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = round((1:8)*10/P, 3), bty = "n")
```

*Larix decidua* seems to be a particular case of the the model @eq-yang2024. It seems to be a simpler convex function with only two stages in its ratio rather than three: First an sharp increase, as for any other species, which corresponds to the juvenile stage, and then a stable phase, which is the asymptotic ratio of adults. There seems to be no intermediate stage. By adding some constraints, and choosing $s = k$ since the intermediate stage merged with the stable stage, and $j = 1$, we get the following type of function:

$$
	r = \alpha + \exp[-\beta x](\gamma x + \delta)
$${#eq-yang2024_noIntermediate}

```{r}
#| label: fig-subcase
#| fig-cap: Particular case of the generic function @eq-yang2024, that reduces to @eq-yang2024_noIntermediate when $j = 1$, $s = k$, $c = \alpha$, $k = \beta$, $m = e^{-1}\frac{\gamma}{\beta} + \alpha$, and $n = \delta + \alpha$.
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

alpha = 0.85
beta = 0.02
gamma = -0.01
delta = -0.7

j = 1
s = beta
c = alpha
k = beta
m = exp(-1)*gamma/k + alpha
n = delta + alpha

curve(f(x, c, j, k, m, n, s), from = 0, to = 500, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
```

Lastly, I use some beta priors in my model. As I am not that familiar with this distribution, it worth it to have some basic plots (Fig. @fig-beta_distrib).

```{r}
#| label: fig-beta_distrib
#| fig-cap: Different shapes of the beta distribution
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

curve(dbeta(x, shape1 = 5, shape2 = 2), from = 0, to = 1, lwd = 3, axes = FALSE, xlab = "x",
	ylab = "Beta pdf", col = Homer[1], ylim = c(0, 3.6))
axis(1)
axis(2, las = 1)

curve(dbeta(x, shape1 = 8, shape2 = 2), add = TRUE, col = Homer[2], lwd = 3)

curve(dbeta(x, shape1 = 8, shape2 = 1), add = TRUE, col = Homer[3], lwd = 3)

curve(dbeta(x, shape1 = 5, shape2 = 5), add = TRUE, col = Homer[4], lwd = 3)

curve(dbeta(x, shape1 = 1.5, shape2 = 1.5), add = TRUE, col = Homer[5], lwd = 3)

curve(dbeta(x, shape1 = 2, shape2 = 5), add = TRUE, col = Homer[8], lwd = 3)

curve(dbeta(x, shape1 = 2, shape2 = 8), add = TRUE, col = Homer[7], lwd = 3)

curve(dbeta(x, shape1 = 1, shape2 = 8), add = TRUE, col = Homer[6], lwd = 3)

legend(x = "topright", fill = met.brewer("Homer1")[c(1:5, 8, 7, 6)],
	legend = c("5, 2", "8, 2", "8, 1", "5, 5", "1.5, 1.5", "2, 5", "2, 8", "1, 8"), bty = "n")
```

## Check the overlap of the ratio data
The Swiss data have an abnormal number of ratio values near one... It happens on very concentrated years and plots, which makes me think 
```{r}
library(terra)
swiss_coords = vect(paste0(mnt_point, "data/coords_switzerland"))

tree_dt[, r := bole_volume_m3/total_volume_m3]

aa = tree_dt[origin != "swiss"][r > 0.98][, .N, by = .(plot_id, year)]
bb = tree_dt[origin != "swiss"][, .(n_tot = .N), by = .(plot_id, year)]

aa = merge.data.table(aa, bb, by = c("plot_id", "year"))
aa[, prop := round(100*N/n_tot, 1)]
aa[, mean(prop)] # In average for plots having trees with r > 0.98, it is 10% of the indiv that are concerned

ls_pb = tree_dt[origin == "swiss"][r > 0.98][, .N, by = .(plot_id, year)]
n_indiv = tree_dt[origin == "swiss"][, .(n_tot = .N), by = .(plot_id, year)]

ls_pb = merge.data.table(ls_pb, n_indiv, by = c("plot_id", "year"))
ls_pb[, prop := round(100*N/n_tot, 1)]

ls_pb[prop > 20]

ls_pb[N > 20, sum(N)]*100/ls_pb[, sum(N)]
ls_pb[N > 50][order(year)]

pb_66 = ls_pb[N > 50, unique(plot_id)] # 2/3rd of the problematic trees come from theses plots

setkey(ls_pb, plot_id, year)

ls_pb_10p = ls_pb[prop > 10, .(plot_id, year)] # "01015004"
ls_pb[ls_pb_10p]
tree_dt[plot_id %in% ls_pb_10p$plot_id, .N, by = speciesName_sci]

switzerland = project(vect(paste0(mnt_point, "data/shapefiles/switzerland/CHE_adm0.shp")), swiss_coords)

## Plots
swiss_cols = met.brewer("Egypt")[1:4]
names(swiss_cols) = unique(swiss_coords$region)
swiss_coords$colour = swiss_cols[swiss_coords$region]

plot(switzerland, lwd = 2, axes = FALSE)
points(swiss_coords, pch = 20, cex = 0.55)
points(swiss_coords[swiss_coords$plot_id %in% ls_pb_10p$plot_id],
	col = swiss_coords[swiss_coords$plot_id %in% ls_pb_10p$plot_id]$colour,
	pch = 4, cex = 3)

plot(switzerland, lwd = 2, axes = FALSE)
points(swiss_coords, pch = 20, cex = 0.55)
points(swiss_coords[swiss_coords$plot_id %in% pb_66],
	col = swiss_coords[swiss_coords$plot_id %in% pb_66]$colour,
	pch = 4, cex = 4)

dt = as.data.table(swiss_coords[swiss_coords$plot_id %in% pb_66])

swiss_coords[(swiss_coords$plot_id %in% ls_pb_10p$plot_id) & (swiss_coords$year %in% ls_pb_10p$year)]

plot(tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), bole_volume_m3],
	tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), r],
	pch = 19, xlab = "Bole volume", ylab = "Ratio")

plot(tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), circumference_m],
	tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), r],
	pch = 19, xlab = "Circumference", ylab = "Ratio")

100*tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year))][r > 0.98, .N]/
	tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year))][, .N]
```

## Removing some trees from the Swiss data

As demonstrated in the previous section, the overlap is not that good for some very particular plots where many trees have a ratio greater than 0.98! I decided to remove trees with a ratio $r > 0.98$ and belonging to plots with more than 10% of such trees.

```{r}
tree_dt_copy = copy(tree_dt)
tree_dt = tree_dt[!((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98))]
# tree_dt[plot_id == "01015004"]
```

I also decided to remove comletely *Fagus sylvatica* as we have many in France and the two populations do not overlap well nor does it extend the data range (contrary to *Larix decidua* for instance).

```{r}
tree_dt = tree_dt[!((speciesName_sci == "Fagus sylvatica") & (origin == "swiss"))]
```

```{r}
ls_species = tree_dt[, unique(speciesName_sci)]
for (sp in ls_species)
{
	plot(tree_dt_copy[sp, bole_volume_m3], tree_dt_copy[sp, bole_volume_m3/total_volume_m3],
		col = tree_dt_copy[sp, colour], xlab = "Bole volume", ylab = "r", axes = FALSE, pch = 19,
		cex = 0.85, main = sp)
	axis(1)
	axis(2, las = 1)
	plot(tree_dt[sp, bole_volume_m3], tree_dt[sp, bole_volume_m3/total_volume_m3], col = tree_dt[sp, colour],
		xlab = "Bole volume", ylab = "r", axes = FALSE, pch = 19, cex = 0.85, main = sp)
	axis(1)
	axis(2, las = 1)
}
```

## First run brute force!

```{r}
## Compile model
model = cmdstan_model("./yang2024.stan")

## Common variables
n_chains = 4

savePath = "./yang/"
if (!dir.exists(savePath))
	dir.create(savePath)

## Run model
for (sp in ls_species)
{
	print(paste("Running", sp))
	
	# Subset to targeted species for stanData
	stanData = list(
		N = tree_dt[.(sp)][, .N],
		bole_volume_m3 = tree_dt[.(sp)][, bole_volume_m3],
		total_volume_m3 = tree_dt[.(sp)][, total_volume_m3]
	)

	filename = stri_replace(str = sp, regex = " ", replacement = "-")

	## Run model
	if (file.exists(paste0(savePath, filename, ".rds")))
	{
		fit = readRDS(paste0(savePath, filename, ".rds"))
	} else {
		fit = model$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
			iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

		## Save results
		fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
		saveRDS(fit, paste0(savePath, filename, ".rds"))
	}

	div = plot_sp(fit, sp, simplif = FALSE, n_bins = 4, pal = "Hiroshige", selected_variable = "height",
		print_plot = FALSE)

	if (div[["any_div"]])
		warning(paste("Divergences for", sp))
}
```

I found there are divergences for *Pinus laricio*. This is a warning to remove this species! Here are a bit more investigations

```{r}
## Load Pinus laricio results
sp = "Pinus laricio"
filename = stri_replace(str = sp, regex = " ", replacement = "-")
fit = readRDS(paste0(savePath, filename, ".rds"))

div = plot_sp(fit, sp, simplif = FALSE, n_bins = 4, pal = "Hiroshige", selected_variable = "height",
	print_plot = FALSE)

plot_divergences(fit, div[["loc"]], simplif = FALSE)
```

## Simplified run for species that have multimodal likelihood

Some species have multimodal likelihood due to the merge of intermediate stage with "adult" stage. For this species, a simpler sub-model is required, defined by @eq-yang2024_noIntermediate.

```{r}
# tree_dt[, any(is.na(bole_volume_m3))]
# tree_dt[is.na(bole_volume_m3), 100*circumference_m]
## Compile model
model = cmdstan_model("./yang2024_submodel.stan")

for (sp in c("Fraxinus excelsior", "Larix decidua", "Pinus laricio"))
{
	print(paste("Running", sp))
	
	# Subset to targeted species for stanData
	stanData = list(
		N = tree_dt[.(sp)][, .N],
		bole_volume_m3 = tree_dt[.(sp)][, bole_volume_m3],
		total_volume_m3 = tree_dt[.(sp)][, total_volume_m3]
	)

	filename = paste0(stri_replace(str = sp, regex = " ", replacement = "-"), "_submodel")

	## Run model
	if (file.exists(paste0(savePath, filename, ".rds")))
	{
		fit = readRDS(paste0(savePath, filename, ".rds"))
	} else {
		fit = model$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
			iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

		## Save results
		fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
		saveRDS(fit, paste0(savePath, filename, ".rds"))
	}

	div = plot_sp(fit, sp, simplif = TRUE, n_bins = 4, pal = "Hiroshige", selected_variable = "height",
		print_plot = TRUE)

	if (div[["any_div"]])
		warning(paste("Divergences for", sp))
}
```
