---
title: "Simple tests of bole volume models"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
# rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(inventR)
library(interp)
# library(DHARMa)
library(shape)
library(loo)

## Tool function
source("./toolFunctions.R")

plot_sp = function(fit, sp, count, stanData, fun, args, forest = forest_dt[.(sp)], pred = FALSE,
	n_subsample = 10, folder = NULL, add = TRUE, n_bins = 12, pal = "Hiroshige", selected_variable = NULL)
{
	# Get parameters
	start = stanData[["start"]][count]
	end = stanData[["end"]][count]
	n_sampling = fit$metadata()$iter_sampling
	n_chains = fit$num_chains()

	if (n_chains != nrow(fit$metadata()$time))
	{
		warning("Some chains are missing probably because they could not start! Using the number of successful chains")
		n_chains = nrow(fit$metadata()$time)
	}

	# Add colour to the measured points according to selected variable
	if (!is.null(selected_variable))
	{
		if (!(selected_variable %in% colnames(forest)))
			stop(paste0("'", selected_variable, "' is not a column of forest"))
		
		colours = MetBrewer::met.brewer(pal, n_bins)[1:n_bins]
		forest[, colour_ind := as.numeric(cut(.SD[[selected_variable]], breaks = n_bins))] # Indices per category
		forest[, colour := colours[colour_ind]] # Map selected variable to colour using indices
	} else {
		forest[, colour := "#3355AA33"] # Default colour
	}

	# Plot figure
	if (!is.null(folder))
		png(paste0("./", folder, "/", sp, ".png"))
	
	plot(forest[, bole_volume_m3], forest[, total_ratio], pch = 19, col = forest[, colour], axes = FALSE,
		xlab = "Bole volume", ylab = "Ratio bole/tot volumes", lwd = 0, cex = 0.75,
		main = ifelse(is.null(selected_variable), "", selected_variable))
	if (!is.null(selected_variable))
		legend("topright", legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins)]),
			fill = colours, title = paste(selected_variable, "classes"))
	axis(1)
	axis(2, las = 1)

	# fwrite(forest[, .(bole_volume_m3, total_ratio)],
	# 	"../../../subventions/2025-ademe_graine/niko_doc/data/ratio.csv") # pgfplots Niko

	# Plot prediction
	if (pred)
	{
		sim = fit$draws("r_gen")[sample(x = 1:n_sampling, size = n_subsample, replace = FALSE), , start:end]
		for (i in 1:forest[, .N])
			points(rep(forest[i, bole_volume_m3], n_chains*n_subsample),
				as.numeric(sim[, , i]), col = "#335577AA", pch = ".")
	}

	# Plot fitted curve
	if (length(args) == 2 && all(c("k", "lambda") == args))
	{
		selected_params = c(paste0("k[", count, "]"), paste0("lambda[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, params[1], params[2]), col = "#CD212A", lwd = 2, add = add, n = 501)
	}
	
	if (length(args) == 3 && all(c("beta_", "gamma", "b") == args))
	{
		selected_params = c(paste0("beta_[", count, "]"),
			paste0("gamma[", count, "]"), paste0("b[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, 1, params[1], params[2], params[3]), col = "#CD212A", lwd = 2, n = 501, add = add)
	}

	if (length(args) == 3 && all(c("m", "k", "d") == args))
	{
		selected_params = c(paste0("m[", count, "]"), paste0("k[", count, "]"), paste0("d[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, params[1], params[2], params[3]), col = "#CD212A", lwd = 2, n = 501, add = add)
	}

	if (length(args) == 4 && all(c("alpha", "beta_", "gamma", "b") == args))
	{
		selected_params = c(paste0("alpha[", count, "]"), paste0("beta_[", count, "]"),
			paste0("gamma[", count, "]"), paste0("b[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, params[1], params[2], params[3], params[4]), col = "#CD212A", lwd = 2, add = add, n = 501)
	}
	if (!is.null(folder))
		dev.off()
}
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

opt = "_full"

## Loading
forest_dt = readRDS(paste0(path_data, "nfi-inra_dt", opt, ".rds"))
ind_species = readRDS(paste0(path_data, "ind_species_nfi-inra", opt, ".rds"))

colours = paste0(met.brewer("Egypt")[1:4])
names(colours) = c("conifer", "broadleaf", "inra", "nfi")

forest_dt[fct_type == "broadleaf", fct_colour := colours["broadleaf"]]
forest_dt[fct_type == "conifer", fct_colour := colours["conifer"]]
forest_dt[origin_data == "inra", origin_colour := colours["inra"]]
forest_dt[origin_data == "nfi", origin_colour := colours["nfi"]]

## Add vernacular english names
# vernacularNames = sci2comm(sci = forest_dt[, unique(speciesName_sci)], db = "itis")
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script try to parametrise the total volume in function of the bole volume, that is to say, to use the basic definition of allometry: the dimension of an organ $y$ is proportional to an organ $x$, such as in this power-law $y = a x^b$. Here, $y$ is the total volume and $x$ is the bole volume. In @Yang2024, they do it with above-ground biomass as $y$ and stem biomass as $x$. I expect to be able to apply the same type of functions to volume, as volume and biomass are related by wood density. It should be the same if wood density is constant between the crown and the bole, and only the scale should change! According to @Knapic2011 there is little variation within tree oaks, while there can be for pines @Repola2006.

Here is an intuition on how the function from @Yang2024 works. The parameter $j$ influences the curvature of the function as claimed in the appendix S1 of that paper, but not intuitive: The lower $j$, the stepper the function! However, it is correct to assume that $j < 1$ as otherwise the function is first convex and then concave. So the stan code should reflect this constraint, although the model is perfectly defined for $j > 1$.

```{r}
#| label: fig-var_j
#| fig-cap: Different shape of the function $r = \exp[j - kx] \left( \frac{kx}{j} \right)^j$ for k = 0.06 and j between 0 and 1.6. Dashed lines are for j > 1
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

f = function(x, j, k)
	return (exp(j - k*x) * (k*x/j)^j)

k = 0.06
Homer = met.brewer("Homer1")
curve(f(x, 0.2, k), from = 0, to = 25, lwd = 3, axes = FALSE, xlab = "SB", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (j in 2:8)
	curve(f(x, j/5, k), add = TRUE, col = Homer[j], lwd = 3, lty = ifelse(j/5 < 1, 1, 2))
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = (1:8)/5, bty = "n")
```

```{r}
#| label: fig-var_k
#| fig-cap: Different shape of the function $r = \exp[j - kx] \left( \frac{kx}{j} \right)^j$ for j = 0.5 and k between 0.02 and 0.16.
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

j = 0.6
curve(f(x, j, 0.02), from = 0, to = 25, lwd = 3, axes = FALSE, xlab = "SB", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (k in 2:8)
	curve(f(x, j, k/50), add = TRUE, col = Homer[k], lwd = 3)
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = (1:8)/50, bty = "n")
```
