---
title: "Simple tests of bole volume models"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(inventR)
library(interp)
library(shape)
library(loo)

## Tool function
source("./toolFunctions.R")

plot_sp = function(fit, sp, count, stanData, fun, args, forest = forest_dt[.(sp)], pred = FALSE,
	n_subsample = 10, folder = NULL, add = TRUE, n_bins = 12, pal = "Hiroshige", selected_variable = NULL)
{
	# Get parameters
	start = stanData[["start"]][count]
	end = stanData[["end"]][count]
	n_sampling = fit$metadata()$iter_sampling
	n_chains = fit$num_chains()

	if (n_chains != nrow(fit$metadata()$time))
	{
		warning("Some chains are missing probably because they could not start! Using the number of successful chains")
		n_chains = nrow(fit$metadata()$time)
	}

	# Add colour to the measured points according to selected variable
	if (!is.null(selected_variable))
	{
		if (!(selected_variable %in% colnames(forest)))
			stop(paste0("'", selected_variable, "' is not a column of forest"))
		
		colours = MetBrewer::met.brewer(pal, n_bins)[1:n_bins]
		forest[, colour_ind := as.numeric(cut(.SD[[selected_variable]], breaks = n_bins))] # Indices per category
		forest[, colour := colours[colour_ind]] # Map selected variable to colour using indices
	} else {
		forest[, colour := "#3355AA33"] # Default colour
	}

	# Plot figure
	if (!is.null(folder))
		png(paste0("./", folder, "/", sp, ".png"))
	
	plot(forest[, bole_volume_m3], forest[, total_ratio], pch = 19, col = forest[, colour], axes = FALSE,
		xlab = "Bole volume", ylab = "Ratio bole/tot volumes", lwd = 0, cex = 0.75,
		main = ifelse(is.null(selected_variable), "", selected_variable))
	if (!is.null(selected_variable))
		legend("topright", legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins)]),
			fill = colours, title = paste(selected_variable, "classes"))
	axis(1)
	axis(2, las = 1)

	# fwrite(forest[, .(bole_volume_m3, total_ratio)],
	# 	"../../../subventions/2025-ademe_graine/niko_doc/data/ratio.csv") # pgfplots Niko

	# Plot prediction
	if (pred)
	{
		sim = fit$draws("r_gen")[sample(x = 1:n_sampling, size = n_subsample, replace = FALSE), , start:end]
		for (i in 1:forest[, .N])
			points(rep(forest[i, bole_volume_m3], n_chains*n_subsample),
				as.numeric(sim[, , i]), col = "#335577AA", pch = ".")
	}

	# Plot fitted curve
	if (length(args) == 2 && all(c("k", "lambda") == args))
	{
		selected_params = c(paste0("k[", count, "]"), paste0("lambda[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, params[1], params[2]), col = "#CD212A", lwd = 2, add = add, n = 501)
	}
	
	if (length(args) == 3 && all(c("beta_", "gamma", "b") == args))
	{
		selected_params = c(paste0("beta_[", count, "]"),
			paste0("gamma[", count, "]"), paste0("b[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, 1, params[1], params[2], params[3]), col = "#CD212A", lwd = 2, n = 501, add = add)
	}

	if (length(args) == 3 && all(c("m", "k", "d") == args))
	{
		selected_params = c(paste0("m[", count, "]"), paste0("k[", count, "]"), paste0("d[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, params[1], params[2], params[3]), col = "#CD212A", lwd = 2, n = 501, add = add)
	}

	if (length(args) == 4 && all(c("alpha", "beta_", "gamma", "b") == args))
	{
		selected_params = c(paste0("alpha[", count, "]"), paste0("beta_[", count, "]"),
			paste0("gamma[", count, "]"), paste0("b[", count, "]"))
		params = apply(X = fit$draws(args)[, , selected_params], MARGIN = 3, FUN = mean)

		curve(fun(x, params[1], params[2], params[3], params[4]), col = "#CD212A", lwd = 2, add = add, n = 501)
	}
	if (!is.null(folder))
		dev.off()
}
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

opt = "_inra"

# Loading Emerge data
inra = readRDS(paste0(path_data, "nfi-inra_dt", opt, ".rds"))
inra = inra[, .(speciesName_sci, tree_id = unique_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = taper_height_flo, bole_volume_m3, total_volume_m3)]
setkey(inra, speciesName_sci)

# Loading Swiss data
swiss = readRDS(paste0(mnt_point, "data/switzerland.rds"))
swiss = swiss[, .(speciesName_sci, tree_id, plot_id, fct_type, year, circumference_m, height, taper_height = hdecro,
	bole_volume_m3, total_volume_m3)]
setkey(swiss, speciesName_sci)

## Modern Emerge data
emerge = readRDS(paste0(mnt_point, "data/emerge_2009-2010.rds"))
emerge[, year := as.integer(stringi::stri_replace_all(str = dataset, replacement = "", regex = "emerge_"))]
emerge = emerge[, .(speciesName_sci, tree_id, fct_type, year, circumference_m, height, taper_height,
	bole_volume_m3 = bole_volume_conic_m3, total_volume_m3)]

## Bind everything
tree_dt = rbindlist(list(inra = inra, swiss = swiss, emerge = emerge), idcol = "origin", fill = TRUE)

# Add colours France vs Swiss
ls_colours = met.brewer("Egypt")[1:4]
names(ls_colours) = c("fr_comp", "fr", "ch_comp", "ch")
tree_dt[, colour := ls_colours["fr"]] # Inra and Emerge, in blue
tree_dt[origin == "swiss", colour := ls_colours["ch"]] # Swiss, in yellow

# Remove NAs
tree_dt = tree_dt[!is.na(bole_volume_m3)]

# Remove species with less than 150 individuals
tree_dt[, n_indiv := .N, by = speciesName_sci]
tree_dt = tree_dt[n_indiv > 150]

tree_dt[, .N, by = .(speciesName_sci, colour)][order(speciesName_sci)]

# Set key to sort by species and colour (i.e., France vs Switzerland)
setkey(tree_dt, speciesName_sci, colour)
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script try to parametrise the total volume in function of the bole volume, that is to say, to use the basic definition of allometry: the dimension of an organ $y$ is proportional to an organ $x$, such as in this power-law $y = a x^b$. Here, $y$ is the total volume and $x$ is the bole volume. In @Yang2024, they do it with above-ground biomass as $y$ and stem biomass as $x$. I expect to be able to apply the same type of functions to volume, as volume and biomass are related by wood density. It should be the same if wood density is constant between the crown and the bole, and only the scale should change! According to @Knapic2011 there is little variation within tree oaks, while there can be for pines @Repola2006.

Here is an intuition on how the function @eq-yang2024 from @Yang2024, works.
$$
	r = (m - c) \exp[j - kx] \left( \frac{kx}{j} \right)^j + c - (c - n)\exp[-sx]
$${#eq-yang2024}
The parameter $j$ influences the curvature of the function as claimed in the appendix S1 of that paper, but not intuitive: The lower $j$, the stepper the function! However, it is correct to assume that $j < 1$ as otherwise the function is first convex and then concave. So the stan code should reflect this constraint, although the model is perfectly defined for $j > 1$.

```{r}
#| label: fig-var_j
#| fig-cap: Different shape of the function $r = \exp[j - kx] \left( \frac{kx}{j} \right)^j$ for k = 0.06 and j between 0 and 1.6. Dashed lines are for j > 1
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

f = function(x, j, k)
	return (exp(j - k*x) * (k*x/j)^j)

k = 0.06
Homer = met.brewer("Homer1")
curve(f(x, 0.2, k), from = 0, to = 25, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (j in 2:8)
	curve(f(x, j/5, k), add = TRUE, col = Homer[j], lwd = 3, lty = ifelse(j/5 < 1, 1, 2))
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = (1:8)/5, bty = "n")
```

```{r}
#| label: fig-var_k
#| fig-cap: Different shape of the function $r = \exp[j - kx] \left( \frac{kx}{j} \right)^j$ for j = 0.5 and k between 0.02 and 0.16.
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

j = 0.6
curve(f(x, j, 0.02), from = 0, to = 25, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (k in 2:8)
	curve(f(x, j, k/50), add = TRUE, col = Homer[k], lwd = 3)
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = (1:8)/50, bty = "n")
```

```{r}
#| label: fig-var_s
#| fig-cap: Different shape of the function @eq-yang2024 for j = 0.5, k = , m = , n = , c = and s between 0.02 and 0.16.
#| fig-subcap:
#|   - "For Mammea"
#|   - "For Triplochiton"   
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 4
#| layout-ncol: 2

f = function(x, c, j, k, m, n, s)
	return ((m - c)*exp(j - k*x) * (k*x/j)^j + c - (c - n)*exp(-s*x))

# Values for Mammea taken from Yang et al. 2024
m = 0.65
n = 0.4
c = 0.6
P = 150
j = 1 - 0.5*exp(-0.001*P)
k = 0.004557
print(paste("s =", round(10/P, 3)))

curve(f(x, c, j, k, m, n, 10/P), from = 0, to = 500, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (s in 2:8)
	curve(f(x, c, j, k, m, n, s*10/P), add = TRUE, col = Homer[s], lwd = 3)
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = round((1:8)*10/P, 3), bty = "n")


curve(f(x, c, j, k, m, n, 10/P), from = 500, to = 2000, lwd = 3, axes = FALSE, xlab = "Bole volume",
	ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
for (s in 2:8)
	curve(f(x, c, j, k, m, n, s*10/P), add = TRUE, col = Homer[s], lwd = 3)
legend(x = "bottomright", fill = met.brewer("Homer1"), legend = round((1:8)*10/P, 3), bty = "n")
```

*Larix decidua* seems to be a particular case of the the model @eq-yang2024. It seems to be a simpler convex function with only two stages in its ratio rather than three: First an sharp increase, as for any other species, which corresponds to the juvenile stage, and then a stable phase, which is the asymptotic ratio of adults. There seems to be no intermediate stage. By adding some constraints, and choosing $s = k$ since the intermediate stage merged with the stable stage, and $j = 1$, we get the following type of function:

$$
	r = \alpha + \exp[-\beta x](\gamma x + \delta)
$${#eq-yang2024_noIntermediate}

```{r}
#| label: fig-subcase
#| fig-cap: Particular case of the generic function @eq-yang2024, that reduces to @eq-yang2024_noIntermediate when $j = 1$, $s = k$, $c = \alpha$, $k = \beta$, $m = e^{-1}\frac{\gamma}{\beta} + \alpha$, and $n = \delta + \alpha$.
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

alpha = 0.85
beta = 0.02
gamma = -0.01
delta = -0.7

j = 1
s = beta
c = alpha
k = beta
m = exp(-1)*gamma/k + alpha
n = delta + alpha

curve(f(x, c, j, k, m, n, s), from = 0, to = 500, lwd = 3, axes = FALSE, xlab = "Bole volume", ylab = "r", col = Homer[1])
axis(1)
axis(2, las = 1)
```

Lastly, I use some beta priors in my model. As I am not that familiar with this distribution, it worth it to have some basic plots (Fig. @fig-beta_distrib).

```{r}
#| label: fig-beta_distrib
#| fig-cap: Different shapes of the beta distribution
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

curve(dbeta(x, shape1 = 5, shape2 = 2), from = 0, to = 1, lwd = 3, axes = FALSE, xlab = "x",
	ylab = "Beta pdf", col = Homer[1], ylim = c(0, 3.6))
axis(1)
axis(2, las = 1)

curve(dbeta(x, shape1 = 8, shape2 = 2), add = TRUE, col = Homer[2], lwd = 3)

curve(dbeta(x, shape1 = 8, shape2 = 1), add = TRUE, col = Homer[3], lwd = 3)

curve(dbeta(x, shape1 = 5, shape2 = 5), add = TRUE, col = Homer[4], lwd = 3)

curve(dbeta(x, shape1 = 1.5, shape2 = 1.5), add = TRUE, col = Homer[5], lwd = 3)

curve(dbeta(x, shape1 = 2, shape2 = 5), add = TRUE, col = Homer[8], lwd = 3)

curve(dbeta(x, shape1 = 2, shape2 = 8), add = TRUE, col = Homer[7], lwd = 3)

curve(dbeta(x, shape1 = 1, shape2 = 8), add = TRUE, col = Homer[6], lwd = 3)

legend(x = "topright", fill = met.brewer("Homer1")[c(1:5, 8, 7, 6)],
	legend = c("5, 2", "8, 2", "8, 1", "5, 5", "1.5, 1.5", "2, 5", "2, 8", "1, 8"), bty = "n")
```

## Check the overlap of the ratio data
The Swiss data have an abnormal number of ratio values near one... It happens on very concentrated years and plots, which makes me think 
```{r}
library(terra)
swiss_coords = vect(paste0(mnt_point, "data/coords_switzerland"))

tree_dt[, r := bole_volume_m3/total_volume_m3]

aa = tree_dt[origin != "swiss"][r > 0.98][, .N, by = .(plot_id, year)]
bb = tree_dt[origin != "swiss"][, .(n_tot = .N), by = .(plot_id, year)]

aa = merge.data.table(aa, bb, by = c("plot_id", "year"))
aa[, prop := round(100*N/n_tot, 1)]
aa[, mean(prop)] # In average for plots having trees with r > 0.98, it is 10% of the indiv that are concerned

ls_pb = tree_dt[origin == "swiss"][r > 0.98][, .N, by = .(plot_id, year)]
n_indiv = tree_dt[origin == "swiss"][, .(n_tot = .N), by = .(plot_id, year)]

ls_pb = merge.data.table(ls_pb, n_indiv, by = c("plot_id", "year"))
ls_pb[, prop := round(100*N/n_tot, 1)]

ls_pb[prop > 20]

ls_pb[N > 20, sum(N)]*100/ls_pb[, sum(N)]
ls_pb[N > 50][order(year)]

pb_66 = ls_pb[N > 50, unique(plot_id)] # 2/3rd of the problematic trees come from theses plots

setkey(ls_pb, plot_id, year)

ls_pb_10p = ls_pb[prop > 10, .(plot_id, year)] # "01015004"
ls_pb[ls_pb_10p]
tree_dt[plot_id %in% ls_pb_10p$plot_id, .N, by = speciesName_sci]

switzerland = project(vect(paste0(mnt_point, "data/shapefiles/switzerland/CHE_adm0.shp")), swiss_coords)

## Plots
swiss_cols = met.brewer("Egypt")[1:4]
names(swiss_cols) = unique(swiss_coords$region)
swiss_coords$colour = swiss_cols[swiss_coords$region]

plot(switzerland, lwd = 2, axes = FALSE)
points(swiss_coords, pch = 20, cex = 0.55)
points(swiss_coords[swiss_coords$plot_id %in% ls_pb_10p$plot_id],
	col = swiss_coords[swiss_coords$plot_id %in% ls_pb_10p$plot_id]$colour,
	pch = 4, cex = 3)

plot(switzerland, lwd = 2, axes = FALSE)
points(swiss_coords, pch = 20, cex = 0.55)
points(swiss_coords[swiss_coords$plot_id %in% pb_66],
	col = swiss_coords[swiss_coords$plot_id %in% pb_66]$colour,
	pch = 4, cex = 4)

dt = as.data.table(swiss_coords[swiss_coords$plot_id %in% pb_66])

swiss_coords[(swiss_coords$plot_id %in% ls_pb_10p$plot_id) & (swiss_coords$year %in% ls_pb_10p$year)]

plot(tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), bole_volume_m3],
	tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), r],
	pch = 19, xlab = "Bole volume", ylab = "Ratio")

plot(tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), circumference_m],
	tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98)), r],
	pch = 19, xlab = "Circumference", ylab = "Ratio")

100*tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year))][r > 0.98, .N]/
	tree_dt[((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year))][, .N]
```

## Removing some trees from the Swiss data

As demonstrated in the previous section, the overlap is not that good for some very particular plots where many trees have a ratio greater than 0.98! I decided to remove trees with a ratio $r > 0.98$ and belonging to plots with more than 10% of such trees.

```{r}
tree_dt_copy = copy(tree_dt)
tree_dt = tree_dt[!((plot_id %in% ls_pb_10p$plot_id) & (year %in% ls_pb_10p$year) & (r > 0.98))]
# tree_dt[plot_id == "01015004"]
```

I also decided to remove comletely *Fagus sylvatica* as we have many in France and the two populations do not overlap well nor does it extend the data range (contrary to *Larix decidua* for instance).

```{r}
tree_dt = tree_dt[!((speciesName_sci == "Fagus sylvatica") & (origin == "swiss"))]
```

```{r}
ls_species = tree_dt[, unique(speciesName_sci)]
for (sp in ls_species)
{
	plot(tree_dt_copy[sp, bole_volume_m3], tree_dt_copy[sp, bole_volume_m3/total_volume_m3],
		col = tree_dt_copy[sp, colour], xlab = "Bole volume", ylab = "r", axes = FALSE, pch = 19,
		cex = 0.85, main = sp)
	axis(1)
	axis(2, las = 1)
	plot(tree_dt[sp, bole_volume_m3], tree_dt[sp, bole_volume_m3/total_volume_m3], col = tree_dt[sp, colour],
		xlab = "Bole volume", ylab = "r", axes = FALSE, pch = 19, cex = 0.85, main = sp)
	axis(1)
	axis(2, las = 1)
}
```

## First run brute force!

```{r}
## Compile model
model = cmdstan_model("./yang2024.stan")

## Prepare data
# Choose species
sp = "Fagus sylvatica"

# Subset to targeted species for stanData
stanData = list(
	N = tree_dt[.(sp)][, .N],
	bole_volume_m3 = tree_dt[.(sp)][, bole_volume_m3],
	total_volume_m3 = tree_dt[.(sp)][, total_volume_m3]
)

## Common variables
n_chains = 4

savePath = "./yang/"
if (!dir.exists(savePath))
	dir.create(savePath)

filename = stri_replace(str = sp, regex = " ", replacement = "-")

## Run model
if (file.exists(paste0(savePath, filename, ".rds")))
{
	fit = readRDS(paste0(savePath, filename, ".rds"))
} else {
	fit = model$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
		iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

	## Save results
	fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
	saveRDS(fit, paste0(savePath, filename, ".rds"))
}
```

### Graphics to check the run

```{r}
#| label: fig-pred-allom
#| fig-cap: Predictions of above-ground volume in function of bole volume and the parametrised fitting curve
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

psi = function(x, pars)
{
	r = (pars["m"] - pars["c"]) * exp(pars["j"] - pars["k"]*x) * (pars["k"]*x/pars["j"])^pars["j"] +
		pars["c"] - (pars["c"] - pars["n"])*exp(-pars["s"]*x)

	return ( 1/pars["c"] * x^( 1 - (log(r) - log(pars["c"])) / log(x) ) )
}


paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")

plot(stanData$bole_volume_m3, stanData$total_volume_m3, pch = 19, cex = 0.65, axes = FALSE,
	col = tree_dt[sp, colour], xlab = "Bole volume", ylab = "Above-ground volume")
curve(psi(x, paramsVec), lwd = 4, add = TRUE)
abline(a = 0, b = 1, lty = "dashed")
axis(1)
axis(2, las = 1)
```

```{r}

lazyTrace(fit$draws("c"), main = "c")
lazyTrace(fit$draws("j"), main = "j")
lazyTrace(fit$draws("k"), main = "k")
lazyTrace(fit$draws("m"), main = "m")
lazyTrace(fit$draws("n"), main = "n")
lazyTrace(fit$draws("s"), main = "s")

params = getParams(model_cmdstan = fit, params_names = c("c", "j", "m", "k"), type = "all")
div = posterior::subset_draws(fit$sampler_diagnostics(), variable = "divergent__")

if (any(div != 0))
{
	div_c = posterior::subset_draws(params[, , "c"], draw = which(div == 1))
	div_j = posterior::subset_draws(params[, , "j"], draw = which(div == 1))
	div_m = posterior::subset_draws(params[, , "m"], draw = which(div == 1))
	div_k = posterior::subset_draws(params[, , "k"], draw = which(div == 1))

	# Plot divergence for c and j
	plot(params[, , "c"], params[, , "j"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "j", main = "c and j")
	points(div_c, div_j, pch = 18, col = "#0F7BA2")

	# Plot divergence for c and k
	plot(params[, , "c"], params[, , "k"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "k", main = "c and k")
	points(div_c, div_k, pch = 18, col = "#0F7BA2")

	# Plot divergence for c and m
	plot(params[, , "c"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "m", main = "c and m")
	points(div_c, div_m, pch = 18, col = "#0F7BA2")

	# Plot divergence for j and k
	plot(params[, , "j"], params[, , "k"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "j", ylab = "k", main = "j and k")
	points(div_j, div_k, pch = 18, col = "#0F7BA2")

	# Plot divergence for j and m
	plot(params[, , "j"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "j", ylab = "m", main = "j and m")
	points(div_j, div_m, pch = 18, col = "#0F7BA2")

	# Plot divergence for k and m
	plot(params[, , "k"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "k", ylab = "m", main = "k and m")
	points(div_k, div_m, pch = 18, col = "#0F7BA2")
}
```

```{r}
# if (any(div != 0))
# 	fit$cmdstan_diagnose()
```

## Predictions ...
### ... in ratio
```{r}
#| label: fig-pred-ratio
#| fig-cap: Predictions of ratio with non-uniqueness
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

f = function(x, pars)
{
	# The vector of parameters, pars, is in this order:
	# 1 -> c,
	# 2 -> j,
	# 3 -> k,
	# 4 -> m,
	# 5 -> n,
	# 6 -> s
	return ((pars["m"] - pars["c"]) * exp(pars["j"] - pars["k"]*x) * (pars["k"]*x/pars["j"])^pars["j"] +
		pars["c"] - (pars["c"] - pars["n"])*exp(-pars["s"]*x));
}

paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")
paramsVec1 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
	type = "chain", chain = 1)
paramsVec2 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
	type = "chain", chain = 2)
paramsVec3 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
	type = "chain", chain = 3)
paramsVec4 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
	type = "chain", chain = 4)

plot(stanData$bole_volume_m3, stanData$bole_volume_m3/stanData$total_volume_m3, pch = 19, cex = 0.65,
	col = tree_dt[sp, colour], xlab = "Bole volume", ylab = "r", axes = FALSE)
curve(f(x, paramsVec), lwd = 4, add = TRUE)
curve(f(x, paramsVec_fr), lwd = 4, add = TRUE, col = ls_colours["fr_comp"])
curve(f(x, paramsVec_ch), lwd = 4, add = TRUE, col = ls_colours["ch_comp"])
curve(f(x, paramsVec1), lwd = 4, add = TRUE, col = met.brewer("Egypt")[1])
curve(f(x, paramsVec2), lwd = 4, add = TRUE, col = met.brewer("Egypt")[2])
curve(f(x, paramsVec3), lwd = 4, add = TRUE, col = met.brewer("Egypt")[3])
curve(f(x, paramsVec4), lwd = 4, add = TRUE, col = met.brewer("Egypt")[4])
abline(v = paramsVec["j"]/paramsVec["k"], lty = "dashed", lwd = 0.6)
abline(h = paramsVec["m"] - (paramsVec["c"] - paramsVec["n"])*exp(-paramsVec["s"]*paramsVec["j"]/paramsVec["k"]),
	lty = "dashed")
legend("bottomright", legend = paste("Chain", 1:4), fill = met.brewer("Egypt"))
axis(1)
axis(2, las = 1)

# curve(f(x, paramsVec), lwd = 4, to = 8)
# abline(v = paramsVec["j"]/paramsVec["k"], lty = "dashed")
# abline(h = paramsVec["m"] - (paramsVec["c"] - paramsVec["n"])*exp(-paramsVec["s"]*paramsVec["j"]/paramsVec["k"]),
# 	lty = "dashed")

# (paramsVec["c"] - paramsVec["n"])*exp(-paramsVec["s"]*paramsVec["j"]/paramsVec["k"])

# paramsVec["s"] - 3*paramsVec["k"]/paramsVec["j"]

# lazyPosterior(fit$draws("c"), fun = dbeta, shape1 = 8, shape2 = 1)

## Save results
params_dt = as.data.table(x = paramsVec, keep.rownames = "parameter")
params_dt = merge.data.table(params_dt, as.data.table(x = paramsVec1, keep.rownames = "parameter"),
	by = "parameter")
params_dt = merge.data.table(params_dt, as.data.table(x = paramsVec2, keep.rownames = "parameter"),
	by = "parameter")
params_dt = merge.data.table(params_dt, as.data.table(x = paramsVec3, keep.rownames = "parameter"),
	by = "parameter")
params_dt = merge.data.table(params_dt, as.data.table(x = paramsVec4, keep.rownames = "parameter"),
	by = "parameter")

setnames(params_dt, old = paste0("paramsVec", c("", 1:4)), new = c("avg", paste0("chain", 1:4)))
fwrite(params_dt, "~/work/writing/article_allometry/data/params_Larix-decidua_yang2024.csv")

fwrite(data.table(bole_volume_m3 = stanData$SB, ratio = stanData$SB/stanData$AGB),
	"~/work/writing/article_allometry/data/larix-decidua_data.csv")
```

### ... in volume

```{r}
#| label: fig-pred-volume
#| fig-cap: Predictions in volume
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

vol_pred = apply(X = fit$draws("v_gen_mean"), MARGIN = 3, FUN = mean)

plot(vol_pred, stanData$AGB, pch = 19, cex = 0.65, axes = FALSE,
	xlab = "Above-ground volume (predicted)", ylab = "Above-ground volume (observed)")
abline(a = 0, b = 1, lwd = 3, col = "#CD212A")
axis(1)
axis(2, las = 1)
```

```{r}
#| label: fig-pred-volume-chains
#| fig-cap: Predictions in volume, per chain
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

vol_pred1 = apply(X = fit$draws("v_gen_mean")[, 1, ], MARGIN = 3, FUN = mean)
vol_pred2 = apply(X = fit$draws("v_gen_mean")[, 2, ], MARGIN = 3, FUN = mean)
vol_pred3 = apply(X = fit$draws("v_gen_mean")[, 3, ], MARGIN = 3, FUN = mean)
vol_pred4 = apply(X = fit$draws("v_gen_mean")[, 4, ], MARGIN = 3, FUN = mean)

plot(vol_pred1, stanData$AGB, pch = 19, axes = FALSE, col = met.brewer("Egypt")[1],
	xlab = "Above-ground volume (predicted)", ylab = "Above-ground volume (observed)")
points(vol_pred2, stanData$AGB, pch = 19, cex = 0.65, col = met.brewer("Egypt")[2])
points(vol_pred3, stanData$AGB, pch = 19, cex = 0.45, col = met.brewer("Egypt")[3])
points(vol_pred4, stanData$AGB, pch = 19, cex = 0.25, col = met.brewer("Egypt")[4])
abline(a = 0, b = 1, lwd = 3)
legend("bottomright", legend = paste("Chain", 1:4), fill = met.brewer("Egypt"))
axis(1)
axis(2, las = 1)
```




## Simplified run brute force!

```{r}
# tree_dt[, any(is.na(bole_volume_m3))]
# tree_dt[is.na(bole_volume_m3), 100*circumference_m]
## Compile model
model = cmdstan_model("./yang2024_simplified.stan")

## Prepare data
tree_dt = tree_dt[!is.na(bole_volume_m3)]

stanData = list(
	N = tree_dt["Fagus sylvatica"][origin == "inra", .N],
	SB = tree_dt["Fagus sylvatica"][origin == "inra", bole_volume_m3],
	AGB = tree_dt["Fagus sylvatica"][origin == "inra", total_volume_m3]
)

## Common variables
n_chains = 4

savePath = "./yang/"
if (!dir.exists(savePath))
	dir.create(savePath)

filename = "inra-Fagsyl_simplified"

## Run model
if (file.exists(paste0(savePath, filename, ".rds")))
{
	fit = readRDS(paste0(savePath, filename, ".rds"))
} else {
	fit = model$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
		iter_warmup = 1500, iter_sampling = 500, max_treedepth = 10)

	## Save results
	fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
	saveRDS(fit, paste0(savePath, filename, ".rds"))
}
```

```{r}
#| label: fig-pred-allom-simplif
#| fig-cap: Predictions of above-ground volume in function of bole volume and the parametrised fitting curve
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

psi = function(x, pars)
{
	r = (pars["m"] - pars["c"]) * exp(1 - pars["k"]*x) * pars["k"]*x + pars["c"]

	return ( 1/pars["c"] * x^( 1 - (log(r) - log(pars["c"])) / log(x) ) )
}

paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "mean")

# ff = function(x)
# 	return (10.24*exp(-3.74*x))

# curve(ff, from = 0, to = 5)

plot(stanData$SB, stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
	xlab = "Bole volume", ylab = "Above-ground volume", axes = FALSE)
curve(psi(x, paramsVec), lwd = 4, add = TRUE)
abline(a = 0, b = 1, lty = "dashed")
axis(1)
axis(2, las = 1)
```

```{r}
lazyTrace(fit$draws("c"), main = "c")
lazyTrace(fit$draws("k"), main = "k")
lazyTrace(fit$draws("m"), main = "m")

lazyPosterior(fit$draws("phi0"), fun = dgamma, shape = 3, rate = 0.1)
lazyPosterior(fit$draws("phi1"), fun = dnorm, mean = 3, sd = 1)

params = getParams(model_cmdstan = fit, params_names = c("c", "m", "k"), type = "all")
div = posterior::subset_draws(fit$sampler_diagnostics(), variable = "divergent__")

if (any(div != 0))
{
	div_c = posterior::subset_draws(params[, , "c"], draw = which(div == 1))
	div_m = posterior::subset_draws(params[, , "m"], draw = which(div == 1))
	div_k = posterior::subset_draws(params[, , "k"], draw = which(div == 1))

	# Plot divergence for c and k
	plot(params[, , "c"], params[, , "k"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "k", main = "c and k")
	points(div_c, div_k, pch = 18, col = "#0F7BA2")

	# Plot divergence for c and m
	plot(params[, , "c"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "m", main = "c and m")
	points(div_c, div_m, pch = 18, col = "#0F7BA2")

	# Plot divergence for k and m
	plot(params[, , "k"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "k", ylab = "m", main = "k and m")
	points(div_k, div_m, pch = 18, col = "#0F7BA2")
}
```

```{r}
if (any(div != 0))
	fit$cmdstan_diagnose()
```

## Predictions ...
### ... in ratio
```{r}
#| label: fig-pred-ratio-simplif
#| fig-cap: Predictions of ratio, simplified model
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

f = function(x, pars)
{
	# The vector of parameters, pars, is in this order:
	# 1 -> c,
	# 2 -> k,
	# 3 -> m,
	return ((pars["m"] - pars["c"]) * exp(1 - pars["k"]*x) * (pars["k"]*x) + pars["c"]);
}

paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "mean")
paramsVec1 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 1)
paramsVec2 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 2)
paramsVec3 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 3)
paramsVec4 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 4)

plot(stanData$SB, stanData$SB/stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
	xlab = "Bole volume", ylab = "r", axes = FALSE)
curve(f(x, paramsVec), lwd = 4, add = TRUE)
curve(f(x, paramsVec1), lwd = 4, add = TRUE, col = met.brewer("Egypt")[1])
curve(f(x, paramsVec2), lwd = 4, add = TRUE, col = met.brewer("Egypt")[2])
curve(f(x, paramsVec3), lwd = 4, add = TRUE, col = met.brewer("Egypt")[3])
curve(f(x, paramsVec4), lwd = 4, add = TRUE, col = met.brewer("Egypt")[4])
abline(v = paramsVec["j"]/paramsVec["k"], lty = "dashed", lwd = 0.6)
abline(h = paramsVec["m"] - (paramsVec["c"] - paramsVec["n"])*exp(-paramsVec["s"]*paramsVec["j"]/paramsVec["k"]),
	lty = "dashed")
legend("bottomright", legend = paste("Chain", 1:4), fill = met.brewer("Egypt"))
axis(1)
axis(2, las = 1)

# curve(f(x, paramsVec), lwd = 4, to = 8)
# abline(v = paramsVec["j"]/paramsVec["k"], lty = "dashed")
# abline(h = paramsVec["m"] - (paramsVec["c"] - paramsVec["n"])*exp(-paramsVec["s"]*paramsVec["j"]/paramsVec["k"]),
# 	lty = "dashed")

# (paramsVec["c"] - paramsVec["n"])*exp(-paramsVec["s"]*paramsVec["j"]/paramsVec["k"])

# paramsVec["s"] - 3*paramsVec["k"]/paramsVec["j"]

# lazyPosterior(fit$draws("c"), fun = dbeta, shape1 = 8, shape2 = 1)
```

### ... in volume

```{r}
#| label: fig-pred-volume-simplif
#| fig-cap: Predictions in volume
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

vol_pred = apply(X = fit$draws("v_gen_mean"), MARGIN = 3, FUN = mean)

plot(vol_pred, stanData$AGB, pch = 19, cex = 0.65, axes = FALSE,
	xlab = "Above-ground volume (predicted)", ylab = "Above-ground volume (observed)")
abline(a = 0, b = 1, lwd = 3, col = "#CD212A")
axis(1)
axis(2, las = 1)
```

```{r}
#| label: fig-pred-volume-chains-simplif
#| fig-cap: Predictions in volume, per chain
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8.142857

vol_pred1 = apply(X = fit$draws("v_gen_mean")[, 1, ], MARGIN = 3, FUN = mean)
vol_pred2 = apply(X = fit$draws("v_gen_mean")[, 2, ], MARGIN = 3, FUN = mean)
vol_pred3 = apply(X = fit$draws("v_gen_mean")[, 3, ], MARGIN = 3, FUN = mean)
vol_pred4 = apply(X = fit$draws("v_gen_mean")[, 4, ], MARGIN = 3, FUN = mean)

plot(vol_pred1, stanData$AGB, pch = 19, axes = FALSE, col = met.brewer("Egypt")[1],
	xlab = "Above-ground volume (predicted)", ylab = "Above-ground volume (observed)")
points(vol_pred2, stanData$AGB, pch = 19, cex = 0.65, col = met.brewer("Egypt")[2])
points(vol_pred3, stanData$AGB, pch = 19, cex = 0.45, col = met.brewer("Egypt")[3])
points(vol_pred4, stanData$AGB, pch = 19, cex = 0.25, col = met.brewer("Egypt")[4])
abline(a = 0, b = 1, lwd = 3)
legend("bottomright", legend = paste("Chain", 1:4), fill = met.brewer("Egypt"))
axis(1)
axis(2, las = 1)
```
