---
title: "Fitting two models for Vtot"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(cmdstanr)
	register_knitr_engine(override = TRUE)

## Tool function
source("./toolFunctions.R")

plot_sp = function(fit, sp, stanData, chains = FALSE, sep_chains = FALSE, popol_plot = FALSE)
{
	# Ratio equation
	r = function(x, pars)
		return ((pars["m"] - pars["c"]) * exp(1 - pars["k"]*x) * (pars["k"]*x) + pars["c"]);
	
	# Allometric equation
	psi = function(x, pars)
	{
		r = (pars["m"] - pars["c"]) * exp(1 - pars["k"]*x) * pars["k"]*x + pars["c"]
		return ( 1/pars["c"] * x^( 1 - (log(r) - log(pars["c"])) / log(x) ) )
	}

	pol = function(x, pars)
		return (pars["beta1"]*x + pars["beta2"]*x^2 + pars["beta3"]*x^3 + pars["beta4"]*x^4)
	
	pol_ratio = function(x, pars)
		return (1/(pars["beta1"] + pars["beta2"]*x + pars["beta3"]*x^2 + pars["beta4"]*x^3))

	if (popol_plot)
	{
		# Extract mean values parameters
		paramsVec = getParams(model_cmdstan = fit, params_names = c("beta1", "beta2", "beta3", "beta4"),
			type = "mean")

		Delta = 4*paramsVec["beta3"]^2 - 12*paramsVec["beta2"]*paramsVec["beta4"]
		Delta = ifelse(Delta >= 0, Delta, NA)
		P1 = (-2*paramsVec["beta3"] + sqrt(Delta))/(6*paramsVec["beta4"])
		P2 = (-2*paramsVec["beta3"] - sqrt(Delta))/(6*paramsVec["beta4"])

		# Plot fitting data and fitted curve (ratio)
		plot(stanData$SB, stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
			xlab = "SB", ylab = "r", axes = FALSE, main = sp)
		curve(pol(x, paramsVec), lwd = 4, add = TRUE)
		legend("bottomright", legend = paste("Chain", 1:4), fill = met.brewer("Egypt"))
		abline(a = 0, b = 1, lty = "dashed")
		axis(1)
		axis(2, las = 1)
		
		plot(stanData$SB, stanData$SB/stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
			xlab = "SB", ylab = "r", axes = FALSE, main = sp)
		curve(pol_ratio(x, paramsVec), lwd = 4, add = TRUE)
		legend("bottomright", legend = paste("Chain", 1:4), fill = met.brewer("Egypt"))
		abline(v = P1, lty = "dashed")
		abline(h = pol_ratio(P1, paramsVec), lty = "dashed")
		abline(v = P2, lty = "dashed")
		abline(h = pol_ratio(P2, paramsVec), lty = "dashed")
		axis(1)
		axis(2, las = 1)

		return (list(P_est_1 = P1, P_est_2 = P2, fP1 = pol_ratio(P1, paramsVec), fP2 = pol_ratio(P2, paramsVec)))
	}

	# Plot chains of the three params
	if (chains)
	{
		lazyTrace(fit$draws("c"), main = paste("c,", sp))
		lazyTrace(fit$draws("k"), main = paste("k,", sp))
		lazyTrace(fit$draws("m"), main = paste("m,", sp))
	}

	# Extract mean values parameters
	paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "mean")

	# Plot fitting data and fitted curve (ratio)
	plot(stanData$SB, stanData$SB/stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
		xlab = "SB", ylab = "r", axes = FALSE, main = sp)
	curve(r(x, paramsVec), lwd = 4, add = TRUE)
	legend("bottomright", legend = paste("Chain", 1:4), fill = met.brewer("Egypt"))
	axis(1)
	axis(2, las = 1)

	if (sep_chains)
	{
		paramsVec1 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 1)
		paramsVec2 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 2)
		paramsVec3 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 3)
		paramsVec4 = getParams(model_cmdstan = fit, params_names = c("c", "k", "m"), type = "chain", chain = 4)
		curve(r(x, paramsVec1), lwd = 4, add = TRUE, col = met.brewer("Egypt")[1])
		curve(r(x, paramsVec2), lwd = 4, add = TRUE, col = met.brewer("Egypt")[2])
		curve(r(x, paramsVec3), lwd = 4, add = TRUE, col = met.brewer("Egypt")[3])
		curve(r(x, paramsVec4), lwd = 4, add = TRUE, col = met.brewer("Egypt")[4])
	}

	# Plot fitting data and fitted curve (above-ground volume)
	plot(stanData$SB, stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
		xlab = "Bole volume", ylab = "Above-ground volume", axes = FALSE, main = sp)
	curve(psi(x, paramsVec), lwd = 4, add = TRUE)
	abline(a = 0, b = 1, lty = "dashed")
	axis(1)
	axis(2, las = 1)

	if (sep_chains)
	{
		curve(psi(x, paramsVec1), lwd = 4, add = TRUE, col = met.brewer("Egypt")[1])
		curve(psi(x, paramsVec2), lwd = 4, add = TRUE, col = met.brewer("Egypt")[2])
		curve(psi(x, paramsVec3), lwd = 4, add = TRUE, col = met.brewer("Egypt")[3])
		curve(psi(x, paramsVec4), lwd = 4, add = TRUE, col = met.brewer("Egypt")[4])
	}


	# Results to return
	params = getParams(model_cmdstan = fit, params_names = c("c", "m", "k"), type = "all")
	div = posterior::subset_draws(fit$sampler_diagnostics(), variable = "divergent__")

	if (all(div == 0))
		return(list(is_div = FALSE))
	
	if (any(div != 0))
		return (list(params = params, div = div, is_div = any(div != 0)))	
}

## Function to export traces data for pgfplots
pgfTrace = function(fit, sp, params, modelsname)
{
	traces = setDT(posterior::as_draws_df(fit$draws(params)))[, c(".iteration", ".chain", ..params)]
	setnames(traces, new = c("iter", "chain", "value"))
	fwrite(x = traces,
		file = paste0("traces_", stringi::stri_replace_all(str = sp, replacement = "-", regex = " "),
		"_", modelsname, ".csv"))
}
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

opt = "_inra"

# Loading Emerge data
inra = readRDS(paste0(path_data, "nfi-inra_dt", opt, ".rds"))
inra = inra[, .(speciesName_sci, tree_id = unique_id, fct_type, year, circumference_m, height,
	taper_height = taper_height_flo, bole_volume_m3, total_volume_m3)]

inra[, .(min_c = min(circumference_m), max_c = max(circumference_m)), by = speciesName_sci]

# Loading Swiss data
swiss = readRDS(paste0(mnt_point, "data/switzerland.rds"))
swiss = swiss[, .(speciesName_sci, tree_id, fct_type, year, circumference_m, height, taper_height = hdecro,
	bole_volume_m3, total_volume_m3)]

swiss = swiss[circumference_m >= 0.17] # To harmonise with Oudin's data and Emerge, 0.17 seems a good one!

## Modern Emerge data
emerge = readRDS(paste0(mnt_point, "data/emerge_2009-2010.rds"))
emerge[, year := as.integer(stringi::stri_replace_all(str = dataset, replacement = "", regex = "emerge_"))]
emerge = emerge[, .(speciesName_sci, tree_id, fct_type, year, circumference_m, height, taper_height,
	bole_volume_m3 = bole_volume_conic_m3, total_volume_m3)]

## Merge three datasets
tree_dt = rbindlist(list(inra = inra, swiss = swiss, emerge = emerge), idcol = "origin")
tree_dt = tree_dt[!is.na(bole_volume_m3)] # Remove NAs, check why Swiss has some

setkey(tree_dt, speciesName_sci)

tree_dt[, year := as.integer(year)]

## Add colours based on origin
tree_dt[origin == "emerge", orig_col := "#E76254"]
tree_dt[origin == "inra", orig_col := "#F7AA58"]
tree_dt[origin == "swiss", orig_col := "#1E466E"]

## Compute ratio
tree_dt[, ratio := bole_volume_m3/total_volume_m3]

## Remove strange ratio, i.e., some years contains a high range of ratios near 1
tree_dt[ratio >= 0.99, hist(year)]

N_per_year = tree_dt[, .(N_tot = .N), by = year][order(year)]

lala = tree_dt[ratio >= 0.99, .N, by = year]
lala = merge.data.table(lala, N_per_year, by = "year")
lala[, perc := round(N/N_tot*100, 1)]

strange_years = lala[perc > 30, year]

lala = tree_dt[(ratio >= 0.99) & (year %in% strange_years)]
lala[, .N, by = origin]
plot(lala[, circumference_m], lala[, total_volume_m3])


tree_dt[(ratio >= 0.99) & (year %in% strange_years), orig_col := "#43B284"]

plot(tree_dt[, circumference_m], tree_dt[, total_volume_m3], pch = 19, col = tree_dt[, orig_col])

points(tree_dt[orig_col == "#43B284", circumference_m], tree_dt[orig_col == "#43B284", total_volume_m3], pch = 19, col = "#43B284")

tree_dt = tree_dt[!((ratio >= 0.99) & (year %in% strange_years))]
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script parametrises the total volume with two models. The first one is taken from @Yang2024, and is expected to work with volumes despite it was done on biomass, simply because we heuristically assume a constant density within the tree. This often made assumption holds for *Quercus* [@Knapic2011], but not for *Pinus* [@Repola2006].

## Simplified model of @Yang2024

In this model, the above-ground volume is derived from the bole volume as an allometric function. Basically, allometry assumes that the dimension of an organ $y$ is proportional to an organ $x$, following a power-law $y = a x^b$. Here, $y$ is the total volume and $x$ is the bole volume.

In @Yang2024, the allometry is applied to above-ground biomass as $y$ and stem biomass as $x$. I expect to be able to apply the same type of functions to volume, as volume and biomass are related by wood density. It should be the same if wood density is constant between the crown and the bole, and only the scale should change! According to @Knapic2011 there is little variation within tree oaks, while there can be for pines @Repola2006.

In this section, I use the simplified model from @Yang2024, equation S4:

$$
	r = (m - c)kx \exp[1 - kx] + c
$$

and then use the allometric equation:

$$
	\Vtot = \frac{1}{c} x^{1 - [\ln(r) - \ln(c)]/\ln(\Vbft)}
$$

```{r}
## Compile model
model = cmdstan_model("./yang2024_simplified.stan")

## Common variables
n_chains = 4

savePath = "./yang/"
if (!dir.exists(savePath))
	dir.create(savePath)

ls_species = tree_dt[, .N, by = speciesName_sci][N > 150, sort(speciesName_sci)]

for (sp in ls_species)
{
	print(paste(" Running species", sp))
	filename = paste0(sp, "_simplified")

	stanData = list(
		N = tree_dt[.(sp)][, .N],
		SB = tree_dt[.(sp)][, bole_volume_m3],
		AGB = tree_dt[.(sp)][, total_volume_m3]
	)

	if (!file.exists(paste0(savePath, filename, ".rds")))
	{
		fit = model$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
			iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

		## Save results
		fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
		saveRDS(fit, paste0(savePath, filename, ".rds"))
	} else {
		fit = readRDS(paste0(savePath, filename, ".rds"))
	}

	plot_sp(fit = fit, sp = sp, stanData = stanData, chains = TRUE)
}
```


## Polynomial fit

```{r}
pol4 = function(x, b1, b2, b3, b4)
	return (x/(b1*x + b2*x^2 + b3*x^3 + b4*x^4))

curve(pol4(x, 1, 1, 2, 1), from = -1.5, to = 0.5, axes = FALSE, xlab = "x", ylab = "y")
abline(v = -1)
abline(h = pol4(-1, 1, 1, 2, 1), lty = "dashed")
abline(v = -1/3)
abline(h = pol4(-1/3, 1, 1, 2, 1), lty = "dashed")
axis(1)
axis(2, las = 1)
```

```{r}
## Compile model
model_pol = cmdstan_model("./yang2024_polynomial.stan")

## Common variables
n_chains = 4

savePath = "./yang/"
if (!dir.exists(savePath))
	dir.create(savePath)

ls_species = tree_dt[, .N, by = speciesName_sci][N > 150, sort(speciesName_sci)]

for (sp in ls_species)
{
	print(paste(" Running species", sp))
	filename = paste0(sp, "_polynomial")

	stanData = list(
		N = tree_dt[.(sp)][, .N],
		SB = tree_dt[.(sp)][, bole_volume_m3],
		AGB = tree_dt[.(sp)][, total_volume_m3]
	)

	if (!file.exists(paste0(savePath, filename, ".rds")))
	{
		fit = model_pol$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
			iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

		## Save results
		fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
		saveRDS(fit, paste0(savePath, filename, ".rds"))
	} else {
		fit = readRDS(paste0(savePath, filename, ".rds"))
	}

	plot_sp(fit = fit, sp = sp, stanData = stanData, chains = TRUE, popol_plot = TRUE)
}
```

## Case study of *Larix decidua*

### Model from @Yang2024
```{r}
## Compile the model
larix = cmdstan_model("./yang2024_simplified.stan") # Simplified model, but c provided

## Common variables
n_chains = 4

savePath = "./yang/"
if (!dir.exists(savePath))
	dir.create(savePath)

sp = "Larix decidua"

print(paste(" Running species", sp))
filename = paste0(sp, "_simplified")

stanData = list(
	N = tree_dt[.(sp)][, .N],
	SB = tree_dt[.(sp)][, bole_volume_m3],
	AGB = tree_dt[.(sp)][, total_volume_m3]
)

if (!file.exists(paste0(savePath, filename, ".rds")))
{
	fit = larix$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
		iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

	## Save results
	fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
	saveRDS(fit, paste0(savePath, filename, ".rds"))
} else {
	fit = readRDS(paste0(savePath, filename, ".rds"))
}

plot_sp(fit, sp, stanData, chains = TRUE, sep_chains = TRUE, popol_plot = FALSE)

pgfTrace(fit, sp, params = "c", modelsname = "yang2024-simplified")
```

### Our model
$$
	r(x) = \alpha + \exp[-\beta*x](\gamma x + \delta)
$$
```{r}
## Compile model
larix = cmdstan_model("./larix.stan") # Different function

## Common variables
filename = paste0(sp, "_amael")

stanData = list(
	N = tree_dt[.(sp)][, .N],
	SB = tree_dt[.(sp)][, bole_volume_m3],
	AGB = tree_dt[.(sp)][, total_volume_m3]
)

if (!file.exists(paste0(savePath, filename, ".rds")))
{
	fit = larix$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
		iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

	## Save results
	fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
	saveRDS(fit, paste0(savePath, filename, ".rds"))
} else {
	fit = readRDS(paste0(savePath, filename, ".rds"))
}

## Posteriors and chains
lazyTrace(fit$draws("asym"))
lazyTrace(fit$draws("b"))
lazyTrace(fit$draws("c"))
lazyTrace(fit$draws("d"))

# Extract mean values parameters
paramsVec = getParams(model_cmdstan = fit, params_names = c("asym", "b", "c", "d"), type = "mean")

# Ratio equation
amael = function(x, pars)
	return (pars["asym"] + exp(-pars["b"]*x) * (pars["c"]*x + pars["d"]));

# Allometric equation
psi = function(x, pars)
{
	r = amael(x, pars)
	return ( 1/pars["asym"] * x^( 1 - (log(r) - log(pars["asym"])) / log(x) ) )
}

# Plot fitting data and fitted curve (ratio)
plot(stanData$SB, stanData$SB/stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
	xlab = "SB", ylab = "r", axes = FALSE, main = sp)
curve(amael(x, paramsVec), lwd = 4, add = TRUE)
axis(1)
axis(2, las = 1)

# Plot fitting data and fitted curve (above-ground volume)
plot(stanData$SB, stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
	xlab = "Bole volume", ylab = "Above-ground volume", axes = FALSE, main = sp)
curve(psi(x, paramsVec), lwd = 4, add = TRUE)
abline(a = 0, b = 1, lty = "dashed")
axis(1)
axis(2, las = 1)

# Plot obs vs pred (above-ground volume)
v_pred = apply(X = fit$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
plot(v_pred, stanData$AGB, pch = 19, cex = 0.65, col = "#A291AF33",
	xlab = "Predicted AGV", ylab = "Observed AGV", axes = FALSE, main = sp)
abline(a = 0, b = 1, lty = "dashed")
axis(1)
axis(2, las = 1)

pgfTrace(fit, sp, params = "asym", modelsname = "amael")
```

## Clayton's copula

```{r}
## Compile the model
copula = cmdstan_model("./copula_noUpdate_Fbft.stan") # Simplified model, but c provided

## Common variables
n_chains = 4

savePath = "./copula_res/"
if (!dir.exists(savePath))
	dir.create(savePath)

sp = "Abies alba"

print(paste(" Running species", sp))
filename = paste0(sp, "_copula_noUpdate")

## Load results Flo++ model (Form factor BFT)
sp_code = "61"
Fbft = readRDS(paste0("../../gohon_florence/", sp_code, "/florence-gamma_heterosked.rds"))

params = getParams(model_cmdstan = Fbft, params_names = c(paste0("beta", 0:4), "sigma0", "sigma_pow"),
	type = "mean")
rm(Fbft)

## Stan data, USING !is.na(taper_height), WAITING FOR the estimation fo taper height in Inra data
stanData = list(
	N = tree_dt[.(sp)][!is.na(taper_height), .N],
	params_Fbft = params,

	circumference_m = tree_dt[.(sp)][!is.na(taper_height), circumference_m],
	taper_height_flo = tree_dt[.(sp)][!is.na(taper_height), taper_height],
	height = tree_dt[.(sp)][!is.na(taper_height), height],

	bole_volume_m3 = tree_dt[.(sp)][!is.na(taper_height), bole_volume_m3],
	total_volume_m3 = tree_dt[.(sp)][!is.na(taper_height), total_volume_m3]
)

if (!file.exists(paste0(savePath, filename, ".rds")))
{
	fit = copula$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
		iter_warmup = 1500, iter_sampling = 1000, max_treedepth = 10)

	## Save results
	fit$save_output_files(dir = savePath, basename = filename, random = FALSE)
	saveRDS(fit, paste0(savePath, filename, ".rds"))
} else {
	fit = readRDS(paste0(savePath, filename, ".rds"))
}

lazyPosterior(fit$draws("alpha0"), fun = dnorm, mean = 0, sd = 1)
lazyPosterior(fit$draws("theta"), fun = dgamma, shape = 1.5^2, rate = 1.5)
```