---
title: "Compare the models (complete and submodel)"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(inventR)
library(interp)
library(terra)
library(shape)
library(loo)

## Tool function
source("./toolFunctions.R")

plot_sp = function(fit, sp, forest = tree_dt[.(sp)], pred = TRUE, simplif = FALSE,
	n_bins = 4, pal = "Hiroshige", selected_variable = NULL, print_plot = TRUE)
{
	# Get parameters
	n_chains = fit$num_chains()

	if (n_chains != nrow(fit$metadata()$time))
	{
		warning("Some chains are missing probably because they could not start! Using the number of successful chains")
		n_chains = nrow(fit$metadata()$time)
	}

	# Add colour to the measured points according to selected variable
	if (!is.null(selected_variable))
	{
		if (!(selected_variable %in% colnames(forest)))
			stop(paste0("'", selected_variable, "' is not a column of forest"))
		
		colours = MetBrewer::met.brewer(pal, n_bins)[1:n_bins]
		forest[, colour_ind := as.numeric(cut(.SD[[selected_variable]], breaks = n_bins))] # Indices per category
		forest[, colour := colours[colour_ind]] # Map selected variable to colour using indices
	} else {
		forest[, colour := "#3355AA33"] # Default colour
	}

	# Define fitted functions (depends if simplified model 2 stages vs 3 stages)
	if (!simplif)
	{
		paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")
	} else {
		paramsVec_simplif = getParams(model_cmdstan = fit,
			params_names = c("alpha", "beta_", "gamma", "delta"), type = "mean")
		paramsVec = c(
			c = unname(paramsVec_simplif["alpha"]),
			j = 1,
			k = unname(paramsVec_simplif["beta_"]),
			m = unname(exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"]),
			n = unname(paramsVec_simplif["delta"] + paramsVec_simplif["alpha"]),
			s = unname(paramsVec_simplif["beta_"])
		)
	}

	r_func = function(x, pars)
	{
		return ((pars["m"] - pars["c"]) * exp(pars["j"] - pars["k"]*x) * (pars["k"]*x/pars["j"])^pars["j"] +
			pars["c"] - (pars["c"] - pars["n"])*exp(-pars["s"]*x));
	}
	
	psi = function(x, pars)
	{
		r = r_func(x, pars)

		return ( 1/pars["c"] * x^( 1 - (log(r) - log(pars["c"])) / log(x) ) )
	}

	# Check rhats
	pars_names = names(paramsVec)
	if (simplif)
		pars_names = c("alpha", "beta_", "gamma", "delta")
	rhats = rhat(object = fit, pars = pars_names)

	# Divergence check up
	params = getParams(model_cmdstan = fit, params_names = pars_names, type = "all")
	div = posterior::subset_draws(fit$sampler_diagnostics(), variable = "divergent__")

	if (!print_plot)
		return (list(any_div = any(div != 0), loc = div, rhats = rhats))

	# Plot figure r vs bole and fitted curve
	plot(forest[, bole_volume_m3], forest[, r], pch = 19, col = forest[, colour], axes = FALSE,
		xlab = "Bole volume", ylab = "Ratio bole/tot volumes", lwd = 0, cex = 0.75,
		main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
	if (!is.null(selected_variable))
		legend("topright", fill = colours, title = paste(selected_variable, "classes"),
			legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
	curve(r_func(x, paramsVec), add = TRUE, lwd = 3)
	axis(1)
	axis(2, las = 1)

	# Plot predictions
	if (pred)
	{
		sim = apply(X = fit$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
		plot(sim, forest[, total_volume_m3], pch = 19, col = forest[, colour], axes = FALSE,
			xlab = "Predicted total volume", ylab = "Observed total volume",
			main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
		if (!is.null(selected_variable))
			legend("topright", fill = colours, title = paste(selected_variable, "classes"),
				legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
		abline(a = 0, b = 1, lty = "dashed", lwd = 1.5, col = "#2E2E2E")
		axis(1)
		axis(2, las = 1)
		
		plot(forest[, bole_volume_m3], forest[, total_volume_m3], pch = 19, col = forest[, colour],
			axes = FALSE, xlab = "Observed bole volume", ylab = "Observed total volume",
			main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
		if (!is.null(selected_variable))
			legend("topright", fill = colours, title = paste(selected_variable, "classes"),
				legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
		abline(a = 0, b = 1, lty = "dashed", lwd = 0.85, col = "#2E2E2E")
		curve(psi(x, paramsVec), add = TRUE, lwd = 3)
		axis(1)
		axis(2, las = 1)
	}

	# Plot chains if any rhat problem
	if (any(rhats > 1.05))
	{
		if (!simplif)
		{
			lazyTrace(fit$draws("c"), main = paste0(sp, ", c"))
			lazyTrace(fit$draws("j"), main = paste0(sp, ", j"))
			lazyTrace(fit$draws("k"), main = paste0(sp, ", k"))
			lazyTrace(fit$draws("m"), main = paste0(sp, ", m"))
			lazyTrace(fit$draws("n"), main = paste0(sp, ", n"))
			lazyTrace(fit$draws("s"), main = paste0(sp, ", s"))

			paramsVec1 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 1)
			paramsVec2 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 2)
			paramsVec3 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 3)
			paramsVec4 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 4)
		} else {
			lazyTrace(fit$draws("alpha"), main = "alpha")
			lazyTrace(fit$draws("beta_"), main = "beta_")
			lazyTrace(fit$draws("gamma"), main = "gamma")
			lazyTrace(fit$draws("delta"), main = "delta")

			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 1)
			paramsVec1 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 2)
			paramsVec2 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 3)
			paramsVec3 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 4)
			paramsVec4 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
		}
		plot(forest[, bole_volume_m3], forest[, r], pch = 19, axes = FALSE, lwd = 0,
			xlab = "Bole volume", ylab = "Ratio bole/tot volumes", main = sp)
		curve(r_func(x, paramsVec), add = TRUE, lwd = 3)
		curve(r_func(x, paramsVec1), lwd = 3, add = TRUE, col = met.brewer("Egypt")[1])
		curve(r_func(x, paramsVec2), lwd = 3, add = TRUE, col = met.brewer("Egypt")[2])
		curve(r_func(x, paramsVec3), lwd = 3, add = TRUE, col = met.brewer("Egypt")[3])
		curve(r_func(x, paramsVec4), lwd = 3, add = TRUE, col = met.brewer("Egypt")[4])
		axis(1)
		axis(2, las = 1)
	}

	return (list(any_div = any(div != 0), loc = div, rhats = rhats))
}

## Function to get NFI data for a given species
get_nfi = function(sp_code_nfi)
{
	## French NFI data query
	db = connect_db()

	# In Ancient Method (am), the inventory was done by cycles rather than homogeneously distributed everywhere
	nfi_data_am = exec_req(conn = db, req = paste0("
		SELECT
			tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.R, tree.V, tree.VTOT

		FROM
			inv_exp_am.g3arbre AS tree

		WHERE
			tree.ESS = '", sp_code_nfi, "' AND -- Selected tree species
			tree.MC IN ('4', '5') AND -- Circumference measured, and not diameter
			(tree.DOM BETWEEN '0' AND '5') -- Alive and no pollarded nor heavily pruned tree

		ORDER BY
			NPP, A;"), DT = TRUE)

	# New method
	nfi_data_nm = exec_req(conn = db, req = paste0("
		SELECT
			tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.R, tree.V, tree.V0 AS VTOT

		FROM
			inv_exp_nm.g3arbre AS tree

		WHERE
			tree.ESS = '", sp_code_nfi, "' AND -- Selected tree species
			(tree.veget = '0' OR tree.veget = 'Z') AND -- Alive
			COALESCE(tree.forme, 'X') NOT IN ('0', 'Z') AND -- No pollarded nor heavily pruned tree
			COALESCE(tree.tige, '1') = '1' -- No twin, forked, or heavily broken trees

		ORDER BY
			NPP, A;"), DT = TRUE)

	disconnect_db(db)

	# Merge data
	nfi_data_am[, method := "ancient"]
	nfi_data_nm[, method := "new"]
	
	nfi_data = rbindlist(list(nfi_data_am, nfi_data_nm))

	# Change names
	setnames(nfi_data, new = c("plot_id", "tree_id", "sp_code", "circumference_m", "height", "cull",
		"bole_volume_m3", "total_volume_m3", "method"))

	# Commpute ratio
	nfi_data[, r_vallet := bole_volume_m3/total_volume_m3]

	return(nfi_data)
}

ratio_vallet_fit = function(fit, sim, nfi_data, n_draws = NULL, simplif = FALSE)
{
	if (!simplif)
	{
		paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")
	} else {
		paramsVec_simplif = getParams(model_cmdstan = fit,
			params_names = c("alpha", "beta_", "gamma", "delta"), type = "mean")
		paramsVec = c(
			c = unname(paramsVec_simplif["alpha"]),
			j = 1,
			k = unname(paramsVec_simplif["beta_"]),
			m = unname(exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"]),
			n = unname(paramsVec_simplif["delta"] + paramsVec_simplif["alpha"]),
			s = unname(paramsVec_simplif["beta_"])
		)
	}

	r_func = function(x, pars)
	{
		return ((pars["m"] - pars["c"]) * exp(pars["j"] - pars["k"]*x) * (pars["k"]*x/pars["j"])^pars["j"] +
			pars["c"] - (pars["c"] - pars["n"])*exp(-pars["s"]*x));
	}
	
	psi = function(x, pars)
	{
		r = r_func(x, pars)

		return ( 1/pars["c"] * x^( 1 - (log(r) - log(pars["c"])) / log(x) ) )
	}

	# Compare Vallet vs my ratio fit
	# plot(x = nfi_data[, bole_volume_m3], y = nfi_data[, r_vallet], pch = 19, col = "#FAB255",
	# 	xlab = "Bole volume", ylab = "Predicted ratio", axes = FALSE)
	# curve(r_func(x, paramsVec), add = TRUE, col = "#0F7BA2", lwd = 4)
	# legend(x = "topright", legend = c("Vallet", "Me"), fill = c("#FAB255", "#0F7BA2"), box.lwd = 0)
	# axis(1)
	# axis(2, las = 1)

	# Mine vs predictor, with uncertainty
	if (!is.null(n_draws))
	{
		n_sampling = fit$metadata()$iter_sampling
		n_chains = fit$num_chains()

		n_draws = min(n_draws, n_sampling*n_chains)
		ind = sample(x = n_sampling*n_chains, size = n_draws)

		aa = sim$draws("r_gen")
		dt = data.table(bole_volume_m3 = rep(nfi_data[, bole_volume_m3], each = n_draws),
			r_me = -Inf)
		for (i in 1:dim(aa)[3])
			dt[((i - 1)*n_draws + 1):(i*n_draws), r_me := aa[, , i][ind]]
		
		rm(aa)
	}

	ylim = range(nfi_data[, r_vallet])

	if (!is.null(n_draws))
		ylim = c(min(dt[, r_me], nfi_data[, r_vallet]), max(dt[, r_me], nfi_data[, r_vallet]))
	
	plot(x = nfi_data[, bole_volume_m3], y = nfi_data[, r_vallet], pch = 19, col = "#FAB255", ylim = ylim,
		xlab = "Predicted bole volume", ylab = "Predicted ratio (me)", axes = FALSE)
	if (!is.null(n_draws))
		points(x = dt[, bole_volume_m3], y = dt[, r_me], pch = ".")
	curve(r_func(x, paramsVec), add = TRUE, col = "#0F7BA2", lwd = 4)
	axis(1)
	axis(2, las = 1)
	legend(x = "topright", legend = c("Vallet", "Me"), fill = c("#FAB255", "#0F7BA2"), box.lwd = 0)

	# Predicted volumes vs circumference
	plot(x = stanData$bole_volume_m3_new, y = stanData$total_volume_m3_new, pch = 19, col = "#FAB255",
		xlab = "Predicted total volume", ylab = "Bole volume", axes = FALSE)
	abline(a = 0, b = 1, lwd = 0.85, lty = "dashed", col = "#4A4A4A")
	points(x = stanData$bole_volume_m3_new, y = v_tot_pred)
	legend(x = "topright", legend = c("Ancient", "New"), fill = egypt[1:2], box.lwd = 0)
	axis(1)
	axis(2, las = 1)
}

## Function to compare models, based on PSIS-LOO
comparison = function(sp, tree_dt, woodstock_seed = 1969 - 08 - 18, printPlot = FALSE)
{
	#### Generate quantitities
	## Compile generator models
	genQ_old = cmdstan_model("./yang2024-genQ_old.stan")
	genQ_new = cmdstan_model("./yang2024_genQ.stan")
	genQ_sub = cmdstan_model("./yang2024-genQ_submodel.stan")

	## Prepare data
	stanData = list(
		N = tree_dt[.(sp)][, .N],
		N_new = tree_dt[.(sp)][, .N],
		bole_volume_m3 = tree_dt[.(sp)][, bole_volume_m3],
		bole_volume_m3_new = tree_dt[.(sp)][, bole_volume_m3],
		total_volume_m3 = tree_dt[.(sp)][, total_volume_m3],
		total_volume_m3_new = tree_dt[.(sp)][, total_volume_m3]
	)

	## Load fitted model
	old_file = paste0("./yang/", stri_replace(str = sp, regex = " ", replacement = "-"))
	new_file = paste0("./yang/", stri_replace(str = sp, regex = " ", replacement = "-"), "_new")
	sub_file = paste0("./yang/", stri_replace(str = sp, regex = " ", replacement = "-"), "_submodel")

	if (sp == "Quercus sp.")
		new_file = stri_replace(str = new_file, regex = "._", replacement = "_")

	old = readRDS(paste0(old_file, ".rds"))
	new = readRDS(paste0(new_file, ".rds"))
	sub = readRDS(paste0(sub_file, ".rds"))

	n_chains = old$num_chains()

	## Predict on new data
	sim_old = genQ_old$generate_quantities(fitted_params = old, data = stanData,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))

	sim_new = genQ_new$generate_quantities(fitted_params = new, data = stanData,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))
	
	sim_sub = genQ_sub$generate_quantities(fitted_params = sub, data = stanData,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))

	## Save simulations
	v_gen_mean_old = apply(X = sim_old$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
	v_gen_mean_new = apply(X = sim_new$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
	v_gen_mean_sub = apply(X = sim_sub$draws("v_gen_mean"), MARGIN = 3, FUN = mean)

	saveRDS(data.table(old = v_gen_mean_old, new = v_gen_mean_new, sub = v_gen_mean_sub),
		paste0(old_file, "_simV.rds"))

	if (printPlot)
	{
		plot(x = v_gen_mean_old, y = v_gen_mean_new, pch = 19, xlab = "Old", ylab = "New", axes = FALSE)
		abline(a = 0, b = 1, col = "#CD212A")
		axis(1)
		axis(2, las = 1)
		
		plot(x = v_gen_mean_old, y = v_gen_mean_sub, pch = 19, xlab = "Old", ylab = "Sub", axes = FALSE)
		abline(a = 0, b = 1, col = "#CD212A")
		axis(1)
		axis(2, las = 1)

		plot(x = v_gen_mean_sub, y = v_gen_mean_new, pch = 19, xlab = "Sub", ylab = "New", axes = FALSE)
		abline(a = 0, b = 1, col = "#CD212A")
		axis(1)
		axis(2, las = 1)
		
		plot(v_gen_mean_old - v_gen_mean_new, pch = 19, xlab = "Rank", ylab = "Diff")
		abline(a = 1, b = 0, lwd = 2, col = "#CD212A")
	}

	## Compute PSIS-LOO...
	# ... for old
	r_eff_old = loo::relative_eff(exp(sim_old$draws("log_lik")), cores = 8)
	loo_old = loo::loo(x = sim_old$draws("log_lik"), r_eff = r_eff_old, cores = 8)

	# ... for new
	r_eff_new = loo::relative_eff(exp(sim_new$draws("log_lik")), cores = 8)
	loo_new = loo::loo(x = sim_new$draws("log_lik"), r_eff = r_eff_new, cores = 8)
	
	# ... for submodel
	r_eff_sub = loo::relative_eff(exp(sim_sub$draws("log_lik")), cores = 8)
	loo_sub = loo::loo(x = sim_sub$draws("log_lik"), r_eff = r_eff_sub, cores = 8)

	warning_loo_old = FALSE
	if (any(loo_old$diagnostics$pareto_k >= 0.7))
		warning_loo_old = TRUE

	warning_loo_new = FALSE
	if (any(loo_new$diagnostics$pareto_k >= 0.7))
		warning_loo_new = TRUE
	
	warning_loo_sub = FALSE
	if (any(loo_sub$diagnostics$pareto_k >= 0.7))
		warning_loo_sub = TRUE

	warn = warning_loo_old | warning_loo_new | warning_loo_sub

	## Compare both models based on psisloo and weights
	comp = loo::loo_compare(list(old = loo_old, new = loo_new, sub = loo_sub))
	best = rownames(comp)[1]

	weights = head(loo::loo_model_weights(x = list(old = loo_old, new = loo_new, sub = loo_sub),
		method = "stacking")) # I use head to return a named vector

	## R squared
	var_fit_old = apply(X = posterior::as_draws_matrix(sim_old$draws("v_gen_mean")),
		MARGIN = 1, FUN = var)
	var_res_old = sim_old$draws("sigma_var")

	var_fit_new = apply(X = posterior::as_draws_matrix(sim_new$draws("v_gen_mean")),
		MARGIN = 1, FUN = var)
	var_res_new = sim_new$draws("sigma_var")
	
	var_fit_sub = apply(X = posterior::as_draws_matrix(sim_sub$draws("v_gen_mean")),
		MARGIN = 1, FUN = var)
	var_res_sub = sim_sub$draws("sigma_var")

	rsq_old = var_fit_old / (var_fit_old + var_res_old)
	rsq_new = var_fit_new / (var_fit_new + var_res_new)
	rsq_sub = var_fit_sub / (var_fit_sub + var_res_sub)

	## Return results
	return (list(
		best = best, comploo = comp, weights = weights,
		warning_loo_old = warning_loo_old, warning_loo_new = warning_loo_new, warning_loo_sub = warning_loo_sub,
		avg_old = mean(rsq_old), avg_new = mean(rsq_new), avg_sub = mean(rsq_sub),
		warning = warn))
}

## Function to compare model new version with submodel, based on PSIS-LOO
comparison_yang_sub = function(sp, tree_dt, woodstock_seed = 1969 - 08 - 18, printPlot = FALSE)
{
	#### Generate quantitities
	## Compile generator models
	genQ_new = cmdstan_model("./yang2024_genQ.stan")
	genQ_sub = cmdstan_model("./yang2024-genQ_submodel.stan")

	## Prepare data
	stanData = list(
		N = tree_dt[.(sp)][, .N],
		N_new = tree_dt[.(sp)][, .N],
		bole_volume_m3 = tree_dt[.(sp)][, bole_volume_m3],
		bole_volume_m3_new = tree_dt[.(sp)][, bole_volume_m3],
		total_volume_m3 = tree_dt[.(sp)][, total_volume_m3],
		total_volume_m3_new = tree_dt[.(sp)][, total_volume_m3]
	)

	## Load fitted model
	new_file = paste0("./yang/", stri_replace(str = sp, regex = " ", replacement = "-"), "_new")
	sub_file = paste0("./yang/", stri_replace(str = sp, regex = " ", replacement = "-"), "_submodel")

	if (sp == "Quercus sp.")
		new_file = stri_replace(str = new_file, regex = "._", replacement = "_")

	new = readRDS(paste0(new_file, ".rds"))
	sub = readRDS(paste0(sub_file, ".rds"))

	n_chains = new$num_chains()

	## Compute max rhat for Yang and submodel
	rhat_yang = posterior::summarise_draws(new$draws(), posterior::rhat)
	rhat_yang_m = max(rhat_yang[, "posterior::rhat"])
	rhat_sub = posterior::summarise_draws(sub$draws(), posterior::rhat)
	rhat_sub_m = max(rhat_sub[, "posterior::rhat"])

	## Predict on new data
	sim_new = genQ_new$generate_quantities(fitted_params = new, data = stanData,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))
	
	sim_sub = genQ_sub$generate_quantities(fitted_params = sub, data = stanData,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))

	## Save simulations
	v_gen_mean_new = apply(X = sim_new$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
	v_gen_mean_sub = apply(X = sim_sub$draws("v_gen_mean"), MARGIN = 3, FUN = mean)

	if (printPlot)
	{	
		plot(x = v_gen_mean_sub, y = v_gen_mean_new, pch = 19, xlab = "Sub", ylab = "New", axes = FALSE)
		abline(a = 0, b = 1, col = "#CD212A")
		axis(1)
		axis(2, las = 1)
	}

	## Compute PSIS-LOO...
	# ... for new
	r_eff_new = loo::relative_eff(exp(sim_new$draws("log_lik")), cores = 8)
	loo_new = loo::loo(x = sim_new$draws("log_lik"), r_eff = r_eff_new, cores = 8)
	
	# ... for submodel
	r_eff_sub = loo::relative_eff(exp(sim_sub$draws("log_lik")), cores = 8)
	loo_sub = loo::loo(x = sim_sub$draws("log_lik"), r_eff = r_eff_sub, cores = 8)

	warning_loo_new = FALSE
	if (any(loo_new$diagnostics$pareto_k >= 0.7))
		warning_loo_new = TRUE
	
	warning_loo_sub = FALSE
	if (any(loo_sub$diagnostics$pareto_k >= 0.7))
		warning_loo_sub = TRUE

	warn = warning_loo_new | warning_loo_sub

	## Compare both models based on psisloo and weights
	comp = loo::loo_compare(list(yang = loo_new, sub = loo_sub))
	best = rownames(comp)[1]

	weights = head(loo::loo_model_weights(x = list(yang = loo_new, sub = loo_sub),
		method = "stacking")) # I use head to return a named vector

	## R squared
	var_fit_new = apply(X = posterior::as_draws_matrix(sim_new$draws("v_gen_mean")),
		MARGIN = 1, FUN = var)
	var_res_new = sim_new$draws("sigma_var")
	
	var_fit_sub = apply(X = posterior::as_draws_matrix(sim_sub$draws("v_gen_mean")),
		MARGIN = 1, FUN = var)
	var_res_sub = sim_sub$draws("sigma_var")

	rsq_new = var_fit_new / (var_fit_new + var_res_new)
	rsq_sub = var_fit_sub / (var_fit_sub + var_res_sub)

	## Return results
	return (list(
		best = best, comploo = comp, weights = weights,
		warning_loo_yang = warning_loo_new, warning_loo_sub = warning_loo_sub, warning = warn,
		avg_yang = mean(rsq_new), avg_sub = mean(rsq_sub),
		rhat_yang = rhat_yang_m, rhat_sub = rhat_sub_m))
}

rhat_temp = function(sp, tree_dt)
{
	## Load fitted model
	new_file = paste0("./yang/", stri_replace(str = sp, regex = " ", replacement = "-"), "_new")
	sub_file = paste0("./yang/", stri_replace(str = sp, regex = " ", replacement = "-"), "_submodel")

	if (sp == "Quercus sp.")
		new_file = stri_replace(str = new_file, regex = "._", replacement = "_")

	new = readRDS(paste0(new_file, ".rds"))
	sub = readRDS(paste0(sub_file, ".rds"))

	n_chains = new$num_chains()

	## Compute max rhat for Yang and submodel
	rhat_yang = posterior::summarise_draws(new$draws(), posterior::rhat)
	rhat_yang_m = max(rhat_yang[, "posterior::rhat"])
	rhat_sub = posterior::summarise_draws(sub$draws(), posterior::rhat)
	rhat_sub_m = max(rhat_sub[, "posterior::rhat"])

	## Return
	return(list(rhat_yang = rhat_yang_m, rhat_sub = rhat_sub_m))
}

rebuilt_comp = function(save_ls)
{
	ls_species = names(save_ls)
	
	comp = data.table(species = ls_species, best = "", elpd_diff_old = -Inf, se_diff_old = -Inf,
		elpd_diff_new = -Inf, se_diff_new = -Inf,
		elpd_diff_sub = -Inf, se_diff_sub = -Inf,
		warning = FALSE, key = "species")
	weights_dt = data.table(species = ls_species, best = "", old = -Inf, new = -Inf, sub = -Inf, key = "species")
	R2D2 = data.table(species = ls_species, R2_old = -Inf, R2_new = -Inf, R2_sub = -Inf, key = "species")

	for (sp in ls_species)
	{
		weights_dt[sp, c("old", "new", "sub") := as.list(round(save_ls[[sp]]$weights, digits = 3))]
		R2D2[sp, c("R2_old", "R2_new", "R2_sub") :=
			.(round(save_ls[[sp]]$avg_old, digits = 3),
				round(save_ls[[sp]]$avg_new, digits = 3),
				round(save_ls[[sp]]$avg_sub, digits = 3))]
		
		comp[sp, c("best", "warning") := .(save_ls[[sp]]$best, save_ls[[sp]]$warning)]
		
		comp[sp, c("elpd_diff_old", "se_diff_old") :=
				as.list(round(save_ls[[sp]]$comploo["old", c("elpd_diff", "se_diff")], digits = 2))]
		comp[sp, c("elpd_diff_new", "se_diff_new") :=
				as.list(round(save_ls[[sp]]$comploo["new", c("elpd_diff", "se_diff")], digits = 2))]
		comp[sp, c("elpd_diff_sub", "se_diff_sub") :=
				as.list(round(save_ls[[sp]]$comploo["sub", c("elpd_diff", "se_diff")], digits = 2))]
	}

	return(list(comp = comp, weights_dt = weights_dt, R2D2 = R2D2))
}
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

tree_dt = readRDS(paste0(path_data, "data_Vtot.rds"))

sp_list = readRDS(paste0(path_data, "species_dt.rds"))
sp_list[, species_code_inra := NULL]
sp_list = na.omit(sp_list)
setkey(sp_list, speciesName_sci)
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
## Comparison Old/new Yang 2024 and my simplified version

I will do two comparison:
1. Old vs new Yang 2024 vs sub, in order to check if new and old are the same for species that are working with old version, and to see if new is able to replace sub in some cases.
2. New Yang 2024 vs sub, once I validated that Old and New mostly agree and that performance are the same.

### Old vs new Yang 2024 vs sub

```{r}
ls_species = tree_dt[, unique(speciesName_sci)]
comp = data.table(species = ls_species, best = "", elpd_diff = -Inf, se_diff = -Inf,
	warning = FALSE, key = "species")

weights_dt = data.table(species = ls_species, best = "", old = -Inf, new = -Inf, sub = -Inf, key = "species")

R2D2 = data.table(species = ls_species, R2_old = -Inf, R2_new = -Inf, R2_sub = -Inf, key = "species")

if (file.exists("./yang/comparison.rds"))
{
	save_ls = readRDS("./yang/comparison.rds")
} else {
	save_ls = vector(mode = "list", length = length(ls_species))
	names(save_ls) = ls_species

	for (sp in ls_species)
	{
		temp = comparison(sp = sp, tree_dt = tree_dt)
		save_ls[[sp]] = temp
		comp[.(sp), c("best", "elpd_diff", "se_diff", "warning") :=
			.(temp$best, round(temp$comploo[2, 1], 2), round(temp$comploo[2, 2], 2), temp$warning)]

		weights_dt[.(sp), c("best", "old", "new", "sub") :=
			.(temp$best, temp$weights["old"], temp$weights["new"], temp$weights["sub"])]

		R2D2[.(sp), c("R2_old", "R2_new", "R2_sub") := .(temp[["avg_old"]], temp[["avg_new"]], temp[["avg_sub"]])]
	}

	saveRDS(save_ls, "./yang/comparison.rds")
}
```

*In fine*, it seems that old and new are quite similare, with non-significative differences. Therefore, I can now process only with the new version, which I will use in the article (and the old one not being reported, as it does not follow eactly the method section, with the hidden dependance of $s$ to $$j$ and $k$).

### New vs sub

```{r}
ls_species = tree_dt[, unique(speciesName_sci)]
comp = data.table(species = ls_species, best = "", elpd_diff = -Inf, se_diff = -Inf,
	warning = FALSE, key = "species")

weights_dt = data.table(species = ls_species, best = "", W_yang = -Inf, W_sub = -Inf, key = "species")

rhat_dt = data.table(species = ls_species, Rhat_yang = -Inf, Rhat_sub = -Inf, key = "species")

R2D2 = data.table(species = ls_species, R2_yang = -Inf, R2_sub = -Inf, key = "species")

if (file.exists("./yang/comparison_new-sub.rds"))
{
	save_ls = readRDS("./yang/comparison_new-sub.rds")
} else {
	save_ls = vector(mode = "list", length = length(ls_species))
	names(save_ls) = ls_species

	for (sp in ls_species)
	{
		temp = comparison_yang_sub(sp = sp, tree_dt = tree_dt)
		save_ls[[sp]] = temp
		comp[.(sp), c("best", "elpd_diff", "se_diff", "warning") :=
			.(temp$best, temp$comploo[2, 1], temp$comploo[2, 2], temp$warning)]

		weights_dt[.(sp), c("best", "W_yang", "W_sub") :=
			.(temp$best, temp$weights["yang"], temp$weights["sub"])]

		R2D2[.(sp), c("R2_yang", "R2_sub") := .(temp[["avg_yang"]], temp[["avg_sub"]])]

		rhat_dt[sp, c("Rhat_yang", "Rhat_sub") := .(temp[["rhat_yang"]], temp[["rhat_sub"]])]
	}

	comp = merge.data.table(comp, rhat_dt, by = "species")
	comp = merge.data.table(comp, tree_dt[, .N, by = speciesName_sci], by.x = "species", by.y = "speciesName_sci")
	saveRDS(save_ls, "./yang/comparison_new-sub.rds")
	fwrite(R2D2, "../../../writing/article_allometry/data/rsquared.csv")
}
```

## Reading @McLatchie2024

I am testing the appendix A of @McLatchie2024, and more specifically the pseudo-BMA+ weights, equation A4. My coding of equation A2 might be surprising, but read well equation A2! It is:
$$
	P[A > B],
$$

which equals

$$
	1 - P[A < B],
$$

and since $\int \text{dnorm}$ returns $P[A < B]$ I need the $1 - $ stuff! Not that it is also equivalent to integrate the function Gauss from 0 to $\infty$...

Finally, note that I use `pnorm` as I noticed the `integrate` function is unstable for large values of $\Delta \text{ELPD}$...

```{r}
diff_elpd = 4
se_elpd = 2
gauss = function(x, diff_elpd, se_elpd)
	return(dnorm(x, mean = diff_elpd, sd = se_elpd))

f = function(x, diff_elpd, se_elpd)
	return(dnorm(x, mean = 0, sd = se_elpd)/(1 + exp(-diff_elpd - x)))

1 - integrate(gauss, lower = -Inf, upper = 0, diff_elpd = diff_elpd, se_elpd = se_elpd)$value # Eq A2

pnorm(0, mean = diff_elpd, sd = se_elpd, lower.tail = FALSE) # Eq A2, equivalent

integrate(f, lower = -Inf, upper = +Inf, diff_elpd = diff_elpd, se_elpd = se_elpd)$value # Eq A4


comp[, eq_A2 := -Inf]
comp[, eq_A2_bis := -Inf]
comp[, pseudo_bma := -Inf]

comp[, eq_A2 := integrate(gauss, lower = 0, upper = +Inf, diff_elpd = abs(elpd_diff), se_elpd = se_diff)$value,
	by = species]
comp[, eq_A2_bis := pnorm(0, mean = abs(elpd_diff), sd = se_diff, lower.tail = FALSE),
	by = species]
comp[, pseudo_bma := integrate(f, lower = -Inf, upper = +Inf, diff_elpd = abs(elpd_diff), se_elpd = se_diff)$value,
	by = species]

fwrite(comp, "../../../writing/article_allometry/data/comparison.csv")
```

## Other stuff, to compare Vallet with my model

```{r}
#### Generate quantitities
## Compile generator model
genQ = cmdstan_model("./yang2024_genQ.stan")

## Prepare data
sp = "Fagus sylvatica"
sp_code_nfi = sp_list[sp, species_code_nfi]

nfi_data = get_nfi(sp_code_nfi)
set.seed(1969 - 08 - 18) # woodstock seed
nfi_data = nfi_data[method == "new"][sample(x = seq_len(.N), size = 1e4, replace = FALSE)]

stanData = list(
	N = tree_dt[.(sp)][, .N],
	N_new = nfi_data[, .N],
	bole_volume_m3 = tree_dt[.(sp)][, bole_volume_m3],
	bole_volume_m3_new = nfi_data[, bole_volume_m3],
	total_volume_m3 = tree_dt[.(sp)][, total_volume_m3],
	total_volume_m3_new = nfi_data[, total_volume_m3]
)

## Load fitted model
sp_file = stri_replace(str = sp, regex = " ", replacement = "-")
filename = paste0("./yang/", sp_file)
fit = readRDS(paste0(filename, ".rds"))

## Common variables
n_chains = fit$num_chains()
woodstock_seed = 1969 - 08 - 18
egypt = met.brewer("Egypt")

nfi_data[, .N, by = method]
nfi_data[, colour := ifelse(method == "ancient", egypt[1], egypt[2])]

## Predict on new data
sim = genQ$generate_quantities(fitted_params = fit, data = stanData,
	seed = woodstock_seed, parallel_chains = min(n_chains, 4))

v_tot_pred = apply(X = sim$draws("v_gen_mean"), MARGIN = 3, FUN = mean)

## Plot
# Mine vs Vallet
plot(x = stanData$total_volume_m3_new, y = v_tot_pred, pch = 19, col = nfi_data[, colour],
	xlab = "Predicted total volume (Vallet)", ylab = "Predicted total volume (me)", axes = FALSE)
abline(a = 0, b = 1, lwd = 0.85, lty = "dashed", col = "#4A4A4A")
legend(x = "topright", legend = c("Ancient", "New"), fill = egypt[1:2], box.lwd = 0)
axis(1)
axis(2, las = 1)

# Ratio vallet and fit vs predictor
ratio_vallet_fit(fit, sim, nfi_data, n_draws = 10)

# Predicted volumes vs circumference
plot(x = nfi_data[, circumference_m], y = nfi_data[, total_volume_m3], pch = 19, col = "#FAB255",
	xlab = "Circumference", ylab = "Total volume", axes = FALSE)
points(x = nfi_data[, circumference_m], y = v_tot_pred, col = "#0F7BA2", pch = 19, cex = 0.65)
legend(x = "topleft", legend = c("M1", "M2"), fill = c("#FAB255", "#0F7BA2"), box.lwd = 0)
axis(1)
axis(2, las = 1)
```
