---
title: "Generate figures and pgfplots data from model outputs"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(inventR)
library(interp)
library(loo)

## Tool function
source("./toolFunctions.R")

## Function that computes PSIS-LOO, residuals, predictions, R squared on validation dataset
pred_fct = function(sp, florence_dt, nfi_data, woodstock_seed = 1969 - 08 - 18)
{	
	#### Generate quantities...
	## Prepare data
	stanData_gen = list(
		N = florence_dt[sp][(type == "ajust"), .N],
		
		circumference_m = florence_dt[sp][(type == "ajust"), c13],
		taper_height_flo = florence_dt[sp][(type == "ajust"), hdecro],
		height = florence_dt[sp][(type == "ajust"), htot],

		N_new = nfi_data[sp][, .N],

		circumference_m_new = nfi_data[sp][, circumference_m],
		taper_height_flo_new = nfi_data[sp][, taper_height_flo],
		height_new = nfi_data[sp][, height],
		bole_volume_m3_new = nfi_data[sp][, bole_volume_m3]
	)

	## ... for Florence++
	heterosked_genQ = cmdstan_model("../../gohon_florence/florence-genQ_gamma.stan")
	fit_flopp = readRDS(paste0("../../gohon_florence/", sp, "/florence-gamma_heterosked.rds"))
	n_chains = fit_flopp$num_chains()

	sim_flopp = heterosked_genQ$generate_quantities(fit_flopp, data = stanData_gen,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))

	rm(fit_flopp)

	## Predictions and residuals
	pred_flopp = apply(X = sim_flopp$draws("sim_bole"), MARGIN = 3, FUN = mean)
	pred_flopp_quant = apply(X = sim_flopp$draws("sim_bole"), MARGIN = 3, FUN = quantile,
		probs = c(0.1, 0.9))

	return (list(pred = pred_flopp, quant = pred_flopp_quant))
}
```

```{r}
#| output: false

#### Load the training data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading training dataset
training_dt = readRDS(paste0(path_data, "data_Vtot.rds"))

## List parametrised species and get their NFI code
ls_species = training_dt[, unique(speciesName_sci)]

nfi_codes = readRDS(paste0(path_data, "species_dt.rds"))[!is.na(species_code_nfi),
	.(speciesName_sci, species_code_nfi)]
setkey(nfi_codes, speciesName_sci)

nfi_codes = nfi_codes[ls_species]
setkey(nfi_codes, speciesName_sci)

nfi_codes["Fraxinus excelsior", species_code_nfi := 17] # Manually assigned Fraxinus sp. code

sp_code_nfi = paste0("'", paste(nfi_codes[, species_code_nfi], collapse = "','"), "'")

#### Loading NFI data
## French NFI data query
db = connect_db()

## In Ancient Method (am), the inventory was done by cycles rather than homogeneously distributed everywhere
nfi_data_am = exec_req(conn = db, req = paste0("
	SELECT
		tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.HDEC, tree.DDEC, tree.V

	FROM
		inv_exp_am.g3arbre AS tree

	WHERE
		tree.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		tree.MC IN ('4', '5') AND -- Circumference measured, and not diameter
		tree.CUBE = 'M' AND -- Cubed trees, remove 439 trees that have hdec but were not cubed, so ok
		(tree.DOM BETWEEN '0' AND '5') -- Alive and no pollarded nor heavily pruned tree

	ORDER BY
		NPP, A;"), DT = TRUE)

## New method
nfi_data_nm = exec_req(conn = db, req = paste0("
	SELECT
		tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.HDEC, tree.DDEC, tree.V, tree.DECOUPE

	FROM
		inv_exp_nm.g3arbre AS tree

	WHERE
		tree.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		(tree.veget = '0' OR tree.veget = 'Z') AND -- Alive
		COALESCE(tree.forme, 'X') NOT IN ('0', 'Z') AND -- No pollarded nor heavily pruned tree
		COALESCE(tree.tige, '1') = '1' -- No twin, forked, or heavily broken trees

	UNION ALL

	SELECT
		treep.NPP, treep.A, treep.ESS, treep.C13, treep.HTOT, treep.HDEC, treep.DDEC, treep.V, treep.DECOUPE

	FROM
		inv_exp_nm.p3arbre AS treep

	WHERE
		treep.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		(treep.veget = '0' OR treep.veget = 'Z') AND -- Alive
		COALESCE(treep.forme, 'X') NOT IN ('0', 'Z') AND -- No pollarded nor heavily pruned tree
		COALESCE(treep.tige, '1') = '1' -- No twin, forked, or heavily broken trees

	ORDER BY
		NPP, A;"), DT = TRUE)

disconnect_db(db)

## Remove NAs and compute hdecro (hdec from Florence, never knew why it was called hdecro...)
nfi_data_am = nfi_data_am[!is.na(ddec)]
nfi_data_am[, hdecro := hdec]
nfi_data_am[ddec == 0.07, hdecro := htot]

nfi_data_nm = nfi_data_nm[!is.na(hdec)]
nfi_data_nm[, hdecro := hdec]
nfi_data_nm[decoupe == "0", hdecro := htot] # Regulare trees

## Merge data
# Add origin
nfi_data_am[, method := "ancient"]
nfi_data_nm[, method := "new"]

# Remove non-common columns
nfi_data_am[, ddec := NULL]
nfi_data_nm[, c("decoupe", "ddec") := NULL]

# Merging
nfi_data = rbindlist(list(nfi_data_am, nfi_data_nm))

## Change names
setnames(nfi_data, new = c("plot_id", "tree_id", "sp_code", "circumference_m", "height",
	"taper_height", "bole_volume_m3", "taper_height_flo", "method"))

## Add scientific names
nfi_data = merge.data.table(nfi_data, nfi_codes, by.x = "sp_code", by.y = "species_code_nfi")

## Remove null volumes as judged irrealistic
nfi_data = nfi_data[bole_volume_m3 > 0]
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script is generates data to plot figures in pgfplots in the article. The figures that I have currently in mind are:
1. $\Vtot$ vs $\Vbft$ with 90% credibility interval (CI). Ideally, there would be two colours, one for the training range, and one for the NFI range if larger than training range (which I expect). Note that I will use the new formula for bole volumes rather than the values provided in the database...

## Figure `Vtot` vs `Vbole` with CI

I noticed that the range of values provided by the NFI are sometimes incoherent, such as 40 cubic meters for *Pinus laricio*, which is far beyond its usual range I think (although it is a tree 42.4 m tall and 5.52 m circumference)...

```{r}
range_dt = merge.data.table(
	training_dt[, .(min_t = min(bole_volume_m3), max_t = max(bole_volume_m3)), by = speciesName_sci],
	nfi_data[, .(min_nfi = min(bole_volume_m3), max_nfi = max(bole_volume_m3)), by = speciesName_sci])
```

### Predict the bole volumes based on Flopp model

This is done in ?? steps:
1. Load the training dataset of Florence
2. 

```{r}
#### Reading data
florence_dt = readRDS("../../gohon_florence/arbComp.rds")
setkey(florence_dt, "ess")

florence_dt = florence_dt[nfi_codes[, species_code_nfi]]
setkey(florence_dt, "ess")

for (sp in nfi_codes[, species_code_nfi])
	pred_fct(sp, florence_dt, nfi_data)
```
