---
title: "Run the ratio models (complete and submodel)"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(inventR)
library(interp)
library(loo)

## Tool function
source("./toolFunctions.R")
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

opt = "_inra"

# Loading Emerge data
inra = readRDS(paste0(path_data, "nfi-inra_dt", opt, ".rds"))
inra = inra[, .(speciesName_sci, tree_id = unique_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = taper_height_flo, bole_volume_m3, total_volume_m3)]
setkey(inra, speciesName_sci)

# Loading Swiss data
swiss = readRDS(paste0(mnt_point, "data/switzerland.rds"))
swiss = unique(swiss[, .(speciesName_sci, tree_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = hdec, bole_volume_m3, total_volume_m3)])
setkey(swiss, speciesName_sci)

## Modern Emerge data
emerge = readRDS(paste0(mnt_point, "data/emerge_2009-2010.rds"))
emerge[, year := as.integer(stringi::stri_replace_all(str = dataset, replacement = "", regex = "emerge_"))]
emerge = emerge[, .(speciesName_sci, tree_id, fct_type, year, circumference_m, height, taper_height,
	bole_volume_m3 = bole_volume_conic_m3, total_volume_m3)]

## Bind everything
tree_dt = rbindlist(list(inra = inra, swiss = swiss, emerge = emerge), idcol = "origin", fill = TRUE)

# Add colours France vs Swiss
ls_colours = met.brewer("Egypt")[1:4]
names(ls_colours) = c("fr_comp", "fr", "ch_comp", "ch")
tree_dt[, colour := ls_colours["fr"]] # Inra and Emerge, in blue
tree_dt[origin == "swiss", colour := ls_colours["ch"]] # Swiss, in yellow

# Remove NAs
tree_dt = tree_dt[!is.na(bole_volume_m3)]

# Remove species with less than 150 individuals
tree_dt[, n_indiv := .N, by = speciesName_sci]
tree_dt = tree_dt[n_indiv > 150]

tree_dt[, .N, by = .(speciesName_sci, colour)][order(speciesName_sci)]

# Set key to sort by species and colour (i.e., France vs Switzerland)
setkey(tree_dt, speciesName_sci, colour)
```


<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script is generates data to plot figures in pgfplots in the article. The figures that I have currently in mind are:
1. `Vtot` vs `Vbole` with 90% credibility interval (CI). Ideally, there would be two colours, one for the training range, and one for the NFI range if larger than training range (which I expect)

## Figure `Vtot` vs `Vbole` with CI

```{r}

```