---
title: "Generate figures and pgfplots data from model outputs"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(inventR)
library(interp)
library(loo)

## Tool function
source("./toolFunctions.R")
```

```{r}
#| output: false

#### Load the training data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading training dataset
training_dt = readRDS(paste0(path_data, "data_Vtot.rds"))

## List parametrised species and get their NFI code
ls_species = training_dt[, unique(speciesName_sci)]

nfi_codes = readRDS(paste0(path_data, "species_dt.rds"))[!is.na(species_code_nfi),
	.(speciesName_sci, species_code_nfi)]
setkey(nfi_codes, speciesName_sci)

nfi_codes = nfi_codes[ls_species]
setkey(nfi_codes, speciesName_sci)

nfi_codes["Fraxinus excelsior", species_code_nfi := 17] # Manually assigned Fraxinus sp. code

sp_code_nfi = paste0("'", paste(nfi_codes[, species_code_nfi], collapse = "','"), "'")

#### Loading NFI data
## French NFI data query
db = connect_db()

# In Ancient Method (am), the inventory was done by cycles rather than homogeneously distributed everywhere
nfi_data_am = exec_req(conn = db, req = paste0("
	SELECT
		tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.R, tree.V, tree.VTOT

	FROM
		inv_exp_am.g3arbre AS tree

	WHERE
		tree.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		tree.MC IN ('4', '5') AND -- Circumference measured, and not diameter
		(tree.DOM BETWEEN '0' AND '5') -- Alive and no pollarded nor heavily pruned tree

	ORDER BY
		NPP, A;"), DT = TRUE)

# New method
nfi_data_nm = exec_req(conn = db, req = paste0("
	SELECT
		tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.R, tree.V, tree.V0 AS VTOT

	FROM
		inv_exp_nm.g3arbre AS tree

	WHERE
		tree.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		(tree.veget = '0' OR tree.veget = 'Z') AND -- Alive
		COALESCE(tree.forme, 'X') NOT IN ('0', 'Z') AND -- No pollarded nor heavily pruned tree
		COALESCE(tree.tige, '1') = '1' -- No twin, forked, or heavily broken trees

	ORDER BY
		NPP, A;"), DT = TRUE)

disconnect_db(db)

# Merge data
nfi_data_am[, method := "ancient"]
nfi_data_nm[, method := "new"]

nfi_data = rbindlist(list(nfi_data_am, nfi_data_nm))

# Change names
setnames(nfi_data, new = c("plot_id", "tree_id", "sp_code", "circumference_m", "height", "cull",
	"bole_volume_m3", "total_volume_m3", "method"))

# Commpute ratio
nfi_data[, r_vallet := bole_volume_m3/total_volume_m3]
```


<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script is generates data to plot figures in pgfplots in the article. The figures that I have currently in mind are:
1. $\Vtot$ vs $\Vbft$ with 90% credibility interval (CI). Ideally, there would be two colours, one for the training range, and one for the NFI range if larger than training range (which I expect)

## Figure `Vtot` vs `Vbole` with CI

```{r}

```