---
title: "Generate figures and pgfplots data from model outputs"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(bayesplot)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(inventR)
library(interp)
library(loo)

## Tool function
source("./toolFunctions.R")

## Function to predict bole volume, using Flo++ model 
pred_fct_flopp = function(sp, stanData_gen, woodstock_seed = 1969 - 08 - 18)
{
	## Generate quantities for Florence++
	heterosked_genQ = cmdstan_model("../../gohon_florence/florence-genQ_gamma.stan")
	fit_flopp = readRDS(paste0("../../gohon_florence/", sp, "/florence-gamma_heterosked.rds"))
	n_chains = fit_flopp$num_chains()

	sim_flopp = heterosked_genQ$generate_quantities(fit_flopp, data = stanData_gen,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))

	rm(fit_flopp)

	## Predictions and quantiles (includes the residual variance for pred_flopp_quant_uncertainty)
	pred_flopp = apply(X = sim_flopp$draws("sim_bole"), MARGIN = 3, FUN = mean)
	pred_flopp_quant_uncertainty = quantile(sim_flopp$draws("sim_bole"), probs = c(0, 0.005, 0.025, 0.975, 0.995, 1))
	pred_flopp_quant = quantile(pred_flopp, probs = c(0, 0.005, 0.025, 0.975, 0.995, 1))

	return (list(pred = pred_flopp, quant_mean_pred = pred_flopp_quant,
		quant_uncertain_pred = pred_flopp_quant_uncertainty))
}

## Function to predict total above-ground volume, using adapted Yang model
pred_fct_yang = function(sp, stanData_gen, simplif = FALSE, n_points = 5, woodstock_seed = 1969 - 08 - 18)
{	
	## Generate quantities for Yang 2024
	genQ = cmdstan_model("./yang2024-genQ.stan")
	if (simplif)
		genQ = cmdstan_model("./yang2024-genQ_submodel.stan")
	
	fitname = paste0("./yang/", sp, "_new.rds")
	if (simplif)
		fitname = paste0("./yang/", sp, "_submodel.rds")
	
	fit = readRDS(fitname)
	n_chains = fit$num_chains()
	n_iter = fit$metadata()$iter_sampling

	sim = genQ$generate_quantities(fit, data = stanData_gen,
		seed = woodstock_seed, parallel_chains = min(n_chains, 4))

	if (!simplif)
	{
		paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")
	} else {
		paramsVec_simplif = getParams(model_cmdstan = fit,
			params_names = c("alpha", "beta_", "gamma", "delta"), type = "mean")
		paramsVec = c(
			c = unname(paramsVec_simplif["alpha"]),
			j = 1,
			k = unname(paramsVec_simplif["beta_"]),
			m = unname(exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"]),
			n = unname(paramsVec_simplif["delta"] + paramsVec_simplif["alpha"]),
			s = unname(paramsVec_simplif["beta_"])
		)
	}

	rm(fit)

	## Predictions on volume
	pred = apply(X = sim$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
	pred_quant_mean = apply(X = sim$draws("v_gen_mean"), MARGIN = 3, FUN = quantile, probs = c(0.05, 0.95))
	pred_points = sim$draws("v_gen")
	pred_quant = data.table(Vbole = stanData_gen$bole_volume_m3_new,
		q05 = rep(-Inf, stanData_gen$N_new), q95 = rep(-Inf, stanData_gen$N_new))

	## Predictions on ratio
	pred_r = apply(X = sim$draws("r_gen_mean"), MARGIN = 3, FUN = mean)
	pred_r_quant_mean = apply(X = sim$draws("r_gen_mean"), MARGIN = 3, FUN = quantile, probs = c(0.05, 0.95))
	pred_r_points = sim$draws("r_gen")
	pred_r_quant = data.table(Vbole = stanData_gen$bole_volume_m3_new,
		q05 = rep(-Inf, stanData_gen$N_new), q95 = rep(-Inf, stanData_gen$N_new))

	# pred_quant = apply(X = sim$draws("v_gen"), MARGIN = 3, FUN = quantile,
	# 	probs = c(0, 0.005, 0.025, 0.975, 0.995, 1))
	
	dt_v = matrix(data = -Inf, nrow = stanData_gen$N_new, ncol = n_points)
	dt_r = matrix(data = -Inf, nrow = stanData_gen$N_new, ncol = n_points)
	set.seed(woodstock_seed)
	for (i in seq_len(stanData_gen$N_new))
	{
		rand_ind = sample(x = 1:(n_iter*n_chains), size = n_points, replace = FALSE)
		dt_v[i,] = pred_points[, , i][rand_ind]
		dt_r[i,] = pred_points[, , i][rand_ind]
		pred_quant[i, c("q05", "q95") := as.list(quantile(pred_points[, , i], prob = c(0.05, 0.95)))]
		pred_r_quant[i, c("q05", "q95") := as.list(quantile(pred_r_points[, , i], prob = c(0.05, 0.95)))]
	}
		
	return (list(pred = pred, quant_mean = pred_quant_mean, quant = pred_quant,
		pred_r = pred_r, quant_mean_r = pred_r_quant_mean, quant_r = pred_r_quant,
		paramsVec = paramsVec, randomSample_v = dt_v, randomSample_r = dt_r, n_points = n_points))
}

## Function that compute the average above-ground pred volume without params/residual uncertainty
yang2024 = function(x, pars)
{
	return((pars["m"] - pars["c"]) * exp(pars["j"] - pars["k"]*x) * (pars["k"]*x/pars["j"])^pars["j"] +
		pars["c"] - (pars["c"] - pars["n"])*exp(-pars["s"]*x));
}

yang2024_vol = function(x, pars)
{
	if (is.data.table(pars))
	{
		if (pars[, .N] != 1)
			stop("pars shoud correspond to one species only")
		pars = c(c = pars[, c], j = pars[, j], k = pars[, k], m = pars[, m], n = pars[, n], s = pars[, s])
	}

	return(x/yang2024(x, pars));
}
```

```{r}
#| output: false

#### Load the training data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading training dataset
training_dt = readRDS(paste0(path_data, "data_Vtot.rds"))

## List parametrised species and get their NFI code
ls_species = training_dt[, unique(speciesName_sci)]

nfi_codes = readRDS(paste0(path_data, "species_dt.rds"))[!is.na(species_code_nfi),
	.(speciesName_sci, species_code_nfi)]
setkey(nfi_codes, speciesName_sci)

nfi_codes = nfi_codes[ls_species]
setkey(nfi_codes, speciesName_sci)

nfi_codes["Fraxinus excelsior", species_code_nfi := 17] # Manually assigned Fraxinus sp. code

sp_code_nfi = paste0("'", paste(nfi_codes[, species_code_nfi], collapse = "','"), "'")
```

```{r}
#### Loading NFI data
## French NFI data query
db = connect_db()

## In Ancient Method (am), the inventory was done by cycles rather than homogeneously distributed everywhere
nfi_data_am = exec_req(conn = db, req = paste0("
	SELECT
		tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.HDEC, tree.DDEC, tree.V

	FROM
		inv_exp_am.g3arbre AS tree

	WHERE
		tree.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		tree.MC IN ('4', '5') AND -- Circumference measured, and not diameter
		tree.CUBE = 'M' AND -- Cubed trees, remove 439 trees that have hdec but were not cubed, so ok
		(tree.DOM BETWEEN '0' AND '5') -- Alive and no pollarded nor heavily pruned tree

	ORDER BY
		NPP, A;"), DT = TRUE)

## New method
nfi_data_nm = exec_req(conn = db, req = paste0("
	SELECT
		tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.HDEC, tree.DDEC, tree.V, tree.DECOUPE

	FROM
		inv_exp_nm.g3arbre AS tree

	WHERE
		tree.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		(tree.veget = '0' OR tree.veget = 'Z') AND -- Alive
		COALESCE(tree.forme, 'X') NOT IN ('0', 'Z') AND -- No pollarded nor heavily pruned tree
		COALESCE(tree.tige, '1') = '1' -- No twin, forked, or heavily broken trees

	UNION ALL

	SELECT
		treep.NPP, treep.A, treep.ESS, treep.C13, treep.HTOT, treep.HDEC, treep.DDEC, treep.V, treep.DECOUPE

	FROM
		inv_exp_nm.p3arbre AS treep

	WHERE
		treep.ESS IN (", sp_code_nfi, ") AND -- Selected tree species
		(treep.veget = '0' OR treep.veget = 'Z') AND -- Alive
		COALESCE(treep.forme, 'X') NOT IN ('0', 'Z') AND -- No pollarded nor heavily pruned tree
		COALESCE(treep.tige, '1') = '1' -- No twin, forked, or heavily broken trees

	ORDER BY
		NPP, A;"), DT = TRUE)

disconnect_db(db)

## Remove NAs and compute hdecro (hdec from Florence, never knew why it was called hdecro...)
nfi_data_am = nfi_data_am[!is.na(ddec)]
nfi_data_am[, hdecro := hdec]
nfi_data_am[ddec == 0.07, hdecro := htot]

nfi_data_nm = nfi_data_nm[!is.na(hdec)]
nfi_data_nm[, hdecro := hdec]
nfi_data_nm[decoupe == "0", hdecro := htot] # Regulare trees

## Merge data
# Add origin
nfi_data_am[, method := "ancient"]
nfi_data_nm[, method := "new"]

# Remove non-common columns
nfi_data_am[, ddec := NULL]
nfi_data_nm[, c("decoupe", "ddec") := NULL]

# Merging
nfi_data = rbindlist(list(nfi_data_am, nfi_data_nm))

## Change names
setnames(nfi_data, new = c("plot_id", "tree_id", "sp_code", "circumference_m", "height",
	"taper_height", "bole_volume_m3", "taper_height_flo", "method"))

## Add scientific names
nfi_data = merge.data.table(nfi_data, nfi_codes, by.x = "sp_code", by.y = "species_code_nfi")

## Remove null volumes as judged irrealistic
nfi_data = nfi_data[bole_volume_m3 > 0]
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script is generates data to plot figures in pgfplots in the article. The figures that I have currently in mind are:
1. $\Vtot$ vs $\Vbft$ with 90% credibility interval (CI). Ideally, there would be two colours, one for the training range, and one for the NFI range if larger than training range (which I expect). Note that I will use the new formula for bole volumes rather than the values provided in the database...

## Figure `Vtot` vs `Vbole` with CI

I noticed that the range of values provided by the NFI are sometimes incoherent, such as 40 cubic meters for *Pinus laricio*, which is far beyond its usual range I think (although it is a tree 42.4 m tall and 5.52 m circumference)...

```{r}
range_dt = merge.data.table(
	training_dt[, .(min_t = min(bole_volume_m3), max_t = max(bole_volume_m3)), by = speciesName_sci],
	nfi_data[, .(min_nfi = min(bole_volume_m3), max_nfi = max(bole_volume_m3),
		q025_nfi = quantile(bole_volume_m3, 0.025), q975_nfi = quantile(bole_volume_m3, 0.975),
		q005_nfi = quantile(bole_volume_m3, 0.005), q995_nfi = quantile(bole_volume_m3, 0.995)), by = speciesName_sci])
```

### Predict the bole volumes based on Flopp model

This is done in ?? steps:
1. Load the training dataset of Florence
2. Use the function `pred_fct_flopp`, which returns the range of bole volume predictions on new data (here the NFI dataset already prepaired)

```{r}
#### Reading data
## Training data
florence_dt = readRDS("../../gohon_florence/arbComp.rds")
setkey(florence_dt, "ess")

florence_dt = florence_dt[nfi_codes[, species_code_nfi]]
setkey(florence_dt, "ess")

## Keeping for predictions only small and large trees as otherwise I overload tempdir() folder i.e., /tmp/Rtmp***
nfi_data[, c("q15", "q85") := as.list(quantile(bole_volume_m3, prob = c(0.15, 0.85))), by = speciesName_sci]
nfi_data_full = copy(nfi_data)
nfi_data = nfi_data[bole_volume_m3 < q15 | bole_volume_m3 > q85]
filename = "./range_nfi-pred_Vbole.rds"

if (file.exists(filename))
{
	range_dt = readRDS(filename)
} else {
	select_cols = c("min_nfi_flopp", "q005_flopp", "q025_flopp", "q975_flopp", "q995_flopp", "max_nfi_flopp")
	select_cols_u = c("min_nfi_flopp_u", "q005_flopp_u", "q025_flopp_u", "q975_flopp_u",
		"q995_flopp_u", "max_nfi_flopp_u")
	range_dt[, c(select_cols) :=
		.(-Inf, -Inf,-Inf, -Inf, -Inf, -Inf)]
	range_dt[, c(select_cols_u) :=
		.(-Inf, -Inf,-Inf, -Inf, -Inf, -Inf)]
	for (sp in nfi_codes[, species_code_nfi])
	{
		current_sp = nfi_codes[species_code_nfi == sp, speciesName_sci]
		print(current_sp)

		if (sp == "01")
		{
			warning("Not possible to run it simply for this species, so skiping")
			next
		}

		## Prepare data
		stanData_gen = list(
			N = florence_dt[sp][(type == "ajust"), .N],
			
			circumference_m = florence_dt[sp][(type == "ajust"), c13],
			taper_height_flo = florence_dt[sp][(type == "ajust"), hdecro],
			height = florence_dt[sp][(type == "ajust"), htot],

			N_new = nfi_data[sp][, .N],

			circumference_m_new = nfi_data[sp][, circumference_m],
			taper_height_flo_new = nfi_data[sp][, taper_height_flo],
			height_new = nfi_data[sp][, height],
			bole_volume_m3_new = nfi_data[sp][, bole_volume_m3]
		)

		temp = pred_fct_flopp(sp, stanData_gen)
		range_dt[current_sp, c(select_cols) := as.list(temp$quant_mean_pred)] # Without residual variance
		range_dt[current_sp, c(select_cols_u) := as.list(temp$quant_uncertain_pred)] # Incl. residual var.
		rm(temp)
	}
	saveRDS(range_dt, filename)
}

range_dt["Quercus sp.", q995_flopp_u := q995_nfi] # Because there is nothing, and I will use q995_flopp_u after
range_dt["Quercus sp.", min_nfi_flopp := min_nfi]

# Species need an extension if 110% max_t is below q995_flopp
range_dt[, need_extension := 1.1*max_t < q995_flopp_u]
```

As can be seen, for some species, the NFI goes below the minimum limit of the training dataset. However, since the limits are so close to zero, I will never been able to distinguish any prolungation outside of the lower range! So I simply plot up to 0.

### Compute total volumes and quantiles

The total above-ground volumes are computed on the largest range the training dataset and the 99.5 percentile of NFI bole volume (predicted by Flopp rather than Morneau). I only use 500 points equally spaced, spanning the whole range. For the generated data, a column is added to know whether or not we are in extrapolation.

```{r}
#### Make figures and generate data for pgfplots
n_indiv = 500

for (sp in nfi_codes[, species_code_nfi])
{
	current_sp = nfi_codes[species_code_nfi == sp, speciesName_sci]

	sp_filename = stri_replace(str = current_sp, replacement = "-", regex = " ")
	
	if (current_sp == "Quercus sp.")
		sp_filename = stri_replace(str = sp_filename, replacement = "", regex = "\\.")

	if (file.exists(paste0("../../../writing/article_allometry/data/pred/", sp_filename, ".csv")))
		next

	## Simulate total above ground volume on nfi or training range (depends which is largest)
	stanData = list(
		N = training_dt[current_sp, .N],
		bole_volume_m3 = training_dt[current_sp, bole_volume_m3],
		total_volume_m3 = training_dt[current_sp, total_volume_m3],
		N_new = n_indiv,
		bole_volume_m3_new = seq(range_dt[current_sp, min(min_t, min_nfi_flopp)],
			range_dt[current_sp, max(max_t, q995_flopp_u)], length.out = n_indiv)
	)

	stanData$total_volume_m3_new = 1.3*stanData$bole_volume_m3_new # Useless here, but must be provided for genQ to run

	fwrite(data.table(obs_bole_volume = stanData$bole_volume_m3, obs_total_volume = stanData$total_volume_m3),
		paste0("../../../writing/article_allometry/data/pred/", sp_filename, "_obs-data.csv"))

	simplif = FALSE
	if ((current_sp == "Fraxinus excelsior") || (current_sp == "Picea abies"))
		simplif = TRUE
	
	temp = pred_fct_yang(sp = sp_filename, stanData_gen = stanData, simplif = simplif)

	dt = data.table(bole_volume_m3 = stanData$bole_volume_m3_new, pred_avg = temp$pred, pred_avg_r = temp$pred_r,
		q05_v_u = temp$quant[, q05], q95_v_u = temp$quant[, q95],
		q05_v = temp$quant_mean["5%", ], q95_v = temp$quant_mean["95%", ],
		q05_r_u = temp$quant_r[, q05], q95_r_u = temp$quant_r[, q95],
		q05_r = temp$quant_mean_r["5%", ], q95_r = temp$quant_mean_r["95%", ]
	)
	dt[, beyond_rg := bole_volume_m3 > range_dt[current_sp, max_t]]
	dt[, beyond_rg_int := as.integer(beyond_rg)]

	## Plot
	# Compute boundaries
	max_x = dt[, max(bole_volume_m3)]
	max_y = max(temp$quant[, q95])
	min_y_r = min(temp$quant_r[, q05])
	max_y_r = max(temp$quant_r[, q95])
	if (max_y_r > 1)
		warning("Uncertainty leads beyond 1! How come with a Beta distribution??")

	# Smooth out the uncertainties, only to rmove small wiggles in the quantiles...
	m05_v = mgcv::gam(q05_v_u ~ s(bole_volume_m3, k = 30), data = dt)
	m95_v = mgcv::gam(q95_v_u ~ s(bole_volume_m3, k = 30), data = dt)
	m05_r = mgcv::gam(q05_r_u ~ s(bole_volume_m3, k = 30), data = dt)
	m95_r = mgcv::gam(q95_r_u ~ s(bole_volume_m3, k = 30), data = dt)

	# ... of volumes
	dt[, q05_v_u_s := predict(m05_v)]
	if (dt[, any(q05_v_u_s < 0)])
	{
		warning("Negative values due to smoothing out. Replacing by original value")
		dt[q05_v_u_s < 0, q05_v_u_s := q05_v_u]
	}
	
	dt[, q95_v_u_s := predict(m95_v)]
	if (dt[, any(q95_v_u_s < 0)])
	{
		warning("Negative values due to smoothing out. Replacing by original value")
		dt[q95_v_u_s < 0, q95_v_u_s := q95_v_u]
	}
	
	# ... of ratios
	dt[, q05_r_u_s := predict(m05_r)]
	if (dt[, any(q05_r_u_s < 0)])
	{
		warning("Negative values due to smoothing out. Replacing by original value")
		dt[q05_r_u_s < 0, q05_r_u_s := q05_r_u]
	}
	
	dt[, q95_r_u_s := predict(m95_r)]
	if (dt[, any(q95_r_u_s < 0)])
	{
		warning("Negative values due to smoothing out. Replacing by original value")
		dt[q95_r_u_s < 0, q95_r_u_s := q95_r_u]
	}

	# Empty plot with good extent
	plot(0, type = "n", xlim = c(0, max_x), ylim = c(0, max_y), axes = FALSE,
		xlab = "Bole volume", ylab = "Above-ground volume")
	
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), pred_avg], lwd = 2)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), pred_avg], lwd = 2, lty = "dashed")

	abline(a = 0, b = 1, lwd = 0.85, col = "#5A5A5A", lty = "dotted")
	
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q95_v_u_s], col = "#0F7BA2", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q95_v_u_s], col = "#0F7BA2", lwd = 1, lty = "dashed")
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q05_v_u_s], col = "#0F7BA2", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q05_v_u_s], col = "#0F7BA2", lwd = 1, lty = "dashed")
	
	polygon(c(dt[, bole_volume_m3], rev(dt[, bole_volume_m3])),
		c(dt[, q95_v_u_s], rev(dt[, q05_v_u_s])), col = "#0F7BA266", lty = 0)
	
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q95_v], col = "#DD5129", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q95_v], col = "#DD5129", lwd = 1, lty = "dashed")
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q05_v], col = "#DD5129", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q05_v], col = "#DD5129", lwd = 1, lty = "dashed")
	
	polygon(c(dt[, bole_volume_m3], rev(dt[, bole_volume_m3])),
		c(dt[, q95_v], rev(dt[, q05_v])), col = "#DD512966", lty = 0)
	
	# Empty plot with good extent for ratio
	plot(0, type = "n", xlim = c(0, max_x), ylim = c(min_y_r, max_y_r), axes = FALSE,
		xlab = "Bole volume", ylab = "Above-ground volume")
	
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), pred_avg_r], lwd = 2)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), pred_avg_r], lwd = 2, lty = "dashed")
	
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q95_r_u_s], col = "#0F7BA2", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q95_r_u_s], col = "#0F7BA2", lwd = 1, lty = "dashed")
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q05_r_u_s], col = "#0F7BA2", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q05_r_u_s], col = "#0F7BA2", lwd = 1, lty = "dashed")
	
	polygon(c(dt[, bole_volume_m3], rev(dt[, bole_volume_m3])),
		c(dt[, q95_r_u_s], rev(dt[, q05_r_u_s])), col = "#0F7BA266", lty = 0)
	
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q95_r], col = "#DD5129", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q95_r], col = "#DD5129", lwd = 1, lty = "dashed")
	lines(dt[!(beyond_rg), bole_volume_m3], dt[!(beyond_rg), q05_r], col = "#DD5129", lwd = 1)
	lines(dt[(beyond_rg), bole_volume_m3], dt[(beyond_rg), q05_r], col = "#DD5129", lwd = 1, lty = "dashed")
	
	polygon(c(dt[, bole_volume_m3], rev(dt[, bole_volume_m3])),
		c(dt[, q95_r], rev(dt[, q05_r])), col = "#DD512966", lty = 0)
	
	axis(1)
	axis(2, las = 1)

	## Save data for pgfplots
	fwrite(dt, paste0("../../../writing/article_allometry/data/pred/", sp_filename, ".csv"))
}
```

## Ratio figure

The asymptotic value of the ratio is surprisingly stable, around 0.8. Therefore, it would worth it to plot it for the four species of interest, *Abies alba*, *Fagus sylvatica*, *Picea abies*, and *Quercus petraea*. The aim is to compare at which bole volume do they reach the asymptote, and also check difference boradleaves/conifers. This could be discussed in the article, together with other traits

```{r}
#### Make figures and generate data for pgfplots
n_points = 750
vbole = seq(0, 7, length.out = n_points) # 7 is a good compromise for the 6 species I am interestied in

threshold_dt = data.table(speciesName_sci = training_dt[, unique(speciesName_sci)], simplif = FALSE,
	x1 = NA_real_, x2 = NA_real_, wp = NA_real_, c = NA_real_, eps = NA_real_, div_by = NA_integer_,
	delta_mc = NA_real_, diff_eps = NA_real_, key = "speciesName_sci")

params_dt = data.table(speciesName_sci = training_dt[, unique(speciesName_sci)], c = NA_real_,
	j = NA_real_, k = NA_real_, m = NA_real_, n = NA_real_, s = NA_real_, key = "speciesName_sci")

fct_output = matrix(data = NA_real_, nrow = threshold_dt[, .N] + 1, ncol = n_points)
rownames(fct_output) = c("vbole", threshold_dt[, speciesName_sci])
fct_output["vbole", ] = vbole

loaded = FALSE

if (!file.exists("./yang/lambert_calculus.rds"))
{
	for (sp in training_dt[, unique(speciesName_sci)])
	{
		print(paste("Doing species", sp))
		
		## Get params
		if ((sp == "Fraxinus excelsior") || (sp == "Picea abies"))
			next # Simplified species, done after
		
		sp_filename = stri_replace(str = sp, replacement = "-", regex = " ")
		if (sp == "Quercus sp.")
			sp_filename = stri_replace(str = sp_filename, replacement = "", regex = "\\.")
		
		filename = paste0("./yang/", sp_filename, "_new", ".rds")
		fit = readRDS(filename)

		paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")
		rm(fit)

		m = paramsVec["m"]
		j = paramsVec["j"]
		c = paramsVec["c"]
		k = paramsVec["k"]

		epsilon = 2/100*c # 2% of the asymptotic rate
		div_by = 1

		if (epsilon > (m - c))
		{
			warning(paste("The condition epsilon < m - c is not respeced for", sp))
			count = 0
			while (epsilon > (m - c) && count < 10)
			{
				count = count + 1
				epsilon = epsilon/2
			}

			if (epsilon > (m - c))
			{
				warning(paste("Could not solve the problem for", sp, "by setting epsilon = epsilon/1024"))
				next
			}
			warning(paste("Divided epsilon by", 2^count, "for", sp))
			div_by = 2^count
		}

		xi = -exp(-1) * (epsilon/(m - c))^(1/j)
		wp = -j/k * lamW::lambertW0(xi)
		wm = -j/k * lamW::lambertWm1(xi)

		if (wm < j/k)
			warning("My analytical calculus showed that this should not be possible...")

		s = paramsVec["s"]
		n = paramsVec["n"]

		params_dt[sp, c("c", "j", "k", "m", "n", "s") := as.list(paramsVec)]

		x2 = 1/s*(log(c - n) - log(epsilon))

		diff = abs(yang2024(wm, paramsVec) - c) - epsilon

		threshold_dt[sp, c("x1", "x2","wp", "c", "eps", "div_by", "delta_mc", "diff_eps") :=
			.(wm, ..x2, ..wp, ..c, epsilon, ..div_by, (..m - ..c), diff)]

		fct_output[sp, ] = yang2024(vbole, paramsVec)
	}
} else {
	threshold_dt = readRDS("./yang/lambert_calculus.rds")
	fct_output = readRDS("./yang/ratio_output.rds")
	params_dt = readRDS("./yang/avg_params.rds")
	loaded = TRUE
}
```

I am not sure about the interpretation of these threshold, but:
1. Among the three broadleaves (*Fagus sylvatica*, *Fraxinus excelsior*, and *Quercus petraea*), the earliest to reach stability. Could that be related to pioneering characteristics?
2. Any difference Broadleves/Conifers?
3. Any stuff on shade tolerance?

### Lambert function for *Fraxinus excelsior* and *Picea abies*

When I am using the simplified model, and despite it is a submodel of the general case, it seems that I need some cautiousness on the computation with Lambert function, especially with $\varepsilon$... Check p.12--13 Notebook 4.

```{r}
if (!loaded)
{
	for (sp in c("Fraxinus excelsior", "Picea abies"))
	{
		print(paste("Doing species", sp))

		sp_filename = stri_replace(str = sp, replacement = "-", regex = " ")
		
		filename = paste0("./yang/", sp_filename, "_submodel", ".rds")
		fit = readRDS(filename)
		
		paramsVec_simplif = getParams(model_cmdstan = fit,
			params_names = c("alpha", "beta_", "gamma", "delta"), type = "mean")
		
		rm(fit)
		
		alpha = paramsVec_simplif["alpha"] # Frax: 0.7159374        Picea: 0.8859609
		beta = paramsVec_simplif["beta_"]  # Frax: 5.184622         Picea: 13.23916
		gamma = paramsVec_simplif["gamma"] # Frax: 2.829555         Picea: 0.36472
		delta = paramsVec_simplif["delta"] # Frax: -0.02979472      Picea: -0.4933718
		
		epsilon = 2*alpha/100
		epsilon < gamma/beta*exp(beta*delta/gamma - 1)
		
		if (sp == "Picea abies")
			epsilon = 1e-11
		
		xi = -beta*epsilon/gamma*exp(-beta*delta/gamma)
		
		wp = -1/beta * (lamW::lambertW0(xi) + beta*delta/gamma)
		wm = -1/beta * (lamW::lambertWm1(xi) + beta*delta/gamma)

		paramsVec = c(
			c = unname(paramsVec_simplif["alpha"]),
			j = 1,
			k = unname(paramsVec_simplif["beta_"]),
			m = unname(exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"]),
			n = unname(paramsVec_simplif["delta"] + paramsVec_simplif["alpha"]),
			s = unname(paramsVec_simplif["beta_"])
		)
		fct_output[sp, ] = yang2024(vbole, paramsVec)

		params_dt[sp, c("c", "j", "k", "m", "n", "s") := as.list(paramsVec)]

		diff = abs(yang2024(wm, paramsVec) - alpha) - epsilon
	
		threshold_dt[sp, c("x1", "wp", "c", "eps", "div_by", "delta_mc", "diff_eps") :=
			.(wm, ..wp, ..alpha, epsilon, 1, gamma/beta*exp(-1), diff)]
	}

	saveRDS(threshold_dt, "./yang/lambert_calculus.rds")
	saveRDS(fct_output, "./yang/ratio_output.rds")
	saveRDS(params_dt, "./yang/avg_params.rds")
}

if (!file.exists("../../../writing/article_allometry/data/lambert.csv"))
	fwrite(threshold_dt, "../../../writing/article_allometry/data/lambert.csv")

threshold_dt[, .(speciesName_sci, x1, c)]
```

### Grouped figure

Actually, I decided to group species by type (broadleaf vs conifers), and to compute avg for the conifs are there are too many species.

```{r}
broad = c("Fagus sylvatica", "Fraxinus excelsior", "Quercus petraea") # Do not need Quercus sp.
conif = c("Abies alba", "Larix decidua", "Picea abies", "Pinus laricio", "Pinus nigra", "Pinus pinaster",
	"Pinus strobus", "Pinus sylvestris", "Pinus uncinata", "Pseudotsuga menziesii")

# --- Broadleaves
max_x = max(fct_output[vbole, ])
max_y = max(fct_output[broad, ])

# Empty plot with good extent
plot(0, type = "n", xlim = c(0, max_x), ylim = c(0, max_y), axes = FALSE,
	xlab = "Bole volume", ylab = "Ratio")
ls_colours = met.brewer("Egypt")[1:3]
names(ls_colours) = broad
for (sp in broad)
{
	lines(x = vbole, y = fct_output[sp, ], col = ls_colours[sp], lwd = 4)
	abline(v = threshold_dt[sp, x1], lty = "dashed", col = ls_colours[sp])
}
legend("bottomright", legend = broad, fill = ls_colours, bty = "n")
axis(1)
axis(2, las = 1)

# --- Conifers
selected_conif = c("Abies alba", "Picea abies", "Pinus nigra")
max_y = max(fct_output[conif, ])

# Empty plot with good extent
plot(0, type = "n", xlim = c(0, max_x), ylim = c(0, max_y), axes = FALSE,
	xlab = "Bole volume", ylab = "Ratio")

ls_colours = met.brewer("Egypt")[1:3]
names(ls_colours) = selected_conif

sp_min = params_dt[conif][which.min(c), speciesName_sci]
sp_max = params_dt[conif][which.max(c), speciesName_sci]

for (sp in selected_conif)
{
	lines(x = vbole, y = fct_output[sp, ], col = ls_colours[sp], lwd = 4)
	abline(v = threshold_dt[sp, x1], lty = "dashed", col = ls_colours[sp])
}
lines(x = vbole, y = fct_output[sp_min, ], lty = "dashed", lwd = 2)
lines(x = vbole, y = fct_output[sp_max, ], lty = "dashed", lwd = 2)
legend("bottomright", legend = selected_conif, fill = ls_colours, bty = "n")
axis(1)
axis(2, las = 1)

# --- All together
# Empty plot with good extent
plot(0, type = "n", xlim = c(0, max_x), ylim = c(0, 1), axes = FALSE,
	xlab = "Bole volume", ylab = "Ratio")

ls_colours = rep(met.brewer("Egypt")[1:3], 2)
names(ls_colours) = c(broad, selected_conif)

ls_lty = rep(c("solid", "dashed"), each = 3)
names(ls_lty) = c(broad, selected_conif)
# lims = apply(X = fct_output[conif, ], MARGIN = 2, FUN = quantile, probs = c(0, 0.5, 1))

for (sp in c(broad, selected_conif))
{
	lines(x = vbole, y = fct_output[sp, ], col = ls_colours[sp], lwd = 4, lty = ls_lty[sp])
	abline(v = threshold_dt[sp, x1], lty = ls_lty[sp], col = ls_colours[sp], lwd = 4)
}
# lines(x = vbole, y = fct_output[sp_min, ], lty = "dashed", lwd = 2)
# lines(x = vbole, y = fct_output[sp_max, ], lty = "dashed", lwd = 2)
abline(h = params_dt[sp_min, c], lty = "dashed", lwd = 0.65, col = "#5E5E5E")
abline(h = params_dt[sp_max, c], lty = "dashed", lwd = 0.65, col = "#5E5E5E")

legend("bottomright", legend = c(broad, selected_conif), col = ls_colours,
	lty = ls_lty, bty = "n", lwd = 4)
axis(1)
axis(2, las = 1)

```

### Trying few stuff for detecing flatness

I now try few stuff with the threshod data table.

```{r}
## Add ratio (m - c)/epsilon to know how many times peak above epsilon
threshold_dt[, eps_mc := delta_mc/eps]
threshold_dt[, .(speciesName_sci, x1, wp, eps_mc)]

## Add curvature at x_M (i.e., second derivative) at maximum x
mu_2nd = function(k, j, m, c, s, n)
	return(-k^2/j*(m - c) - s^2*(c - n)*exp(-s*j/k))

params_dt[, curve := mu_2nd(k, j, m, c, s, n), by = speciesName_sci]

mu_2nd_submodel = function(k, m, c, n)
{
	gamma = k*(m - c)*exp(1)
	delta = n - c
	return(-k*gamma*exp(-k*(1/k - delta/gamma)))
}

params_dt["Fraxinus excelsior", curve := mu_2nd_submodel(k, m, c, n)]
params_dt["Picea abies", curve := mu_2nd_submodel(k, m, c, n)]
```

### Export the data for pgfplots

```{r}
## Write fct_ouput
filename = "../../../writing/article_allometry/data/pred/ratio-fct_all-sp.csv"
if (!file.exists(filename))
{
	temp = as.data.table(t(fct_output))
	fwrite(temp, filename)
}

## Write treshold_dt
filename = "../../../writing/article_allometry/data/thresholds.csv"
threshold_dt = merge.data.table(x = threshold_dt, y = params_dt[, .(speciesName_sci, curve)], by = "speciesName_sci")
if (!file.exists(filename))
	fwrite(threshold_dt, filename)
```

## Compare NFI provided total volumes with my predictions

In this section, I request all the individual volumes of living trees recorded from 2005. The aim is to predict on these trees a new total volume and to compare with the current provided value (which is based on @Vallet2006).

```{r}
## List parametrised species and get their NFI code
species_trained = training_dt[, unique(speciesName_sci)]

nfi_codes = readRDS(paste0(path_data, "species_dt.rds"))[!is.na(species_code_nfi),
	.(speciesName_sci, species_code_nfi)]

## Modify fraxinus sp. by excelesior in French NFI
nfi_codes[species_code_nfi == "17", speciesName_sci := "Fraxinus excelsior"]

## Set key
setkey(nfi_codes, speciesName_sci)

sp_code_nfi_trained = paste0("'", paste(nfi_codes[species_trained, species_code_nfi], collapse = "','"), "'")
```

```{r}
#### Loading NFI data
## French NFI data query
db = connect_db()

nfi_data = exec_req(conn = db, req = paste0("
	SELECT
		tree.NPP, tree.A, tree.ESS, tree.C13, tree.HTOT, tree.HDEC, tree.DECOUPE, tree.V, tree.V0

	FROM
		inv_exp_nm.g3arbre AS tree

	WHERE
		tree.ESS IN (", sp_code_nfi_trained, ")

	ORDER BY
		NPP, A;"), DT = TRUE)

disconnect_db(db)

# info = nfi_data[, .(n_indiv = .N), by = ess]
# info[, percent := n_indiv/sum(n_indiv)*100]
```

```{r}
### Predict new volumes
nfi_data = merge.data.table(nfi_data, nfi_codes, by.x = "ess", by.y = "species_code_nfi")
setkey(nfi_data, speciesName_sci)

nfi_data[, any(is.na(v))]
nfi_data[, any(is.na(v0))]
nfi_data[, any(v == 0)]
nfi_data = nfi_data[v > 0]

nfi_data[, yang := yang2024_vol(v, params_dt[speciesName_sci]), by = speciesName_sci]

nfi_data[, .N, by = speciesName_sci]

for (sp in nfi_data[, unique(speciesName_sci)])
{
	plot(nfi_data[sp, c13], nfi_data[sp, v0], pch = 19,
		xlab = "Circumference", ylab = "Predicted total volume", main = sp)
	points(nfi_data[sp, c13], nfi_data[sp, yang], pch = 19,
		col = "#FAB255", cex = 0.65)
	ind = nfi_data[sp, v0 < v]
	points(nfi_data[sp][ind, c13], nfi_data[sp][ind, v0], pch = 19, cex = 2,
		col = "#0F7BA2")
}

min_circum = nfi_data[, min(c13)]

nfi_qt = nfi_data[, as.list(quantile(v0, probs = c(0, 0.025, 0.5, 0.975, 1))), by = speciesName_sci]
setnames(nfi_qt, new = c("species", "min", "q025", "med", "q975", "max"))
nfi_qt[, ind := .N:1]
nfi_qt[, maxRounded := round(max)]

yang_qt = nfi_data[, as.list(quantile(yang, probs = c(0, 0.025, 0.5, 0.975, 1))), by = speciesName_sci]
setnames(yang_qt, new = c("species", "min", "q025", "med", "q975", "max"))
yang_qt[, ind := .N:1]
yang_qt[, maxRounded := round(max)]

# Subset training_dt to compare with NFI, otherwise very small trees
train_qt = training_dt[(circumference_m >= min_circum) & (speciesName_sci != "Quercus sp."),
	as.list(quantile(total_volume_m3, probs = c(0, 0.025, 0.5, 0.975, 1))),
	by = speciesName_sci]
setnames(train_qt, new = c("species", "min", "q025", "med", "q975", "max"))
train_qt[, ind := .N:1]
train_qt[, maxRounded := round(max)]

filename = "../../../writing/article_allometry/data/span_nfi.csv"
if (!file.exists(filename))
	fwrite(nfi_qt, filename)

filename = "../../../writing/article_allometry/data/span_myPred.csv"
if (!file.exists(filename))
	fwrite(yang_qt, filename)

filename = "../../../writing/article_allometry/data/span_Dtot.csv"
if (!file.exists(filename))
	fwrite(train_qt, filename)
```

Some particular cases that deserves investigation.
```{r}
#### Fagus sylvatica
sp = "Fagus sylvatica"

weirdos = nfi_data[sp][(c13 > 4) & (yang < 10), .(speciesName_sci, c13, htot, hdec, decoupe, v, v0, yang)]
weirdos[decoupe == 0, hdec := htot] # Florence hdec
weirdos = na.omit(weirdos)

sp_code = nfi_codes[sp, species_code_nfi]

## Prepare data
stanData_gen = list(
	N = florence_dt[sp_code][(type == "ajust"), .N],
	
	circumference_m = florence_dt[sp_code][(type == "ajust"), c13],
	taper_height_flo = florence_dt[sp_code][(type == "ajust"), hdecro],
	height = florence_dt[sp_code][(type == "ajust"), htot],

	N_new = weirdos[, .N],

	circumference_m_new = weirdos[, c13],
	taper_height_flo_new = weirdos[, hdec],
	height_new = weirdos[, htot],
	bole_volume_m3_new = weirdos[, v]
)

temp = pred_fct_flopp(sp_code, stanData_gen)

weirdos[, v_flopp := temp[["pred"]]]
weirdos[, yang_flopp := yang2024_vol(v_flopp, params_dt[sp])]

fagsyl = weirdos[, .(v, v_flopp, v0, yang_flopp)]


#### Fraxinus excelsior
sp = "Fraxinus excelsior"

weirdos = nfi_data[sp][(c13 > 3) & (yang < 10), .(speciesName_sci, c13, htot, hdec, decoupe, v, v0, yang)]
weirdos[decoupe == 0, hdec := htot] # Florence hdec
weirdos = na.omit(weirdos)

sp_code = nfi_codes[sp, species_code_nfi]

## Prepare data
stanData_gen = list(
	N = florence_dt[sp_code][(type == "ajust"), .N],
	
	circumference_m = florence_dt[sp_code][(type == "ajust"), c13],
	taper_height_flo = florence_dt[sp_code][(type == "ajust"), hdecro],
	height = florence_dt[sp_code][(type == "ajust"), htot],

	N_new = weirdos[, .N],

	circumference_m_new = weirdos[, c13],
	taper_height_flo_new = weirdos[, hdec],
	height_new = weirdos[, htot],
	bole_volume_m3_new = weirdos[, v]
)

temp = pred_fct_flopp(sp_code, stanData_gen)

weirdos[, v_flopp := temp[["pred"]]]
weirdos[, yang_flopp := yang2024_vol(v_flopp, params_dt[sp])]

fraexc = weirdos[, .(v, v_flopp, v0, yang_flopp)]


#### Pinus laricio
sp = "Pinus laricio"

weirdos = nfi_data[sp][(c13 > 3) & (yang < 10), .(speciesName_sci, c13, htot, hdec, decoupe, v, v0, yang)]
weirdos[decoupe == 0, hdec := htot] # Florence hdec
weirdos = na.omit(weirdos)

sp_code = nfi_codes[sp, species_code_nfi]

## Prepare data
stanData_gen = list(
	N = florence_dt[sp_code][(type == "ajust"), .N],
	
	circumference_m = florence_dt[sp_code][(type == "ajust"), c13],
	taper_height_flo = florence_dt[sp_code][(type == "ajust"), hdecro],
	height = florence_dt[sp_code][(type == "ajust"), htot],

	N_new = weirdos[, .N],

	circumference_m_new = weirdos[, c13],
	taper_height_flo_new = weirdos[, hdec],
	height_new = weirdos[, htot],
	bole_volume_m3_new = weirdos[, v]
)

temp = pred_fct_flopp(sp_code, stanData_gen)

weirdos[, v_flopp := temp[["pred"]]]
weirdos[, yang_flopp := yang2024_vol(v_flopp, params_dt[sp])]

pinlar = weirdos[, .(v, v_flopp, v0, yang_flopp)]


#### Quercus petraea
sp = "Quercus petraea"

weirdos = nfi_data[sp][(c13 > 3.5) & (yang < 16), .(speciesName_sci, c13, htot, hdec, decoupe, v, v0, yang)]
weirdos[decoupe == 0, hdec := htot] # Florence hdec
weirdos = na.omit(weirdos)

sp_code = nfi_codes[sp, species_code_nfi]

## Prepare data
stanData_gen = list(
	N = florence_dt[sp_code][(type == "ajust"), .N],
	
	circumference_m = florence_dt[sp_code][(type == "ajust"), c13],
	taper_height_flo = florence_dt[sp_code][(type == "ajust"), hdecro],
	height = florence_dt[sp_code][(type == "ajust"), htot],

	N_new = weirdos[, .N],

	circumference_m_new = weirdos[, c13],
	taper_height_flo_new = weirdos[, hdec],
	height_new = weirdos[, htot],
	bole_volume_m3_new = weirdos[, v]
)

temp = pred_fct_flopp(sp_code, stanData_gen)

weirdos[, v_flopp := temp[["pred"]]]
weirdos[, yang_flopp := yang2024_vol(v_flopp, params_dt[sp])]

quepet = weirdos[, .(v, v_flopp, v0, yang_flopp)]


weirdos = rbindlist(list("Fagus sylvatica" = fagsyl, "Fraxinus excelsior" = fraexc,
	"Pinus laricio" = pinlar, "Quercus petraea" = quepet), idcol = "speciesName_sci")

ls_colours = met.brewer("Egypt")[1:4]
names(ls_colours) = weirdos[, unique(speciesName_sci)]

weirdos[, colour := ls_colours[speciesName_sci], by = speciesName_sci]

## Plot bole volumes NFI vs Florence
plot(weirdos[, v_flopp], weirdos[, v], col = weirdos[, colour], pch = 19,
	axes = FALSE, xlab = "Bole volume Florence", ylab = "Bole volume NFI")
abline(a = 0, b = 1, lty = "dotted", col = "#515151", lwd = 0.65)
axis(1)
axis(2, las = 1)

## Plot total volumes NFI vs me
plot(weirdos[, yang_flopp], weirdos[, v0], col = weirdos[, colour], pch = 19,
	axes = FALSE, xlab = "Total volume me", ylab = "Total volume NFI")
abline(a = 0, b = 1, lty = "dotted", col = "#515151", lwd = 0.65)
axis(1)
axis(2, las = 1)


lim_y = range(weirdos[, .(v0, yang_flopp)])

plot(weirdos[, v_flopp], weirdos[, v0], col = weirdos[, colour], pch = 19, ylim = lim_y,
	axes = FALSE, xlab = "Bole volume Florence", ylab = "Total volume")
points(weirdos[, v_flopp], weirdos[, yang_flopp], col = weirdos[, colour], pch = 4)
abline(a = 0, b = 1, lty = "dotted", col = "#515151", lwd = 0.65)
axis(1)
axis(2, las = 1)
```