---
title: "Comparison with Vallet and output preparation for pgfplots"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(loo)

## Tool function
# source("./toolFunctions.R")
```

```{r}
#| output: false

#### Load the training data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading training dataset
training_dt = readRDS(paste0(path_data, "data_Vtot.rds"))
training_dt = training_dt[(!is.na(circumference_m)) & (!is.na(height))]
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

The aim of this document is to run a Bayesian version of @Vallet2006, and to compare it with:
1. The fitted values from @Vallet2006 in Frequentist (to make sure I have similar predictions and that Bayesian does not deviate from the Frequentist)
2. My model (which hopefully is better at predicting)

Part of this stuff was already done in the exploratory file `06_zhou.qmd`. I just gather everything and add what is necessary.

## Fit @Vallet2006 in Bayesian

Here is the model (taken from `06_zhou.qmd`):

```{stan output.var = "vallet_bayesian", cache = TRUE}
//| echo: true

functions {
	vector vallet(vector circumference, vector hdn, vector params, int N_params)
	{
		if (N_params == 4)
			return (params[1] + params[2]*circumference + params[3]*hdn).*(1 + params[4]./circumference.^2);
		
		if (N_params == 3)
			return params[1] + params[2]*circumference + params[3]*hdn;
		
		return params[1] + params[2]*circumference; // 2 parameters, Norway spruce
	}
}

data{
	// Dimensions
	int <lower = 1> N; // Number of trees // In m, while in cm in Vallet2006
	int <lower = 3, upper = 4> N_params; // Number of parameters

	// Predictors
	vector[N] circumference_cm;
	vector[N] height;

	// Observations
	vector[N] total_volume_m3;
}

transformed data {
	vector[N] hdn = sqrt(circumference_cm) ./ height;
}

parameters {
	vector[N_params] vec_params;
	real<lower = 0> sigma;
}

model {
	target += normal_lpdf(vec_params[1] | 0, 1);
	target += normal_lpdf(vec_params[2] | 0, 0.01);
	target += normal_lpdf(vec_params[3] | 0, 1);

	if (N_params == 4)
		target += normal_lpdf(vec_params[4] | 45, 10);
	
	target += gamma_lpdf(sigma | 0.06^2/0.01, 0.06/0.01);
	
	target += normal_lpdf(total_volume_m3 |
		vallet(circumference_cm, hdn, vec_params, N_params) ./ (40000*pi()) .* circumference_cm.^2 .* height, sigma);
}

```

I now run it on the four species of interest: *Abies alba*, *Fagus sylvatica*, *Picea abies*, and *Quercus petraea*. Note that Douglas fir cannot be parametrised with the way I wrote the model, as it is a particular case with $\alpha$, $\beta$, and $\delta$ estimated (while all the other three-parameter species have $\gamma$ estimated rather than the correction parameter $\delta$).

```{r}
ls_species = c("Abies alba", "Fagus sylvatica", "Picea abies", "Quercus petraea")
training_dt = training_dt[ls_species]


training_dt[is.na(circumference_m)]
training_dt[, .N, by = speciesName_sci]

# Table containing number of parameters (taken from Vallet 2006) as Vallet uses submodels
n_params = data.table(speciesName_sci = ls_species, n_params = c(3, 4, 2, 3), key = "speciesName_sci") 

output_dir = "./bayesian_vallet/"
if (!dir.exists(output_dir))
	dir.create(output_dir)

n_chains = 4

for (sp in ls_species)
{
	sp_filename = stri_replace(str = sp, replacement = "-", regex = " ")
	filename = paste0(output_dir, sp_filename, ".rds")
	if (!file.exists(filename))
	{
		stanData = list(
			N = training_dt[sp, .N],
			N_params = n_params[sp, n_params],

			circumference_cm = training_dt[sp, 100*circumference_m], # Was in cm in Vallet2006!
			height = training_dt[sp, height],

			total_volume_m3 = training_dt[sp, total_volume_m3]
		)

		results = vallet_bayesian$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
			max_treedepth = 12)

		results$save_output_files(dir = output_dir, basename = sp_filename, random = FALSE)
		saveRDS(results, filename)
	} else {
		results = readRDS(filename)
	}
}
```

