---
title: "Comparison with Vallet and output preparation for pgfplots"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(loo)

## Tool function
# source("./toolFunctions.R")

# Function to predict total volume from Vallet 2006
form_vallet = function(circumference_cm, height, params, n_params, is_douglas)
{
	res = params["alpha"] + params["beta"]*circumference_cm

	if (is_douglas == 1)
		return(res*(1 + params["delta"]/circumference_cm^2))
	
	if (n_params >= 3)
		res = res + params["gamma"]*sqrt(circumference_cm)/height

	if (n_params == 4)
		res = res*(1 + params["delta"]/circumference_cm^2)
	
	res = res/(40000*pi) * circumference_cm^2*height # Conversion to volume
	return(res)
}

# Get fixed values parameters (will not work for draws with third dimension > 1)
getParams = function(model_cmdstan, params_names, type = "mean", ...)
{
	if (!(type %in% c("all", "chain", "chain-iter", "mean", "median", "quantile")))
		stop("Unknown type. Please choose all, iter-chain, mean, median, or quantile")

	args = list(...)

	if (type %in% c("chain", "chain-iter", "mean", "median"))
	{
		vals = numeric(length(params_names))
		names(vals) = params_names

		if (type == "chain")
		{
			if (!("chain" %in% names(args)))
				stop("You must provide chain number when using the option chain")

			chain = args[["chain"]]
			if (chain > model_cmdstan$num_chains())
				stop("chain cannot be larger than num_chains")

			for (i in seq_along(params_names))
			{
				draws = model_cmdstan$draws(params_names[i])
				if (dim(draws)[3] != 1)
					stop("This function is not yet designed to handle a multi-params")
				vals[i] = mean(draws[, chain, 1])
			}
			return(vals)
		} else if (type == "chain-iter") {
			if (!all(c("iter", "chain") %in% names(args)))
				stop("You must provide iter and chain when using the option iter-chain")
			iter = args[["iter"]]
			chain = args[["chain"]]

			if (iter > model_cmdstan$metadata()$iter_sampling)
				stop("iter cannot be larger than iter_sampling")

			if (chain > model_cmdstan$num_chains())
				stop("chain cannot be larger than num_chains")

			for (i in seq_along(params_names))
			{
				draws = model_cmdstan$draws(params_names[i])
				if (dim(draws)[3] != 1)
					stop("This function is not yet designed to handle a multi-params")
				vals[i] = draws[iter, chain, 1]
			}
			return(vals)
		} else {
			for (i in seq_along(params_names))
			{
				vals[i] = ifelse(type == "mean",
					mean(model_cmdstan$draws(params_names[i])),
					median(model_cmdstan$draws(params_names[i])))
			}
			return(vals)
		}
	} else if (type == "quantile") {
		if (!("probs" %in% names(args)))
		{
			probs = c(0, 0.05, 0.5, 0.95, 1)
			vals = data.table(parameter = params_names, min = 0, q05 = 0, med = 0, avg = 0, q95 = 0, max = 0)
		}
		if ("probs" %in% names(args))
		{
			warning("I have not yet coded the general output, so far I use probs = c(0, 0.05, 0.5, 0.95, 1)")
			probs = c(0, 0.05, 0.5, 0.95, 1)
			vals = data.table(parameter = params_names, min = 0, q05 = 0, med = 0, avg = 0, q95 = 0, max = 0)
		}
		for (i in seq_along(params_names))
		{
			vals[i, c("min", "q05", "med", "q95", "max") := as.list(quantile(model_cmdstan$draws(params_names[i]), c(0, 0.05, 0.5, 0.95, 1)))]
			vals[i, avg := mean(model_cmdstan$draws(params_names[i]))]
		}
		return(vals)
	}

	vals = model_cmdstan$draws(params_names)

	return(vals)
}
```

```{r}
#| output: false

#### Load the training data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading training dataset
training_dt = readRDS(paste0(path_data, "data_Vtot.rds"))
training_dt = training_dt[(!is.na(circumference_m)) & (!is.na(height))]

## Table 5 from Vallet2006
vallet_dt = data.table(speciesName_sci = c("Abies alba", "Fagus sylvatica", "Picea abies", "Pinus pinaster", "Pinus sylvestris", "Pseudotsuga menziesii", "Quercus petraea"),
	alpha = c(0.550, 0.395, 0.631, 0.235, 0.297, 0.534, 0.471),
	sd_alpha = c(0.015, 0.01, 0.007, 0.041, 0.025, 0.01, 0.014),
	beta = c(7.49e-4, 2.66e-4, -9.46e-4, 9.7e-4, 3.18e-4, -5.3e-4, -3.45e-4),
	sd_beta = c(3.9e-5, 4.9e-5, 7.2e-5, 4.06e-4, 1.4e-4, 9.7e-5, 1.3e-5),
	gamma = c(0.277, 0.421, NA, 0.396, 0.384, NA, 0.377),
	sd_gamma = c(0.034, 0.025, NA, 0.057, 0.058, NA, 0.031),
	delta = c(NA, 45.4, NA, 198.8, 204.0, 56.6, NA),
	sd_delta = c(NA, 4, NA, 40, 26.6, 13.8, NA),
	var_res = c(0.0031, 0.0036, 0.0023, 0.0079, 0.0028, 0.0041 , 0.004),
	n_params = c(3, 4, 2, 4, 4, 3, 3), key = "speciesName_sci")
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

The aim of this document is to run a Bayesian version of @Vallet2006, and to compare it with:
1. The fitted values from @Vallet2006 in Frequentist (to make sure I have similar predictions and that Bayesian does not deviate from the Frequentist)
2. My model (which hopefully is better at predicting)

Part of this stuff was already done in the exploratory file `06_zhou.qmd`. I just gather everything and add what is necessary.

## Fit @Vallet2006 in Bayesian

Here is the model (taken from `06_zhou.qmd`):

```{stan output.var = "vallet_bayesian", cache = TRUE}
//| echo: true

functions {
	vector vallet(vector circumference, vector hdn, vector params, int N_params, int is_douglas)
	{
		if (is_douglas == 1)
			return (params[1] + params[2]*circumference).*(1 + params[3]./circumference.^2);
		
		if (N_params == 4)
			return (params[1] + params[2]*circumference + params[3]*hdn).*(1 + params[4]./circumference.^2);
		
		if (N_params == 3)
			return params[1] + params[2]*circumference + params[3]*hdn;
		
		return params[1] + params[2]*circumference; // 2 parameters, Norway spruce
	}
}

data{
	// Dimensions
	int <lower = 1> N; // Number of trees
	int <lower = 2, upper = 4> N_params; // Number of parameters
	int <lower = 0, upper = 1> is_douglas; // 0 if not, and 1 otherwise

	// Predictors
	vector[N] circumference_cm; // In m, while in cm in Vallet2006
	vector[N] height;

	// Observations
	vector[N] total_volume_m3;
}

transformed data {
	vector[N] hdn = sqrt(circumference_cm) ./ height;
}

parameters {
	vector[N_params] vec_params;
	real<lower = 0> sigma;
}

model {
	target += normal_lpdf(vec_params[1] | 0, 0.4);
	target += normal_lpdf(vec_params[2] | 0, 0.0005);

	if (N_params > 2 && is_douglas != 1)
		target += normal_lpdf(vec_params[3] | 0, 0.4);
	
	if (is_douglas == 1)
		target += normal_lpdf(vec_params[3] | 45, 10);

	if (N_params == 4)
		target += normal_lpdf(vec_params[4] | 45, 10);
	
	target += gamma_lpdf(sigma | 0.06^2/0.01, 0.06/0.01);
	
	target += normal_lpdf(total_volume_m3 | vallet(circumference_cm, hdn, vec_params, N_params, is_douglas) ./
		(40000*pi()) .* circumference_cm.^2 .* height, sigma);
}

```

I now run it on the species parametrised in @Vallet2006

```{r}
#| fig-cap: Bayesian vs Frequentist Vallet 2006
#| fig-subcap:
#|   - "Abies alba"
#|   - "Fagus sylvatica"
#|   - "Picea abies"
#|   - "Pinus pinaster"
#|   - "Pinus sylvestris"
#|   - "Pseudotsuga menziesii"
#|   - "Quercus petraea"
#| layout-ncol: 2
#| label: fig-bayesFreq

ls_species = c("Abies alba", "Fagus sylvatica", "Picea abies", "Pinus pinaster", "Pinus sylvestris",
	"Pseudotsuga menziesii", "Quercus petraea")
training_dt = training_dt[ls_species]

output_dir = "./bayesian_vallet/"
if (!dir.exists(output_dir))
	dir.create(output_dir)

n_chains = 4

for (sp in ls_species)
{
	sp_filename = stri_replace(str = sp, replacement = "-", regex = " ")
	filename = paste0(output_dir, sp_filename, ".rds")
	is_douglas = ifelse(sp == "Pseudotsuga menziesii", 1, 0) # Boolean style compatible with Stan language

	if (!file.exists(filename))
	{
		stanData = list(
			N = training_dt[sp, .N],
			N_params = vallet_dt[sp, n_params],
			is_douglas = is_douglas,

			circumference_cm = training_dt[sp, 100*circumference_m], # Was in cm in Vallet2006!
			height = training_dt[sp, height],

			total_volume_m3 = training_dt[sp, total_volume_m3]
		)

		results = vallet_bayesian$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4),
			max_treedepth = 12)

		results$save_output_files(dir = output_dir, basename = sp_filename, random = FALSE)
		saveRDS(results, filename)
	} else {
		results = readRDS(filename)
	}

	params = getParams(results, paste0("vec_params[", 1:vallet_dt[sp, n_params], "]"), type = "mean")
	
	if (vallet_dt[sp, n_params] == 2)
	{
		names(params) = c("alpha", "beta")
		freq_params = vallet_dt[sp, c(alpha, beta)]
		names(freq_params) = names(params)
	}
	
	if (vallet_dt[sp, n_params] == 3)
	{
		names(params) = c("alpha", "beta", "gamma")
		freq_params = vallet_dt[sp, c(alpha, beta, gamma)]
		names(freq_params) = names(params)

		if (is_douglas)
		{
			names(params) = c("alpha", "beta", "delta")
			freq_params = vallet_dt[sp, c(alpha, beta, delta)]
			names(freq_params) = names(params)
		}
	}
	
	if (vallet_dt[sp, n_params] == 4)
	{
		names(params) = c("alpha", "beta", "gamma", "delta")
		freq_params = vallet_dt[sp, c(alpha, beta, gamma, delta)]
		names(freq_params) = names(params)
	}
	
	training_dt[sp, bayes_vallet :=
			form_vallet(100*circumference_m, height, params, vallet_dt[sp, n_params], is_douglas)]
	training_dt[sp, freq_vallet :=
			form_vallet(100*circumference_m, height, freq_params, vallet_dt[sp, n_params], is_douglas)]

	## Plot Bayes vs Freq
	plot(training_dt[sp, freq_vallet], training_dt[sp, bayes_vallet], pch = 19, cex = 0.55,
		axes = FALSE, xlab = "Frequentist", ylab = "Bayesian")
	abline(a = 0, b = 1, lwd = 1.5, col = "#CD212A")
	axis(1)
	axis(2, las = 1)
}

if (!file.exists("./bayesian_vallet/pred_dt-7species.rds"))
	saveRDS(training_dt, "./bayesian_vallet/pred_dt-7species.rds")
```

## Compare Vallet Bayesian with my model

### Generate quantities Vallet

```{stan output.var = "vallet_genQ", cache = TRUE}
//| echo: true

functions {
	vector vallet(vector circumference, vector hdn, vector params, int N_params, int is_douglas)
	{
		if (is_douglas == 1)
			return (params[1] + params[2]*circumference).*(1 + params[3]./circumference.^2);
		
		if (N_params == 4)
			return (params[1] + params[2]*circumference + params[3]*hdn).*(1 + params[4]./circumference.^2);
		
		if (N_params == 3)
			return params[1] + params[2]*circumference + params[3]*hdn;
		
		return params[1] + params[2]*circumference; // 2 parameters, Norway spruce
	}
}

data{
	// Dimensions
	int <lower = 1> N; // Number of trees
	int <lower = 2, upper = 4> N_params; // Number of parameters
	int <lower = 0, upper = 1> is_douglas; // 0 if not, and 1 otherwise

	// Predictors
	vector[N] circumference_cm; // In m, while in cm in Vallet2006
	vector[N] height;

	// Observations
	vector[N] total_volume_m3;
	
	// --------------   NEW OBSERVATIONS   --------------
	// Dimensions
	int <lower = 1> N_new;

	// Predictors
	vector[N_new] circumference_cm_new;
	vector[N_new] height_new;

	// Observations
	vector[N_new] total_volume_m3_new;
}

transformed data {
	vector[N] hdn = sqrt(circumference_cm) ./ height;
	vector[N_new] hdn_new = sqrt(circumference_cm_new) ./ height_new;
}

parameters {
	vector[N_params] vec_params;
	real<lower = 0> sigma;
}

generated quantities {
	vector[N_new] v_gen_mean = vallet(circumference_cm_new, hdn_new, vec_params, N_params, is_douglas) ./
		(40000*pi()) .* circumference_cm_new.^2 .* height_new;
	array[N_new] real v_gen = normal_rng(v_gen_mean, sigma);
	
	// Log likelihood of newly observed volumes given fitted params on (other) data
	vector [N_new] log_lik;
	
	for (i in 1:N_new)
		log_lik[i] = normal_lpdf(total_volume_m3_new[i] | v_gen_mean[i], sigma);
}
	
```

```{r}
#| fig-cap: Vallet2006 vs mine (i.e., Yang2024 modified)
#| fig-subcap:
#|   - "Abies alba"
#|   - "Abies alba"
#|   - "Fagus sylvatica"
#|   - "Fagus sylvatica"
#|   - "Picea abies"
#|   - "Picea abies"
#|   - "Pinus pinaster"
#|   - "Pinus pinaster"
#|   - "Pinus sylvestris"
#|   - "Pinus sylvestris"
#|   - "Pseudotsuga menziesii"
#|   - "Pseudotsuga menziesii"
#|   - "Quercus petraea"
#|   - "Quercus petraea"
#| layout-ncol: 2
#| label: fig-valletYang
#| output: false

vallet_genQ = cmdstan_model("./vallet.stan")
genQ = cmdstan_model("./yang2024-genQ.stan")
genQ_sub = cmdstan_model("./yang2024-genQ_submodel.stan")

comp_dt = data.table(speciesName_sci = ls_species, best = "", elpd_diff = -Inf, se_diff = -Inf,
	wrn_yang = "", wrn_vallet = "", key = "speciesName_sci")

for (sp in ls_species)
{
	sp_filename = stri_replace(str = sp, replacement = "-", regex = " ")
	filename = paste0(output_dir, sp_filename, ".rds")
	is_douglas = ifelse(sp == "Pseudotsuga menziesii", 1, 0) # Boolean style compatible with Stan language

	stanData_gen = list(
		N = training_dt[sp, .N],
		N_params = vallet_dt[sp, n_params],
		is_douglas = is_douglas,

		bole_volume_m3 = training_dt[sp, bole_volume_m3],
		circumference_cm = training_dt[sp, 100*circumference_m], # Was in cm in Vallet2006!
		height = training_dt[sp, height],

		total_volume_m3 = training_dt[sp, total_volume_m3],

		# New data, which are the same...
		N_new = training_dt[sp, .N],

		bole_volume_m3_new = training_dt[sp, bole_volume_m3],
		circumference_cm_new = training_dt[sp, 100*circumference_m], # Was in cm in Vallet2006!
		height_new = training_dt[sp, height],

		total_volume_m3_new = training_dt[sp, total_volume_m3]
	)

	# Generate data Bayesian vallet
	fit_vallet = readRDS(filename)
	sim_vallet = vallet_genQ$generate_quantities(fitted_params = fit_vallet, data = stanData_gen,
		parallel_chains = min(4, n_chains))
	
	# Generate data my model
	if (sp %in% c("Fraxinus excelsior", "Picea abies"))
	{
		sp_filename = paste0(sp_filename, "_submodel")
		filename = paste0("./yang/", sp_filename, ".rds")
		fit = readRDS(filename)
		sim = genQ_sub$generate_quantities(fitted_params = fit, data = stanData_gen,
			parallel_chains = min(4, n_chains))

	} else {
		filename = paste0("./yang/", sp_filename, ".rds")
		fit = readRDS(filename)
		sim = genQ$generate_quantities(fitted_params = fit, data = stanData_gen,
			parallel_chains = min(4, n_chains))
	}

	## Compute PSIS-LOO...
	# ... for Vallet
	r_eff = loo::relative_eff(exp(sim_vallet$draws("log_lik")), cores = 8)
	loo_vallet = loo::loo(x = sim_vallet$draws("log_lik"), r_eff = r_eff, cores = 8)

	warning_vallet = "none"
	if (any(loo_vallet$diagnostics$pareto_k >= 0.7))
	{
		warning_vallet = "bad"
		if (any(loo_vallet$diagnostics$pareto_k > 1))
			warning_vallet = "very bad"
	}
	
	# ... for my model/submodel
	r_eff = loo::relative_eff(exp(sim$draws("log_lik")), cores = 8)
	loo_yang = loo::loo(x = sim$draws("log_lik"), r_eff = r_eff, cores = 8)

	warning_yang = "none"
	if (any(loo_yang$diagnostics$pareto_k >= 0.7))
	{
		warning_yang = "bad"
		if (any(loo_yang$diagnostics$pareto_k > 1))
			warning_yang = "very bad"
	}

	comp = loo::loo_compare(list(vallet = loo_vallet, yang = loo_yang))

	## Record comparison output
	comp_dt[sp, c("best", "elpd_diff", "se_diff") :=
		.(rownames(comp)[1], comp[2, "elpd_diff"], comp[2, "se_diff"])]
	comp_dt[sp, c("wrn_yang", "wrn_vallet") :=
		.(warning_yang, warning_vallet)]

	## Check pred vs obs
	pred_vallet = apply(X = sim_vallet$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
	pred_yang = apply(X = sim$draws("v_gen_mean"), MARGIN = 3, FUN = mean)
	
	rm(sim, sim_vallet, fit, fit_vallet)
	
	plot(pred_vallet, stanData_gen$total_volume_m3, pch = 19, cex = 0.65,
		xlab = "Predicted", ylab = "Observed", axes = FALSE)
	points(pred_vallet, pred_yang, col = "#FAB255", pch = 19, cex = 0.45)
	axis(1)
	axis(2, las = 1)
	abline(a = 0, b = 1, lwd = 2, col = "#CD212A")
	
	## Check pred vs bole volume, is it always above the 1:1 line?
	plot(stanData_gen$bole_volume_m3, pred_vallet, pch = 19, cex = 0.65,
		xlab = "Predicted", ylab = "Observed", axes = FALSE)
	points(stanData_gen$bole_volume_m3, pred_yang, col = "#FAB255", pch = 19, cex = 0.45)
	ind = pred_vallet < stanData_gen$bole_volume_m3
	if (any(ind))
		points(stanData_gen$bole_volume_m3[ind], pred_vallet[ind], col = "#0F7BA2",
			pch = 19, cex = 0.75)
	axis(1)
	axis(2, las = 1)
	abline(a = 0, b = 1, lty = "dotted")
}

if (!file.exists("./bayesian_vallet/comparison_vallet-yang.rds"))
	saveRDS(comp_dt, "./bayesian_vallet/comparison_vallet-yang.rds")
```
