---
title: "Run the ratio models (complete and submodel) for species groups"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)

## Tool function
source("./toolFunctions.R")

plot_sp = function(fit, sp, forest = tree_dt[.(sp)], pred = TRUE, simplif = FALSE,
	n_bins = 4, pal = "Hiroshige", selected_variable = NULL, print_plot = TRUE)
{
	# Get parameters
	n_sampling = fit$metadata()$iter_sampling
	n_chains = fit$num_chains()

	if (n_chains != nrow(fit$metadata()$time))
	{
		warning("Some chains are missing probably because they could not start! Using the number of successful chains")
		n_chains = nrow(fit$metadata()$time)
	}

	# Add colour to the measured points according to selected variable
	if (!is.null(selected_variable))
	{
		if (!(selected_variable %in% colnames(forest)))
			stop(paste0("'", selected_variable, "' is not a column of forest"))
		
		colours = MetBrewer::met.brewer(pal, n_bins)[1:n_bins]
		forest[, colour_ind := as.numeric(cut(.SD[[selected_variable]], breaks = n_bins))] # Indices per category
		forest[, colour := colours[colour_ind]] # Map selected variable to colour using indices
	} else {
		forest[, colour := "#3355AA33"] # Default colour
	}

	# Define fitted functions (depends if simplified model 2 stages vs 3 stages)
	if (!simplif)
	{
		paramsVec = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "mean")
	} else {
		paramsVec_simplif = getParams(model_cmdstan = fit,
			params_names = c("alpha", "beta_", "gamma", "delta"), type = "mean")
		paramsVec = c(
			c = unname(paramsVec_simplif["alpha"]),
			j = 1,
			k = unname(paramsVec_simplif["beta_"]),
			m = unname(exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"]),
			n = unname(paramsVec_simplif["delta"] + paramsVec_simplif["alpha"]),
			s = unname(paramsVec_simplif["beta_"])
		)
	}

	r_func = function(x, pars)
	{
		return ((pars["m"] - pars["c"]) * exp(pars["j"] - pars["k"]*x) * (pars["k"]*x/pars["j"])^pars["j"] +
			pars["c"] - (pars["c"] - pars["n"])*exp(-pars["s"]*x));
	}
	
	psi = function(x, pars)
	{
		r = r_func(x, pars)

		return ( 1/pars["c"] * x^( 1 - (log(r) - log(pars["c"])) / log(x) ) )
	}

	# Check rhats
	pars_names = names(paramsVec)
	if (simplif)
		pars_names = c("alpha", "beta_", "gamma", "delta")
	rhats = rhat(object = fit, pars = pars_names)

	# Divergence check up
	params = getParams(model_cmdstan = fit, params_names = pars_names, type = "all")
	div = posterior::subset_draws(fit$sampler_diagnostics(), variable = "divergent__")

	if (!print_plot & !pred)
		return (list(any_div = any(div != 0), loc = div, rhats = rhats))

	# Plot figure r vs bole and fitted curve
	if (print_plot)
	{
		plot(forest[, bole_volume_m3], forest[, r], pch = 19, col = forest[, colour], axes = FALSE,
			xlab = "Bole volume", ylab = "Ratio bole/tot volumes", lwd = 0, cex = 0.75,
			main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
		if (!is.null(selected_variable))
			legend("topright", fill = colours, title = paste(selected_variable, "classes"),
				legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
		curve(r_func(x, paramsVec), add = TRUE, lwd = 3)
		axis(1)
		axis(2, las = 1)
	}

	# Plot predictions
	if (pred)
	{
		sim = apply(X = fit$draws("v_gen_mean"), MARGIN = 3, FUN = mean)

		if (print_plot)
		{
			plot(sim, forest[, total_volume_m3], pch = 19, col = forest[, colour], axes = FALSE,
				xlab = "Predicted total volume", ylab = "Observed total volume",
				main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
			if (!is.null(selected_variable))
				legend("topright", fill = colours, title = paste(selected_variable, "classes"),
					legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
			abline(a = 0, b = 1, lty = "dashed", lwd = 1.5, col = "#2E2E2E")
			axis(1)
			axis(2, las = 1)

			plot(forest[, bole_volume_m3], forest[, total_volume_m3], pch = 19, col = forest[, colour],
				axes = FALSE, xlab = "Observed bole volume", ylab = "Observed total volume",
				main = ifelse(is.null(selected_variable), sp, paste(sp, "(", selected_variable, ")")))
			if (!is.null(selected_variable))
				legend("topright", fill = colours, title = paste(selected_variable, "classes"),
					legend = levels(forest[, cut(.SD[[selected_variable]], breaks = n_bins, dig.lab = 1)]))
			abline(a = 0, b = 1, lty = "dashed", lwd = 0.85, col = "#2E2E2E")
			curve(psi(x, paramsVec), add = TRUE, lwd = 3)
			axis(1)
			axis(2, las = 1)
		}
	}

	if (!print_plot & pred)
	{
		return (list(any_div = any(div != 0), loc = div, rhats = rhats, pred_tot = sim,
			obs_bole = forest[, bole_volume_m3], obs_tot = forest[, total_volume_m3], params = paramsVec))
	}

	print("Hello 3")

	# Plot chains if any rhat problem
	if (any(rhats > 1.05))
	{
		if (!simplif)
		{
			lazyTrace(fit$draws("c"), main = paste0(sp, ", c"))
			lazyTrace(fit$draws("j"), main = paste0(sp, ", j"))
			lazyTrace(fit$draws("k"), main = paste0(sp, ", k"))
			lazyTrace(fit$draws("m"), main = paste0(sp, ", m"))
			lazyTrace(fit$draws("n"), main = paste0(sp, ", n"))
			lazyTrace(fit$draws("s"), main = paste0(sp, ", s"))

			paramsVec1 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 1)
			paramsVec2 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 2)
			paramsVec3 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 3)
			paramsVec4 = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"),
				type = "chain", chain = 4)
		} else {
			lazyTrace(fit$draws("alpha"), main = "alpha")
			lazyTrace(fit$draws("beta_"), main = "beta_")
			lazyTrace(fit$draws("gamma"), main = "gamma")
			lazyTrace(fit$draws("delta"), main = "delta")

			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 1)
			paramsVec1 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 2)
			paramsVec2 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 3)
			paramsVec3 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
			paramsVec_simplif = getParams(model_cmdstan = fit, params_names = c("alpha", "beta_", "gamma", "delta"),
				type = "chain", chain = 4)
			paramsVec4 = c(
				c = paramsVec_simplif["alpha"],
				j = 1,
				k = paramsVec_simplif["beta_"],
				m = exp(-1)*paramsVec_simplif["gamma"]/paramsVec_simplif["beta_"] + paramsVec_simplif["alpha"],
				n = paramsVec_simplif["delta"] + paramsVec_simplif["alpha"],
				s = paramsVec_simplif["beta_"]
			)
		}
		plot(forest[, bole_volume_m3], forest[, r], pch = 19, axes = FALSE, lwd = 0,
			xlab = "Bole volume", ylab = "Ratio bole/tot volumes", main = sp)
		curve(r_func(x, paramsVec), add = TRUE, lwd = 3)
		curve(r_func(x, paramsVec1), lwd = 3, add = TRUE, col = met.brewer("Egypt")[1])
		curve(r_func(x, paramsVec2), lwd = 3, add = TRUE, col = met.brewer("Egypt")[2])
		curve(r_func(x, paramsVec3), lwd = 3, add = TRUE, col = met.brewer("Egypt")[3])
		curve(r_func(x, paramsVec4), lwd = 3, add = TRUE, col = met.brewer("Egypt")[4])
		axis(1)
		axis(2, las = 1)
	}

	return (list(any_div = any(div != 0), loc = div, rhats = rhats))
}

plot_divergences = function(fit, div, simplif = FALSE)
{
	params = getParams(model_cmdstan = fit, params_names = c("c", "j", "k", "m", "n", "s"), type = "all")

	div_c = posterior::subset_draws(params[, , "c"], draw = which(div == 1))
	div_j = posterior::subset_draws(params[, , "j"], draw = which(div == 1))
	div_m = posterior::subset_draws(params[, , "m"], draw = which(div == 1))
	div_k = posterior::subset_draws(params[, , "k"], draw = which(div == 1))
	div_s = posterior::subset_draws(params[, , "s"], draw = which(div == 1))

	# Plot divergence for c and j
	plot(params[, , "c"], params[, , "j"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "j", main = "c and j")
	points(div_c, div_j, pch = 18, col = "#0F7BA2")

	# Plot divergence for c and k
	plot(params[, , "c"], params[, , "k"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "k", main = "c and k")
	points(div_c, div_k, pch = 18, col = "#0F7BA2")

	# Plot divergence for c and m
	plot(params[, , "c"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "c", ylab = "m", main = "c and m")
	points(div_c, div_m, pch = 18, col = "#0F7BA2")

	# Plot divergence for j and k
	plot(params[, , "j"], params[, , "k"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "j", ylab = "k", main = "j and k")
	points(div_j, div_k, pch = 18, col = "#0F7BA2")

	# Plot divergence for j and m
	plot(params[, , "j"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "j", ylab = "m", main = "j and m")
	points(div_j, div_m, pch = 18, col = "#0F7BA2")

	# Plot divergence for k and m
	plot(params[, , "k"], params[, , "m"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "k", ylab = "m", main = "k and m")
	points(div_k, div_m, pch = 18, col = "#0F7BA2")
	
	# Bonus plots of susceptible correlated parameters: k and s
	plot(params[, , "k"], params[, , "s"], pch = 19, cex = 0.65, col = "#FAB255",
		xlab = "k", ylab = "s", main = "k and s")
	points(div_k, div_s, pch = 18, col = "#0F7BA2")
}
```

```{r}
#| output: false

#### Load the data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

opt = "_inra"

# Loading Emerge data
inra = readRDS(paste0(path_data, "nfi-inra_dt", opt, ".rds"))
inra = inra[, .(speciesName_sci, tree_id = unique_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = taper_height_flo, bole_volume_m3, total_volume_m3)]
setkey(inra, speciesName_sci)

# Loading Swiss data
swiss = readRDS(paste0(mnt_point, "data/switzerland.rds"))
swiss = unique(swiss[, .(speciesName_sci, tree_id, plot_id, fct_type, year, circumference_m, height,
	taper_height = hdec, bole_volume_m3, total_volume_m3)])
setkey(swiss, speciesName_sci)

## Modern Emerge data
emerge = readRDS(paste0(mnt_point, "data/emerge_2009-2010.rds"))
emerge[, year := as.integer(stringi::stri_replace_all(str = dataset, replacement = "", regex = "emerge_"))]
emerge = emerge[, .(speciesName_sci, tree_id, fct_type, year, circumference_m, height, taper_height,
	bole_volume_m3 = bole_volume_conic_m3, total_volume_m3)]

## Bind everything
tree_dt = rbindlist(list(inra = inra, swiss = swiss, emerge = emerge), idcol = "origin", fill = TRUE)

# Add colours France vs Swiss
ls_colours = met.brewer("Egypt")[1:4]
names(ls_colours) = c("fr_comp", "fr", "ch_comp", "ch")
tree_dt[, colour := ls_colours["fr"]] # Inra and Emerge, in blue
tree_dt[origin == "swiss", colour := ls_colours["ch"]] # Swiss, in yellow

# Remove NAs (bole volumes only, I do not car if circumference or height have NAs)
tree_dt = tree_dt[!is.na(bole_volume_m3)]

# Set key to sort by species and colour (i.e., France vs Switzerland)
setkey(tree_dt, speciesName_sci, colour)

## Prepare species groups
# Load data
sp_sci = readRDS("../../gohon_florence/species.rds")[, .(tarif, speciesName_sci)]
sp_code = readRDS("../../gohon_florence/espar.rds")[, .(ess, Group_ess, espar, nom_espar)]

# Merging
sp_code = merge.data.table(sp_code, sp_sci, by.x = "ess", by.y = "tarif", all.x = TRUE)

# Unambiguous matching (done by chatGPT and checked manually)
unambiguous = data.table(
	speciesName_sci = c(
		"Abies nordmanniana",
		"Alnus sp.",
		"Cedrus atlantica",
		"Chamaecyparis sp.",
		"Eucalyptus sp.",
		"Fraxinus sp.",
		"Pinus cembra",
		"Pinus laricio",
		"Pinus laricio",
		"Pinus mugo",
		"Pinus strobus",
		"Pinus uncinata",
		"Quercus coccinea",
		"Tilia sp.",
		"Tilia sp.",
		"Tilia sp."
	),
	nom_espar = c(
		"Sapin de Nordmann",
		"Aulne vert",
		"Cèdre de l'Atlas",
		"Cyprès de Lawson",
		"Eucalyptus (genre)",
		"Frêne d'Amérique",
		"Pin cembro",
		"Pin laricio de Corse",
		"Pin laricio de Calabre",
		"Pin mugo",
		"Pin Weymouth",
		"Pin à crochets",
		"Chêne écarlate",
		"Tilleul d'Amérique du Nord",
		"Tilleul argenté",
		"Tilleul vert"
	),
	key = "speciesName_sci"
)

# Joining sp_code with unambiguous
sp_code = sp_code[unambiguous, speciesName_sci := fcoalesce(speciesName_sci, i.speciesName_sci),
	on = "nom_espar"]

# Species from tree_dt without group
ls_species = tree_dt[, unique(speciesName_sci)]
no_group = ls_species[!(ls_species %in% sp_code[, speciesName_sci])]

ls_genra = unique(stri_sub(str = no_group, to = stri_locate_first(str = no_group, regex = " ")[, "start"] - 1))
sp_code[, genra := stri_sub(str = speciesName_sci,
	to = stri_locate_first(str = speciesName_sci, regex = " ")[, "start"] - 1)]

matching = unique(sp_code[genra %in% ls_genra, .(genra, Group_ess)])
matching[, n_choices := .N, by = genra]

## Assigning groups manually
# For species that have only one choice anyway
manual_groups = data.table(speciesName_sci = no_group,
	genra = stri_sub(str = no_group, to = stri_locate_first(str = no_group, regex = " ")[, "start"] - 1),
	Group_ess = "", key = "genra")

manual_groups[matching[n_choices == 1], Group_ess := i.Group_ess, on = "genra"]

# For Pinus halepensis, Pinus sp., and Quercus sp.
manual_groups[genra == "Pinus", Group_ess := "B2"] # Rk: Pinus halepensis was listed as Pinus halepensis/Brutia
manual_groups[genra == "Quercus", Group_ess := "E2"] # Rk: Quercus sp. contains petraea, robur, pubescens

## Assigning groups to tree_dt
final_groups = unique(rbindlist(list(sp_code[!is.na(speciesName_sci), .(speciesName_sci, group = Group_ess)],
	manual_groups[,.(speciesName_sci, group = Group_ess)])))

if (!all(ls_species %in% final_groups[, unique(speciesName_sci)]))
	stop("Some species have no group! Do it manually, or whatever, bu do it! Just do it")

tree_dt = merge.data.table(tree_dt, final_groups, by = "speciesName_sci")
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

This script try to parametrise the total volume in function of the bole volume, that is to say, to use the basic definition of allometry: the dimension of an organ $y$ is proportional to an organ $x$, such as in this power-law $y = a x^b$. Here, $y$ is the total volume and $x$ is the bole volume. In @Yang2024, they do it with above-ground biomass as $y$ and stem biomass as $x$. I expect to be able to apply the same type of functions to volume, as volume and biomass are related by wood density. It should be the same if wood density is constant between the crown and the bole, and only the scale should change! According to @Knapic2011 there is little variation within tree oaks, while there can be for pines @Repola2006.

In this script, I parametrise species groups rather than species-specific models. The groups have been defined during the Emerge project (see @Deleuze2014 or @Longuetaud2013). For more details on the functional forms of the ratio, please check `11_yang.qmd`.

## Checking overlap of ratios from French and Swiss data
 
```{r}
```
