---
title: "Comparison with Cedric's data"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\hdec}{h_{\text{dec}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(stringi)
```

```{r}
#### Loading data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading training dataset
training_dt_sp_specific = readRDS(paste0(path_data, "data_Vtot.rds"))
training_dt = readRDS(paste0(path_data, "data_Vtot-group.rds"))

## Get species name (sci) from NFI codes
nfi_codes = readRDS(paste0(path_data, "species_dt.rds"))[!is.na(species_code_nfi),
	.(speciesName_sci, species_code_nfi)]
setkey(nfi_codes, speciesName_sci)

## Cedric's data
cedric = fread("./cedric/QSM_Volume_Ratios_394_Trees.csv")
cedric = cedric[, .(plot_id = NPP, tree_id = A, tree_name = Tree_Name,
	species_code = espar, species_fr = nom,
	dbh = DBH_QSM, c135_mm, height = TreeHeight,
	bole_volume = Total_Volume_trunk_7cm, bole_volume_nfi = v,
	total_volume = TotalVolumeTree, total_volume_nfi = v0)]

cedric = cedric[!is.na(bole_volume_nfi)]
```

```{r}
#### Function from Vallet
## Table 4 from Vallet2006
vallet_dt = data.table(speciesName_sci = c("Abies alba", "Fagus sylvatica", "Picea abies", "Pinus pinaster", "Pinus sylvestris", "Pseudotsuga menziesii", "Quercus petraea"),
	alpha = c(0.550, 0.395, 0.631, 0.235, 0.297, 0.534, 0.471),
	sd_alpha = c(0.015, 0.01, 0.007, 0.041, 0.025, 0.01, 0.014),
	beta = c(-7.49e-4, 2.66e-4, -9.46e-4, 9.7e-4, 3.18e-4, -5.3e-4, -3.45e-4),
	sd_beta = c(3.9e-5, 4.9e-5, 7.2e-5, 4.06e-4, 1.4e-4, 9.7e-5, 1.3e-5),
	gamma = c(0.277, 0.421, NA, 0.396, 0.384, NA, 0.377),
	sd_gamma = c(0.034, 0.025, NA, 0.057, 0.058, NA, 0.031),
	delta = c(NA, 45.4, NA, 198.8, 204.0, 56.6, NA),
	sd_delta = c(NA, 4, NA, 40, 26.6, 13.8, NA),
	var_res = c(0.0031, 0.0036, 0.0023, 0.0079, 0.0028, 0.0041 , 0.004), key = "speciesName_sci")

## Function to predict volumes
form_vallet = function(circumference_cm, height, pars, volume = TRUE)
{
	# print(pars)
	res = pars[, alpha] + pars[, beta]*circumference_cm
	if (!is.na(pars[, gamma]))
		res = res + pars[, gamma]*sqrt(circumference_cm)/height
	
	if (!is.na(pars[, delta]))
		res = res*(1 + pars[, delta]/circumference_cm^2)

	if (volume)
		res = res/(40000*pi) * circumference_cm^2*height
	
	return (res)
}
```


```{r}
#### Function from my work based in Yang 2024
## Table 4 from Vallet2006
params_dt = readRDS("./yang/avg_params.rds")

## Function to predict volumes
amael2026 = function(x, pars) # x = bole_volume_m3
{
	r_func = function(x, pars)
	{
		return ((pars[, m] - pars[, c]) * exp(pars[, j] - pars[, k]*x) * (pars[, k]*x/pars[, j])^pars[, j] +
			pars[, c] - (pars[, c] - pars[, n])*exp(-pars[, s]*x));
	}
	
	r = r_func(x, pars)

	return ( 1/pars[, c] * x^( 1 - (log(r) - log(pars[, c])) / log(x) ) )
}
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Données de Cédric

QSM = Quantitative Structure Model, a 3D geometric model of a tree reconstructed from terrestrial laser scanning (TLS) point clouds.

J'ai remarqué qu'en fait, chaque arbre est dupliqué de 14 à 16 fois pour *Fagus sylvatica* et *Quercus petraea*. Est-ce dû à différentes tentatives de segmentation de l'arbre? Dans tous les cas, je calcule la moyenne des grandeurs estimées pour chaque arbre (`dbh`, volumes, etc).

```{r}
#### Average within individuals
## Keep records of the range for each tree
range_dt = cedric[, as.list(quantile(total_volume/1000, probs = c(0, 0.05, 0.5, 0.95, 1))), by = tree_name]

## Average within each individual
cedric[, dbh := mean(dbh), by = tree_name]
cedric[, height := mean(height), by = tree_name]
cedric[, bole_volume := mean(bole_volume)/1000, by = tree_name]
cedric[, total_volume := mean(total_volume)/1000, by = tree_name] # /1000 conversion to cubic metres

cedric = unique(cedric)

cedric[, dbh*pi - c135_mm/1000] # Compare estimated QSM dbh with measured by NFI (I guess)

cedric = merge.data.table(cedric, range_dt, by = "tree_name")
```

```{r}
cedric[, .N, by = species_fr]
cedric[, .N, by = tree_name]
cedric = cedric[species_fr %in% c("Chene sessile", "Hetre")]

cedric[, speciesName_sci := ifelse(species_fr == "Hetre", "Fagus sylvatica", "Quercus petraea")]
setkey(cedric, speciesName_sci, plot_id, tree_id)
```

```{r}
plot(cedric[, bole_volume], cedric[, bole_volume_nfi], pch = 19, cex = 0.65,
	axes = FALSE, xlab = "QSM bole volume", ylab = "NFI bole volume")
abline(a = 0, b = 1, lty = "dashed", lwd = 0.8, col = "#515151")
axis(1)
axis(2, las = 1)
```

```{r}
cedric[, amael := amael2026(bole_volume, params_dt[speciesName_sci]), by = speciesName_sci]
cedric[, vallet := form_vallet(circumference_cm = c135_mm/10, height = height,
	pars = vallet_dt[speciesName_sci, .(alpha, beta, gamma, delta)]), by = speciesName_sci]
```

```{r}
plot(cedric[, total_volume], cedric[, vallet], pch = 19, cex = 1, col = "#FAB255",
	axes = FALSE, xlab = "QSM total volume", ylab = "Predicted volume",
	ylim = c(0, cedric[, max(vallet, amael)]))
points(cedric[, total_volume], cedric[, amael], pch = 19, cex = 1, col = "#0F7BA2")
abline(a = 0, b = 1, lty = "dashed", lwd = 0.8, col = "#515151")
legend("topleft", legend = c("Vallet", "Amael"), fill = c("#FAB255", "#0F7BA2"), bty = "n")
axis(1)
axis(2, las = 1)
```

## Application des volumes @Vallet2006

Dans cette section, j'applique les équations de Vallet afin de prédire le volume aerien total sur les données Emerge/Suisse.

```{r}
#### Comput Vallet volumes
## Predict
training_dt = training_dt[speciesName_sci %in% vallet_dt[, speciesName_sci]]
training_dt = training_dt[!is.na(circumference_m) &  !is.na(height)]
training_dt[, vallet_pred := form_vallet(circumference_cm = 100*circumference_m, height = height,
	pars = vallet_dt[speciesName_sci, .(alpha, beta, gamma, delta)]), by = speciesName_sci]

training_dt[, r_ipcc_vallet := vallet_pred/bole_volume_m3]
training_dt[, r_ipcc := total_volume_m3/bole_volume_m3]
```

```{r}
dd_ipcc = density(training_dt[, r_ipcc])
dd_vallet = density(training_dt[, r_ipcc_vallet])

ls_colours = met.brewer("Egypt")[1:2]
names(ls_colours) = c("ipcc", "vallet")

plot(dd_ipcc, col = ls_colours["ipcc"], axes = FALSE, lwd = 3, xlab = "Ratio IPCC", main = "",
	xlim = c(0, 5), ylim = c(0, max(dd_ipcc$y, dd_vallet$y)))
polygon(dd_ipcc, col = paste0(ls_colours["ipcc"], 44))

lines(dd_vallet, col = ls_colours["vallet"])
polygon(dd_vallet, col = paste0(ls_colours["vallet"], 44))

legend("topright", legend = c("Measured", "Vallet 2006"), fill = ls_colours, bty = "n")

axis(1)
axis(2, las = 1)

training_dt[, range(r_ipcc)]
training_dt[, range(r_ipcc_vallet)]

training_dt[, as.list(quantile(r_ipcc, probs = c(0, 0.005, 0.5, 0.995, 1))), by = fct_type]
training_dt[, as.list(quantile(r_ipcc_vallet, probs = c(0, 0.005, 0.5, 0.995, 1))), by = fct_type]
```