---
title: "Comparison with Vallet and output preparation for pgfplots"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: true
  echo: false
  warning: false
  message: false
bibliography: /home/ALe-Squin/work/library/bib_file/references.bib
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

<!-- New command -->
\newcommand{\ie}{*i.e.,*}

<!-- Math commands -->
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\Vtot}{V_{\text{tot}}}
\newcommand{\Vzh}{V_{\text{Zhou}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(stringi)
library(loo)
library(gt)
```

```{r}
#### Load data
## Path
os = Sys.info()[['sysname']]
mnt_point = "/mnt/local_share/"
if (os == "Linux" || os == "Darwin")
{
	if (!dir.exists(mnt_point))
		stop(paste0("The mounting point <", mnt_point, "> does not exist"))
} else if (os == "Windows") {
	stop("TO DO!!! No idea how that works on Windows!")
} else {
	stop(paste("Unknown Operating System:", os))
}

path_data = paste0(mnt_point, "data/")
if (!dir.exists(path_data))
	stop(paste0("Folder <", path_data, "> does not exist! Did you run 01_prepare_data.qmd?"))

## Loading training dataset
training_dt = readRDS(paste0(path_data, "data_Vtot-group.rds"))[origin %in% c("inra", "emerge")][
	!is.na(circumference_m) & !is.na(height)]
```

## Introduction

@Longuetaud2013 does not give any average VEF values! I propose to compute the average $m$ of her species-specific functions with:
$$
	m = \frac{1}{S} \int\int_S f(x) \, dx,
$$
where $S = [d_{025}, d_{975}] \times [h_{025}, h_{975}] $ is the surface delimited by the species-specific dbh and height ranges (from her table 1).

## Compute average

```{r}
#### Prepare data
## Compute genus-specific quantiles
training_dt[, genus := tolower(
	stri_sub(speciesName_sci, to = stri_locate_first(speciesName_sci, regex = " ")[, "start"] - 1))]
training_dt[, dbh_cm := 100*circumference_m/pi]
genus = training_dt[, as.list(lapply(.SD, quantile, probs = c(0.025, 0.975))), by = "genus",
	.SDcols = c("dbh_cm", "height")]
genus[, vars := rep(c("q025", "q975"), .N/2)]

genus = dcast.data.table(data = genus, formula = genus ~ vars, value.var = c("dbh_cm", "height"))

## Parameters and range, verified
longuetaud_pars = data.table(
	genus = c("betula", "carpinus", "fagus", "fraxinus", "quercus",
		"abies", "larix", "picea", "pinus", "pseudotsuga"),
	b1 = rep(6.83, 10),
	b2 = c(1.016, 1.133, 0.842, 1.112, 1.320,
		1.496, 0.966, 0.953, 0.566, 0.744),
	b3 = rep(1.009, 10),
	b4 = c(0.448, 0.822, 0.627, 0.470, 0.245,
		0.566, 0.454, 0.473, 0.336, 0.783),
	min_d = c(12, 8, 7, 8, 7,
		7, 9, 8, 7, 7),
	max_d = c(31, 38, 79, 43, 102,
		90, 42, 67, 71, 34),
	min_h = c(14.7, 11.3, 5.5, 11.5, 6,
		6.9, 9, 7.7, 5.1, 7),
	max_h = c(28.5, 30, 42.5, 26.7, 40,
		37.8, 25.6, 41.3, 33.4, 29.5),
	is_broadleaf = rep(c(TRUE, FALSE), each = 5),
	key = "genus"
)

longuetaud_pars = merge.data.table(longuetaud_pars, genus, by = "genus")
```

```{r}
vef_fct = function(dbh, height, pars)
{
	if (!is.data.table(pars))
		stop("Please provide a data table!")
	
	b1 = pars[, b1]
	b2 = pars[, b2]
	b3 = pars[, b3]
	b4 = pars[, b4]
	
	int_broadleaf = ifelse(pars[, is_broadleaf], 1, 0)

	vef = 1 + exp(b2 * (b1 - dbh)) + exp(b3*int_broadleaf + b4)*dbh/height^2
	
	return(vef)
}

vef_mean = function(pars)
{
	f = function(dbh, height)
		vef_fct(dbh, height, pars)

	res = pracma::integral2(
		fun = f,
		xmin = pars$dbh_cm_q025,
		xmax = pars$dbh_cm_q975,
		ymin = pars$height_q025,
		ymax = pars$height_q975
	)

	areaDomain = (pars$dbh_cm_q975 - pars$dbh_cm_q025) * (pars$height_q975 - pars$height_q025)

	return(res$Q/areaDomain)
}

curve(vef_fct(x, height = 20, pars = longuetaud_pars["abies"]), from = 7, to = 50)

longuetaud_pars[, mean_vef := vef_mean(.SD), by = genus]

if (!file.exists("./longuetaud_VEF.rds"))
	saveRDS(longuetaud_pars, "./longuetaud_VEF.rds")

filename = "../../../writing/article_allometry/data/longuetaud_VEF.csv"
if (!file.exists(filename))
	fwrite(longuetaud_pars, filename)
```