---
title: "Reading @Zhou2021"
date: today
author: Amaël Le Squin
date-format: iso
filters:
  - fontawesome
  - siunitx-quarto
  - imagify
imagify:
  header-includes: |
    \usepackage{pgfplots}
    \pgfplotsset{compat=1.18}
    \usepgfplotslibrary{fillbetween}
    \usetikzlibrary{positioning}
    \usetikzlibrary{calc}
  pdf-engine: lualatex
pdf-engine: lualatex
keep-tex: true
execute:
  error: false
  echo: false
bibliography:	
  - "/home/ALe-Squin/work/library/bib_file/references.bib"
lightbox:
  match: auto
css: style.css
knitr:
  opts_chunk: 
    dev: ragg_png
    crop: null
    out.width: "70%"
    fig.width: 6
    fig.asp: 0.618
    fig.align: "center"
format:
  html:
    toc: true
    include-in-header: mathjax.html
    code-fold: true
    df-print: paged
    number-sections: true
    theme:
      light: cerulean
      dark: darkly
    margin: 5% 0;
  pdf:
    mathspec: true
    include-in-header:
      - text: |
          \usepackage{unicode-math}
---

\newcommand{\ie}{\textit{i.e., }}
\newcommand{\Vbft}{V_{\text{BFT}}}
\newcommand{\Fbft}{F_{\text{BFT}}}
\newcommand{\hdec}{h_{\text{dec}}}
\newcommand{\fmax}{f_{\text{max}}}

<!-- Math operators -->
<!-- \DeclareMathOperator{\ln}{ln} -->

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   LOADING DATA, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

```{r}
#| output: false

#### Clear space and load packages
rm(list = ls())
graphics.off()

options(max.print = 500)

library(data.table)
library(MetBrewer)
library(cmdstanr)
	register_knitr_engine(override = TRUE)
library(pdftools)
library(stringi)
library(DHARMa)

source("./toolFunctions.R")
```

```{r}
#### Load appendix pdf file
pdf_file = pdf_text("~/work/library/suppInf/Zhou2021/app_1.pdf")

#### Read the field data for modelling allometric relationship
n_l = 294

lanceolata_dt = data.table(plot_id = integer(n_l), lon = numeric(n_l), lat = numeric(n_l),
	temp = numeric(n_l), precip = integer(n_l), sb = numeric(n_l), agb = numeric(n_l))

counter = 0

for (p in 2:7) # Pages of lanceolata data
{
	current_page = stri_split(str = pdf_file[p], regex = "\n")
	start = 1
	end = length(current_page[[1]])
	if (p == 2)
	{
		while(start < length(current_page[[1]]) && !stri_detect(str = current_page[[1]][start], regex = "Plot No."))
			start = start + 1
		start = start + 1 # To get rid of "Plot No."
	}

	if (p == 7)
	{
		end = start
		while(end < length(current_page[[1]]) &&
			!stri_detect(str = current_page[[1]][end], regex = "2. Field data for model test"))
			end = end + 1
		end = end - 1 # To remove the line "2. Field data for model test"
	}

	current_page = current_page[[1]][start:end]
	for (r in seq_along(current_page)) # r is the row
	{
		current_line = stri_split(str = current_page[r], regex = " ", simplify = TRUE)
		current_line = stri_split(str = current_line, regex = "\t", simplify = FALSE)
		current_line = unlist(current_line[current_line != ""])
		current_line = current_line[current_line != ""]

		if (length(current_line) == 7)
		{
			counter = counter + 1
			values = as.numeric(current_line)
			lanceolata_dt[counter, colnames(lanceolata_dt) := as.list(values)]
		}
	}
}

#### Read the Field data for model test
n = 307

forest_dt = data.table(plot_id = integer(n), dom_sp = character(n), dbh_cm = numeric(n), height_m = numeric(n),
	tree_dens = numeric(n), volume = numeric(n), sb = numeric(n), agb = numeric(n))

counter = 0

for (p in 7:14) # Pages of lanceolata data
{
	current_page = stri_split(str = pdf_file[p], regex = "\n")
	start = 1
	end = length(current_page[[1]])
	if (p == 7)
	{
		while(start < length(current_page[[1]]) && !stri_detect(str = current_page[[1]][start], regex = "Plot No."))
			start = start + 1
		start = start + 1 # To get rid of "Plot No."
	}

	current_page = current_page[[1]][start:end]
	for (r in seq_along(current_page)) # r is the row
	{
		current_line = stri_split(str = current_page[r], regex = " ", simplify = TRUE)
		current_line = stri_split(str = current_line, regex = "\t", simplify = FALSE)
		current_line = unlist(current_line[current_line != ""])
		current_line = current_line[current_line != ""]

		if (length(current_line) == 9)
		{
			counter = counter + 1
			values_num = suppressWarnings(as.numeric(current_line)) # NAs introduced by coercion
			sp = paste(current_line[2:3], collapse = " ")

			forest_dt[counter, colnames(forest_dt) := c(list(values_num[1]), list(sp), as.list(values_num[4:9]))]
		}
	}
}

# Get genus (and group the rare ones)
forest_dt[, genus := stri_split(str = dom_sp, regex = " ", simplify = TRUE)[, 1]]
forest_dt[!(genus %in% c("Abies", "Cunninghamia", "Picea", "Quercus", "Robinia")), genus := "Evergreen"]

forest_dt[genus %in% c("Abies", "Picea"), genus := "Abies and Picea"]
forest_dt[genus %in% c("Quercus", "Robinia"), genus := "Quercus and Robinia"]

setkey(forest_dt, genus)

# Assin colours and pch
ls_colours = met.brewer("Egypt")[1:4]
names(ls_colours) = c("Abies and Picea", "Evergreen", "Cunninghamia", "Quercus and Robinia")

ls_pch = c(0, 20, 3, 0)
names(ls_pch) = c("Abies and Picea", "Evergreen", "Cunninghamia", "Quercus and Robinia")

forest_dt[, colour := ls_colours[genus], by = genus]
forest_dt[, pch := ls_pch[genus], by = genus]
```

<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
<!-- ---------------------------   MAIN CONTENT, R CODE   --------------------------- -->
<!-- -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

## Introduction

The aim of this small note is to sum-up the study of @Zhou2021 and to reproduce part of it. The data can be found in the first appendix of @Zhou2021, in a pdf format. They have been loaded using the `pdftools` package.

## Reproducing the first figure of @Zhou2021

For the Fig. 1 (a), reproduced below (@fig-1a), they use the following linear model (intercept is 0):
$$
	\text{AGB} = \beta_1 \text{SB} + \beta_2 \text{SB}^2 + \beta_3 \text{SB}^3 + \beta_4 \text{SB}^4.
$$
The coefficients' values are identical to what was found in @Zhou2021.

```{r}
## Linear model
agb_mod = lm(agb ~ -1 + sb + I(sb^2) + I(sb^3) + I(sb^4), data = lanceolata_dt)
f_pred = function(x, coeff = coef(agb_mod))
	return (coeff %*% outer(1:4, x, function(p, x) x^p))
```

```{r}
#| label: fig-1a
## Plot
plot(lanceolata_dt[, sb], lanceolata_dt[, agb], axes = FALSE, pch = 20, col = "#FFAF37", xlab = "SB", ylab = "AGB")
axis(1)
axis(2, las = 1)
curve(f_pred, lwd = 2, col = "#CD212A", add = TRUE)
```

I do not know how Fig. 1 (b) was made... I used $k = 0.035$ and $p = 2$...

```{r}
## Fig. 1 (b)
psi = function(x, k, p)
{
	if (p < 1)
		stop("p must be larger than 1")
	
	if ((k < 0) || (k > 1))
		stop("k must be between 0 and 1")
	return (1 - k*x/(p*exp(k*x) - 1))
}

p = 2
k = 0.035

curve(psi(x, k, p), to = 200, xlab = "SB", ylab = "Allometric coeff", axes = FALSE, lwd = 2, col = "#FFAF37")
axis(1)
axis(2, las = 1)
abline(h = 1, lty = 2, col = "#212121")
```

For the Fig. 1 (d), it is noteworthy to say that they removed an extrem value of 1200 for the stand biomass (`plot_id = 267`, *Picea likiangensis*).
```{r}
## Fig. 1 (d)
x_max = forest_dt[sb < 800, max(sb)]

r_zhou = function(x, m, d, k)
	return (m - d*exp(-k*x))

plot(forest_dt[sb < 800, sb], forest_dt[sb < 800, sb/agb], xlab = "SB", lwd = 2,
	ylab = "Ratio SB to AGB (r)", axes = FALSE, col = forest_dt[, colour], pch = forest_dt[, pch])
axis(1)
axis(2, las = 1)
abline(h = 0.9, lty = 2, col = "#212121")
abline(h = 0.95, lty = 2, col = "#212121")
polygon(x = c(0, 0, x_max, x_max), y = c(0.9, 0.95, 0.95, 0.9), col = "#A2A2A244")
legend(x = "bottomright", legend = names(ls_colours), col = ls_colours, pch = ls_pch, bty = "n")

curve(r_zhou(x, 0.893, 0.378, 0.0076), col = ls_colours["Abies and Picea"], lwd = 2, add = TRUE)
curve(r_zhou(x, 0.905, 0.453, 0.0212), col = ls_colours["Cunninghamia"], lwd = 2, add = TRUE)
curve(r_zhou(x, 0.890, 0.420, 0.0166), col = ls_colours["Evergreen"], lwd = 2, add = TRUE)
curve(r_zhou(x, 0.917, 0.237, 0.0053), col = ls_colours["Quercus and Robinia"], lwd = 2, add = TRUE)

# Extract the min, max, and diff of the ratio (m, n, and d params on Fig 1.c)
bounds_dt = forest_dt[, .(m = min(sb/agb), M = max(sb/agb)), by = genus]
bounds_dt[, d := M - m]

bounds_dt[, c("M_print", "d_print") := lapply(X = .SD, FUN = round, digit = 3), by = genus, .SD = c("M", "d")]
```

## Estimating the parameters {#sec-params}
The equation 2 in @Zhou2021 represents the AGB in function of SB:
$$
	\text{AGB} = \frac{\text{SB}}{m - d \exp[-k\text{SB}]},
$$ {#eq-AGB}
where $m$ is the maximum of the ratio SB over AGB, and $d$ is the ratio span (\ie max - min). In @Zhou2021, the parameter $k$ has been estimated with a linear regression, while $m$ and $d$ were obtained by optimising a cost function (Zhou Xiaolu, *discussion*). I try here to estimate all the parameters in a single model (see @tbl-table_1). It is noteworthy that estimating all three parameters simultaneously initially failed, highlighting the need for constraints and informative priors, and a relevant likelihood. The Gaussian is not adapted and was probably part of the failure of the first attempt. I now use a Beta distribution parametrised as follow:
\begin{align*}
	\text{shape}_1 &= \phi\mu, \\
	\text{shape}_2 &= \phi(1 - \mu),
\end{align*}
where $\mu$ is the mean (defined by @eq-AGB), and $\phi$ is the precision (well, kind of... not exactly the inverse of the variance; but it increases when the variance decreases).

To remove the effect of autocorrelation between SB and AGB (after all, SB is included in AGB), the authors defined a new variable, $y = \text{AGB} - \text{SB}$. Watch out, $y$ denotes both this new variable and the old AGB in the article! I did not do anything like that here, and it still works...

I use the following Bayesian model to estimate the parameters $m$, $k$, and $d$:

```{stan output.var = "model", cache = TRUE}
#| echo: true
data {
	// Dimensions
	int <lower = 1> N; // Number of trees
	int <lower = 1, upper = N> G; // Number of genus
	array[G] int <lower = 1, upper = N> start; // Genus index start
	array[G] int <lower = 1, upper = N> end; // Genus index end

	// Predictors
	vector[N] SB;

	// Data
	vector[N] AGB;
}

transformed data {
	vector[N] ratio = SB ./ AGB;
}

parameters {
	// Parameters k of Zhou2021 (Dynamic allometric scaling of tree biomass and size)
	array[G] real <lower = 0, upper = 0.3> k;
	array[G] real <lower = 0.7, upper = 1> m;
	array[G] real <lower = 0.05, upper = m> d;

	// Precision
	real <lower = 0> phi;
}

transformed parameters {
	vector [N] shape1;
	vector [N] shape2;

	for (i in 1:G)
	{
		shape1[start[i]:end[i]] = phi * (m[i] - d[i]*exp(-k[i]*SB[start[i]:end[i]]));
		shape2[start[i]:end[i]] = phi*(1 - (m[i] - d[i]*exp(-k[i]*SB[start[i]:end[i]])));
	}
}

model{
	// Prior linear regression
	target += normal_lpdf(k | 0, 0.1);
	target += normal_lpdf(m | 0.85, 0.05);
	target += normal_lpdf(d | 0.5, 0.15);
	target += gamma_lpdf(phi | 0.01, 0.01); // 0.25^2/0.03, 0.25/0.03

	// Likelihood
	target += beta_lpdf(ratio | shape1, shape2);
}

generated quantities {
	array[N] real r_gen = beta_rng(shape1, shape2);
}
```

```{r}
#| output: false

# Values from Zhou 2021
m_zhou = c("Abies and Picea" = 0.893, "Cunninghamia" = 0.905, "Evergreen" = 0.890, "Quercus and Robinia" = 0.917)
d_zhou = c("Abies and Picea" = 0.378, "Cunninghamia" = 0.453, "Evergreen" = 0.420, "Quercus and Robinia" = 0.237)

## Run model
# Common variables
n_chains = 4

stanData = list(
	N = forest_dt[, .N],
	G = length(ls_colours),
	start = forest_dt[, .(start = .I[1]), by = genus][, start],
	end = forest_dt[, .(end = .I[.N]), by = genus][, end],

	# Parameters from data
	m = bounds_dt[, M], # m_zhou, bounds_dt[, M]
	d = bounds_dt[, d], # d_zhou, bounds_dt[, d]

	# Predictors
	SB = forest_dt[, sb],

	# Data
	AGB = forest_dt[, agb]
)

# Run
fit = model$sample(data = stanData, chains = n_chains, parallel_chains = min(n_chains, 4))

# Get k params
k_vec = apply(X = fit$draws("k"), MARGIN = 3, FUN = mean)
names(k_vec) = forest_dt[, unique(genus)] # Using this is safe, genus is a key

k_vec_print = signif(k_vec, digit = 3)

# Get m params
m_vec = apply(X = fit$draws("m"), MARGIN = 3, FUN = mean)
names(m_vec) = forest_dt[, unique(genus)] # Using this is safe, genus is a key

m_vec_print = signif(m_vec, digit = 3)

# Get d params
d_vec = apply(X = fit$draws("d"), MARGIN = 3, FUN = mean)
names(d_vec) = forest_dt[, unique(genus)] # Using this is safe, genus is a key

d_vec_print = signif(d_vec, digit = 3)
```

## Results

### Posteriors for the parameter $m$

An important difference with @Zhou2021 is that I constrained $m$ to be between 0.7 and 1 (see stan model in @sec-params), which is a much larger interval than theirs (restricted to 0.9 to 0.95---Xiaolu Zhou, *discussion*). I did it because I noticed that with a restricted interval, some values of $m$ had a tendency to accumulate towards the lower bound (0.9).

```{r}
#| label: fig-posterior_m
#| fig-cap: Posterior of $m$ for the four groups
#| fig-subcap:
#|   - "Posterior of $m$ for Abies and Picea"
#|   - "Posterior of $m$ for Cunninghamia"
#|   - "Posterior of $m$ for Evergreen"
#|   - "Posterior of $m$ for Quercus and Robinia"
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 4
#| output: false

lazyPosterior(draws = fit$draws("m[1]"), fun = dnorm, mean = 0.85, sd = 0.05)
lazyPosterior(draws = fit$draws("m[2]"), fun = dnorm, mean = 0.85, sd = 0.05)
lazyPosterior(draws = fit$draws("m[3]"), fun = dnorm, mean = 0.85, sd = 0.05)
lazyPosterior(draws = fit$draws("m[4]"), fun = dnorm, mean = 0.85, sd = 0.05)
```

### Posteriors for the parameter $d$

The parameter $d$, which governs the span of the data and therefore will influence both steepness and maximum, is the one that shows the most uncertainty. I have not tried, but maybe I could rewrite the mean ratio, $\mu$ (see @eq-AGB) differently:

$$
	\mu = m - (m - y_0) \exp[-k\text{SB}],
$$
where $y_0$ would be the minimum.

```{r}
#| label: fig-posterior_d
#| fig-cap: Posterior of $d$ for the four groups
#| fig-subcap:
#|   - "Posterior of $d$ for Abies and Picea"
#|   - "Posterior of $d$ for Cunninghamia"
#|   - "Posterior of $d$ for Evergreen"
#|   - "Posterior of $d$ for Quercus and Robinia"
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 4
#| output: false

lazyPosterior(draws = fit$draws("d[1]"), fun = dnorm, mean = 0.5, sd = 0.15)
lazyPosterior(draws = fit$draws("d[2]"), fun = dnorm, mean = 0.5, sd = 0.15)
lazyPosterior(draws = fit$draws("d[3]"), fun = dnorm, mean = 0.5, sd = 0.15)
lazyPosterior(draws = fit$draws("d[4]"), fun = dnorm, mean = 0.5, sd = 0.15)
```

### Posteriors for the parameter $k$

```{r}
#| label: fig-posterior_k
#| fig-cap: Posterior of $k$ for the four groups
#| fig-subcap:
#|   - "Posterior of $k$ for Abies and Picea"
#|   - "Posterior of $k$ for Cunninghamia"
#|   - "Posterior of $k$ for Evergreen"
#|   - "Posterior of $k$ for Quercus and Robinia"
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 4
#| output: false

lazyPosterior(draws = fit$draws("k[1]"), fun = dnorm, mean = 0, sd = 0.1)
lazyPosterior(draws = fit$draws("k[2]"), fun = dnorm, mean = 0, sd = 0.1)
lazyPosterior(draws = fit$draws("k[3]"), fun = dnorm, mean = 0, sd = 0.1)
lazyPosterior(draws = fit$draws("k[4]"), fun = dnorm, mean = 0, sd = 0.1)
```

::: {#tbl-table_1 .borderless .hover}

| species type        | m     | m (mine) | d     | d (mine) | k    | k (mine) |
|--------------------:|:------|:---------|:------|:---------|:-----|----------|
| Cunninghamia        | 0.905 | `{r} bounds_dt["Cunninghamia", M_print]` | 0.453 | `{r} bounds_dt["Cunninghamia", d_print]` | -0.0212 | `{r} k_vec_print["Cunninghamia"]` |
| Abies and Picea     | 0.893 | `{r} bounds_dt["Abies and Picea", M_print]` | 0.378 | `{r} bounds_dt["Abies and Picea", d_print]` | -0.0076 | `{r} k_vec_print["Abies and Picea"]` |
| Quercus and Robinia | 0.917 | `{r} bounds_dt["Quercus and Robinia", M_print]` | 0.237 | `{r} bounds_dt["Quercus and Robinia", d_print]` | -0.0053 | `{r} k_vec_print["Quercus and Robinia"]` |
| Evergreen     | 0.890 | `{r} bounds_dt["Evergreen", M_print]` | 0.420 | `{r} bounds_dt["Evergreen", d_print]` | -0.0166 | `{r} k_vec_print["Evergreen"]` |

Values from @Zhou2021 *vs* mine (defined as the mean of the posterior. Note that using the median would provide similar values)

:::

## Residuals (computed with DHARMa)

The residuals are simulated with the package `DHARMa`, and uses the posterior simulations from the stan model (see the block `generated quantities`). They are quite good, including when plotted agains SB.

```{r}
#| label: fig-dharma
#| fig-cap: Residuals computed with DHARMa
#| fig-subcap:
#|   - "Abies and Picea"
#|   - "Abies and Picea (with respect to SB)"
#|   - "Cunninghamia"
#|   - "Cunninghamia (with respect to SB)"
#|   - "Evergreen"
#|   - "Evergreen (with respect to SB)"
#|   - "Quercus and Robinia"
#|   - "Quercus and Robinia (with respect to SB)"
#| fig-asp: 1
#| out-width: 95%
#| fig-width: 8

n_sampling = fit$metadata()$iter_sampling

simResp = fit$draws("r_gen")
dim(simResp)

# Reshape simResp as a matrix nobs x n_repetition
simResp_matrix = matrix(data = 0, nrow = forest_dt[, .N], ncol = n_sampling*n_chains)
for (obs in seq_len(forest_dt[, .N]))
	for (chain in 1:n_chains)
		simResp_matrix[obs, ] = simResp[, chain, obs] # One line of n_sampling*n_chains simulations per observation

## Simulated residuals
for (i in seq_along(stanData[["start"]]))
{
	start = stanData[["start"]][i]
	end = stanData[["end"]][i]
	sim_dharma = createDHARMa(simulatedResponse = simResp_matrix[start:end, ],
		observedResponse = forest_dt[start:end, sb/agb],
		fittedPredictedResponse = apply(simResp[, , start:end], MARGIN = 3, median),
		integerResponse = FALSE)

	# testDispersion(sim_dharma)

	plot(sim_dharma, quantreg = TRUE)
	plotResiduals(sim_dharma, forest_dt[start:end, sb], quantreg = TRUE, xlab = "Stand biomass")
}
```

