% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\newcommand{\ie}{*i.e.,*}
\newcommand{\F}{\mathscr{F}}
\newcommand{\N}{\mathbfscr{N}}
\usepackage{unicode-math}
\usepackage{siunitx}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Testing the model of Westfall et al. (2023)},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\title{Testing the model of Westfall et al. (2023)}
\author{Amaël Le Squin}
\date{2024-12-13}

\begin{document}
\maketitle


\subsection{Introduction}\label{introduction}

The European Union developed a
\href{https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX:52021DC0572}{New
EU Forest Strategy for 2030} as part of the plan to adapt to and fight
against climate change and make Europe a climate neutral continent by
2050. This strategy relies on improved monitoring of European forests to
better understand their condition and respond accordingly. Specifically,
it calls for assessing carbon sequestration in forests to evaluate
whether or not Europe reached carbon neutrality. One bottleneck is the
harmonisation of the forest monitoring methods between European member
states, if not within them. The
\href{https://pathfinder-heu.eu/\#top}{PathFinder project} supports
member states in implementing a European Forest Monitoring System in
order to standardise or harmonise forest data collection and reporting
across the EU. This prompted the Institut National de l'information
Géographique et forestière (IGN), which is in charge of the French
National Forest Inventory (NFI) data, to update its methods for
assessing forest carbon storage.

In most countries, carbon content is estimated via two quantities:
(\emph{i}) the volume in \unit{\cubic\metre}, and (\emph{ii}) the wood
density in \unit{\kilogram\per\cubic\metre}. However, often numerous
models and methods are currently used across countries, contingent upon
the tree species and geographic location within each country (Westfall
et al. 2023; Longuetaud et al. 2013; Gregoire and Schabenberger 1996),
but also the definition of tree parts (Gschwantner et al. 2009).

\subsection{Methods}\label{methods}

The original models can be found in Westfall et al. (2023):

Model of Schumacher-Hall
\begin{equation}\phantomsection\label{eq-schumacherHall}{
y_i = a D_i^b H_i^c
}\end{equation}

Segmented model:

\begin{equation}\phantomsection\label{eq-cases}{
y_i =
\begin{cases}
    a D_i^b H_i^c & D_i < k \\
    a k^{b - b_1} D_i^{b_1} H_i^c & D_i \geqslant k
\end{cases}
}\end{equation}

Continuously variable model:
\begin{equation}\phantomsection\label{eq-continuous}{
y_i = a D_i^{a_1(1 - \exp[-b D_i])^c} H_i^c
}\end{equation}

and the modified Wiley model:
\begin{equation}\phantomsection\label{eq-wiley}{
y_i = a D_i^b H_i^c \exp[-b_1 D_i]
}\end{equation}

The Equation~\ref{eq-schumacherHall} is the default model,
\ie \(\F = \N = 2\).

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Gregoire1996}
Gregoire, Timothy G., and Oliver Schabenberger. 1996. {``A Non-Linear
Mixed-Effects Model to Predict Cumulative Bole Volume of Standing
Trees.''} \emph{Journal of Applied Statistics} 23 (2--3): 257--72.
\url{https://doi.org/10.1080/02664769624233}.

\bibitem[\citeproctext]{ref-Gschwantner2009}
Gschwantner, Thomas, Klemens Schadauer, Claude Vidal, Adrian Lanz, Erkki
Tomppo, Lucio di Cosmo, Nicolas Robert, Daisy Englert Duursma, and Mark
Lawrence. 2009. {``Common Tree Definitions for National Forest
Inventories in Europe.''} \emph{Silva Fennica} 43 (2).
\url{https://doi.org/10.14214/sf.463}.

\bibitem[\citeproctext]{ref-Longuetaud2013}
Longuetaud, Fleur, Philippe Santenoise, Frédéric Mothe, Tristan Senga
Kiessé, Michaël Rivoire, Laurent Saint-André, Nina Ognouabi, and
Christine Deleuze. 2013. {``Modeling Volume Expansion Factors for
Temperate Tree Species in France.''} \emph{Forest Ecology and
Management} 292: 111--21.
\url{https://doi.org/10.1016/j.foreco.2012.12.023}.

\bibitem[\citeproctext]{ref-Westfall2023}
Westfall, James A., John W. Coulston, Andrew N. Gray, John D. Shaw,
Philip J. Radtke, David M. Walker, Aaron R. Weiskittel, et al. 2023.
{``A National-Scale Tree Volume, Biomass, and Carbon Modeling System for
the United States.''} \emph{Gen. Tech. Rep. WO-104. Washington, DC: U.S.
Department of Agriculture, Forest Service.} 104.
\url{https://doi.org/10.2737/wo-gtr-104}.

\end{CSLReferences}




\end{document}
